### 2014-08-19 assets/paperdraft/paper_draft.tex
initial sketch of the article

\documentclass[12pt]{article}

% AMS packages:
\usepackage{amsmath, amsthm, amsfonts}
\usepackage[margin=0.5in]{geometry}

% Theorems
%-----------------------------------------------------------------

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}

\def\RR{\mathbb{R}}
\def\ZZ{\mathbb{Z}}
\newcommand{\abs}[1]{\left\vert#1\right\vert}
\newcommand{\sgn}{\operatorname{sgn}}

%-----------------------------------------------------------------
\title{New Approaches to Uniform Boosting }
\author{
}

\begin{document}
\maketitle

% \abstract{Abstract goes here}

\section{Introduction}

Methods of machine learning are playing a important role in particles physics nowadays, in particular, multivariate analysis (MVA). 
Different classifiers like boosted decision trees (BDTs) and artificial neural networks (ANNs) are often used as an important step in analysis selection criteria.

BDTs are now even used in software triggers [link]. The main point of boosting technique is training many simple classifiers and building a composition of their outputs.
Classifiers are trained one-after-another, their inputs are augmented in such a way that new classifier should target more at those events which were poorly classified by previous ones. The resulting classifier obtained by combining them often much more powerful, than individual ones.

In practice, however, there are some restrictions that trained classifier should meet (apart from having good classification quality, which is usually measured by some integrated FOM). For example, in an amplitude analysis obtaining a uniform efficiency in a multivariate space of physics variates, i.e., variates that are of physical interest, is often times more important than any integrated FOM based on the total amount of signal and background. Such analyses often have many variates in which a uniform efficiency is desired. A uniform efficiency reduces systematic uncertainties and helps maintain sensitivity to all hypotheses being tested.

The one approach that was already proposed is uBoost (uniform BOOSTing) --- a modification of AdaBoost algorithm [link]

Why do we need such algorithms:

\begin{enumerate}
	\item The result of classification is stable to check different hypotheses
	\item To get unbiased distribution of signal, this enables us to compute real mass of particle
	\item Not to get fake peak, thus not get a false discovery
\end{enumerate}

The features, along which we want the prediction to be flat, will be referred to as 
\textit{uniform variables}. In high energy physics these are usually masses.

\section{Uniformity Measurement}

In this section we come up with some approaches on how to measure uniformity of prediction. The typical way of 'checking' uniformity of prediction used by physicists 
is fitting the distribution of the events that were classified as signal (or background) over the mass (or some other variable).

This approach is hardly formalizable, and not automatable --- each time you should assume some kind of distribution. 
Ideally we want to have some easy-to-use out-of-the-box metrics like FOMs in machine learning (like area under the ROC, f1 or ).

\subsection{Ideal uniformity}

We start from the simplest case --- when we are fully satisfied by predictions of our classifier. Remember that output of 
classification is probabilities of each event being a signal and background event, only after we select some cut on probability we get classification.

Ideal uniformity of signal prediction means that whichever cut we select, the efficiency (part of signal event that passed the cut) doesn't depend on uniform variables: 
in every region of uniform variables space the part of signal events that passed the cut is the same.

In practice, of course, this never happens.

There is uniformity of predictions on signal and uniformity of efficiency on background, which can be defined one from another by swapping classes. 
In what follows in this section we are writing about the efficiency on signal events.

A good example of classifier that has close to ideal uniformity is classifier which returns a random probability in [0, 1] range. What a pity: in practice it's absolutely useless.

% The most significant drawback of the upcoming metrics is they are ill-defined if there are not too many events of the named class.

\subsection{Some Restrictions on Uniformity Metrics}

There are some additional conditions which we expect metric to meet:
\begin{enumerate}
\item
shouldn't depend much on the number of events (i.e., if we randomly select half of the events, the metrics should roughly be the same)
\item
shouldn't depend on the weights renormalization: if we multiply all the weight by some arbitrary number, it shouldn't change at all.
\item 
depends only on the order of predictions, not the exact values of probabilities.
This is because we care about which events pass the cut and which don't, not about the exact values of predictions.

Example: correlation of prediction and mass doesn't satisfy this restriction.
\item
parameter stability: if it uses bins, changing the number of bins shouldn't affect the metrics value much, if it uses $k$-nearest neighbors, it should be stable to small deviations of $k$.
\end{enumerate}

\subsection{Standard Deviation of Efficiency on Bins (SDE)}

\def\bineff{\text{eff}_\text{bin}}
\def\binweight{\text{weight}_\text{bin}}
\def\globaleff{\text{eff}}
\def\SDE{\text{SDE}}
\def\bin{\text{bin}}

Let's split the space of uniform features into bins, the uniformity in these terms is following: when we select some probability cut, 
the part of signal events that passes the cut is equal in all bins. Assume we selected some cut, then we have global efficiency
\[
	\globaleff = \dfrac{
		\text{total weight of signal events that passed the cut}}
		{\text{total weight of signal events}}
\]

Efficiency in every bin is defined respectively, 
\[
	\bineff = \dfrac{
		\text{weight of signal events in bin that passed the cut}}
		{\text{weight of signal events in this bin}} 
\]

So, basically, what we want to have in our dreams:
\[
	\bineff = \text{global efficiency} \qquad \forall \; \text{bin}
\]

To measure how far we are from ideal situation we use standard deviation:
\[
	\sqrt{\sum_{\bin} \left( \bineff - \globaleff \right)^2  }
\]

What is bad in this formula that every bin has some impact in the result, which does not depend on how many events are there, so metrics becomes very unstable to deviations in bins with only few events. To cure this, we add weights to the bins (note that $\sum_\bin \binweight = 1$):
\[
	\binweight = \dfrac{\text{total weight of signal events in bin}}
		{\text{total weight of signal events}},
\]
so we have SDE formula:
\[
	\SDE(\globaleff) = 
	\sqrt{\sum_{\bin} \binweight \times \left(\bineff - \globaleff \right)^2}. 
\] 
In fact, the expression depends on the cut, but for cuts which produce equal efficiency, this is

Finally we note that the weighted average of $\bineff$ is $\globaleff$:
\[
	\globaleff = < \bineff > =  \sum_{\bin} \binweight \times \bineff,
\]
and this is why the introduced metrics was named SDE --- this is a weighted standard deviations of array of bin efficiencies.

But this is how we measure the non-uniformity for only one fixed cut, to measure the overall non-flatness, we take several global efficiencies (for instance, [0.5, 0.6, 0.7, 0.8, 0.9], because in practice usually we are interested in cuts with high global efficiency) and use 
\[
	\SDE^2  =  \frac{1}{k} 
	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
		\text{SDE}^2(\globaleff)
\]

Some other power $p \neq 2$ can be used as well, but $p=2$ is considered as the default value: 
\[
	\SDE^p(\globaleff) = 
	\sum_{\bin} \binweight \times \abs{\bineff - \globaleff}^p,
\qquad
	\SDE^p  =  \frac{1}{k} 
	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
		\text{SDE}^p(\globaleff).
\]

\subsection{Theil Index of Efficiency}
\def\theil{\text{Theil}}

One more measure uses Theil Index frequently used to measure economic inequality:
\[
	\theil = \frac{1}{N} \sum_i \frac{x_i}{<x>} \ln{\frac{x_i}{<x>}}, 
		\qquad <x> = \frac{1}{N} \sum_i x_i
\]
In our case we have to alter formula a bit to take into account that different bins have different impact, thus the formula turns into
\[
	\theil(\globaleff) = \sum_\bin \binweight \; \frac{\bineff}{\globaleff} \; \ln{\frac{\bineff}{\globaleff}}
\]

TODO how to combine Theil for different global efficiencies?
\[
	\theil = ??? \text{from} \theil(\globaleff)
\]

\subsection{Distribution Similarity Approach}
\label{subsection:similarity}

Let's start from reformulation of what is uniform predictions in signal. First we split all signal events into some bins in uniform variables. There is some empirical distribution $F_\bin$ of predictions in each bin. Ideal uniformity means that all the distributions $F_\bin$ are equal and hence equal to the global distribution $F(x)$.

To 'measure' non-flatness we can use some distribution distance, like Kolmogorov-Smirnov:
\[
	 \sum_{\bin} \binweight \max_x \abs{F_{\bin}(x) - F(x)},
\]
but Cram\'er--von Mises similarity is more informative (usually $p=2$ is used):
\[
	 \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x),
\]

The good point is we don't need to select some global efficiencies like in the other metrics.

\subsection{Connection Between SDE and Distance Similarity Approach}

SDE and DSA based on Cram\'er--von Mises similarity can be shown to have connection.
Let's consider the SDE with global efficiencies $= [1/N, 2/N, \dots, N/N]$. In the limit $N \to \infty$
\[
	\lim_{N \to \infty} \SDE^2 = 
	\lim_{N \to \infty} \frac{1}{N} \sum_{\globaleff}\SDE^2(\globaleff) = 
	\int_0^1 \SDE^2(\globaleff) d\, \globaleff = 
	\int_0^1 \sum_{\bin} \binweight \abs{\bineff - \globaleff}^2 d\, \globaleff
\]

From the other side, we can write the expression for similarity-based measure (for $p=2$) 
\[
	\sum_{\bin} \binweight \int \abs{F_\text{bin}(x) - F(x)}^2 dF(x) =
	\int \sum_{\bin} \binweight \abs{F_\text{bin}(x) - F(x)}^2 dF(x) 
\] The hard thing now is to believe this is literally the same and these two expressions are equal.

\subsection{Knn-based modifications}

\def\knni{\text{knn}(i)}
\def\effknni{\text{eff}_{\knni}}
\def\weightknni{\text{weight}_{\knni}}
\def\Fknn{F_{\knni}}

\def\knnSDE{\text{knnSDE}}

Though operating with bins is usually both simple and very efficient, 
in many cases it is hard to find optimal size of bins in the space of uniform variables (specifically in the case of more than two dimensions).
One more situation when bins-based approach fails, is when we have too few events to obtain a good statistics at least in several bins.

In these cases we can switch to $k$-nearest neighbors: for each signal event we find $k$ nearest signal events (including the event itself) in the space of uniform variables. Now we can compute the efficiency $\effknni$, empirical distribution $\Fknn$ of nearest neighbors. 
The weights for $\knni$ are proportional to the total weight of events in $\knni$:
\[
	\weightknni = \alpha \sum_{j \in \knni} w_j, \qquad \alpha^{-1} = \sum_i \sum_{j \in \knni} w_j,
\]
so again weights are normed to 1: $\sum_{i} \knni = 1$.

Now we are ready to write knn version of SDE:
\[
	\knnSDE^2(\globaleff)
		= \sum_{i \in \text{events}} \weightknni \abs{\effknni - \globaleff}^2
\]
\[
	\knnSDE^2 = \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]}
		\knnSDE^2(\globaleff),
\]
knn version of Theil index of Efficiency
\[
	\text{knnTheil}(\globaleff) = \sum_{i \in \text{events}} \weightknni \; \frac{\effknni}{\globaleff} \; \ln{\frac{\effknni}{\globaleff}}
\]
\[
	\text{knnTheil} = ??? \text{knnTheil}(\globaleff)
\]
and knn version of similarity-based measure:

\[
	 \sum_{i \in \text{events}} \weightknni \int \abs{\Fknn(x) - F(x)}^p dF(x),
\]

$K$-nearest neighbors approach suffers from the other drawback: the impact of different events has very little connection with the weights, because some events are met in knn of other events much more frequently while the other.
This effect can be suppressed by dividing initial weight of the event by the number of times it is met in knn.

\subsection{Advantages and Disadvantages of Different Metrics}

... TODO, here some plots with comparison, maybe timings ...

\section{Approaches Proposed}

\subsection{Mean Ada Boost}

This is a modification of AdaBoost algorithm. In AdaBoost one multiplies weights in such a way:
\[
	w_i = w_i \times \exp[-y_i \, \text{score}_i],
\]
to enlarge the weights of poorly classified events ($y_i$ is $+1$ for signal and $-1$ for background).

But now we use the mean of prediction of $k$ nearest neighbors (of the same class)
\[
	w_i = w_i \times \exp[-y_i \, \dfrac{1}{k} \sum_{j \in \knni} \text{score}_j]
\]
Thus boosting focuses not on the events that were poorly classified, but on the regions with poor classification.

\subsection{Gradient Boosting with AdaLoss Modification (knn-Ada)}

\def\score{\text{score}}
\def\knn{\text{knn}}
\def\FL{\text{FL}}

Gradient boosting on trees is widely used algorithm[link], it's built upon decision tree regressors with usage of some loss function.

Let's start with examples. One of the popular losses used is AdaLoss:
\[
	\sum_{i \in \text{events}} w_i \times \exp [- \score_i \, y_i] 
\]
where $y_i$ is either +1 (for signal events) or -1 (for background events). Good classification supposes that signal events should have large positive scores, while background ones should have large negative.

The predictions of separate regressors are simply summed up to form a score:
\[
	\score_i = \sum_{r \in \text{regressors}} \text{pred}_r(i),
\]
which can be 'translated' into probabilities by logistic function.

So the goal of algorithm is now to minimize the loss function. At each stage it trains one more regressor, which should decrease the value of loss, the most vivid way is to train it on negative gradient of loss. In the case of AdaLoss this can be done pretty easy:
\[
	-\dfrac{\partial \, \text{AdaLoss}}{\partial \, \score_i} = w_i \, y_i \exp[- \score_i \, y_i],
\]
so it is positive for signal events, negative for background events and has larger modulus for the events which are poorly classified.

After a new tree was built, it's output is altered: for each leaf we can compute such a value, that the result will give the smallest possible value for loss.

...computation of optimal value goes here...

This loss function can be easily modified to take in account not only the score for individual elements, but also 'finds' regions with lower-than-average quality.

\[
	\text{knnAdaLoss} = \sum_{i \in events} \exp[-y_i \times \sum_{j \in \text{knn}(i)} \score_j],
\]
where the index $j$ goes over the $k$ nearest neighbors of $i$th event, which belongs to the same class (signal or background).

We can introduce a supplementary sparse matrix $A \in \RR^{N \times N}$ ($N$ is number of events), which is defined as 
\[
a_{ij} = \begin{cases} 
1, & j \in \knn(i), \text{ events $i$ and $j$ belong to the same class} \\
0, & \text{otherwise},
\end{cases}
\] so the loss can be written as
\[
	\text{knnAdaLoss} = \sum_i \exp [- y_i \sum_j a_{ij} \, \score_j ],
\]
it's negative gradient is easily computed:
\[
	-\dfrac{\partial \, \text{knnAdaLoss}} {\partial \, \score_k} = 
	 y_k \, \sum_i a_{ik} \exp [- y_i \sum_j a_{ij} \, \score_j ],
\]
from this we can see that new algorithm will pay more attention to the events, which has poorly classified neighbors (and neighbors of neighbors). The named loss targets to obtain uniformity in both signal and background.

And the post-correction can be handled by using second-order Newton-Raphson optimization step

... TODO computation of optimal value goes here ...

One can note, that the matrix $A$ doesn't need to be necessarily square. One can introduce $M$ groups of events (which may intersect), each group consists of several events with close uniform variables (and close events). Then one introduces $A \in \RR^{M \times N}$:
\[
	a_{mj} = \begin{cases}
		1, & \text{event $j$ is in group $m$} \\
		0, & \text{otherwise}
	\end{cases}
\]

In particular, if we take $A$ to be identity matrix: $A = I$ (each event to be in it's own group), knnAdaLoss turns into a simple AdaLoss.

\subsection{Gradient Boosting with Flatness Loss (FL)}

Let's use the metrics introduces in the section \ref{subsection:similarity}:
\[
	\sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x),
\]
which is good as a measure, but due to the non-smoothness of $F(x)$ it's 
gradient is singular, so we use instead
\[
	\FL = \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dx.
\]

So, the derivative looks like:
\[
	\dfrac{\partial} {\partial \, \score_i} \FL
	= \sum_{\bin} \binweight \frac{\partial }{ \partial \, \score_i} 
			\int \abs{F_\bin(x) - F(x)}^p dx
\]
Let $\bin(i)$ be a bin to which event $i$ belongs, then we can compute:
\def\binIweight{\text{weight}_\text{\bin(i)}}

% TODO sign

\begin{multline*}
	- \dfrac{\partial} {\partial \, \score_i} \FL = 
		- \binIweight
		\frac{\partial }{ \partial \, \score_i} 
			\int \abs{F_{\bin(i)}(x) - F(x)}^p dx \cong \\
	\cong \binIweight \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1} 
		\sgn[F_{\bin(i)}(x) - F(x)]
		\dfrac{w_i}{\binIweight}
		\Bigg|_{x=\score_i} = \\
	= 
		w_i \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1}
		\sgn[F_{\bin(i)}(x) - F(x)]
		\Bigg|_{x=\score_i}
\end{multline*}

TODO give explanations here

The next thing we need to point is FL doesn't take into account the quality of predictions. So what we use in practice is linear combination of FlatnessLoss and AdaLoss:
\[
	\text{loss} = \FL + \alpha \, \text{AdaLoss}
\]

First one penalizes the non-uniformity, second one --- poor predictions.

\section{Dataset}

Here should be some description of $D \to hhh$ sample.

\section{Plots}

\section{Timings}

The main drawback of uBoost technique is it's high computational complexity: 
while simple AdaBoost trains $M$ trees, uBoost builds $100 \times M$ trees (contribution of other operations usually can be neglected).

Presented in this paper classifiers are building $M$ trees, though there is more complicated boosting, it takes at each iteration ($k$ is number of neighbors, $N$ is number of events in training sample)

\begin{itemize}
	\item meanAdaBoost: $O(k \times N)$
	\item knnAdaLoss: $O(k \times N)$, for the arbitrary matrix $A$ it is 
	$O( \text{\#nonzero elements in the matrix})$
	\item FlatnessLoss: $O(N \ln N)$
\end{itemize}

\section{Summary}

\section{Acknowledgments}

\section{Source code}

The code for classifiers ans links to final notebooks with results in the article.

\begin{thebibliography}{99}

\bibitem{Cd94} Author, \emph{Title}, Journal/Editor, (year)

\end{thebibliography}

\end{document}

### 2014-08-19 IPythonWorkflow/gradient_boosting_old.py

@@ -1,0 +1,337 @@
+__author__ = 'axelr'
+from itertools import islice
+
+import numpy
+from scipy.special import expit
+from sklearn.base import BaseEstimator, ClassifierMixin
+from sklearn.cross_validation import train_test_split
+from sklearn.ensemble.forest import RandomForestClassifier, RandomForestRegressor
+from sklearn.random_projection import GaussianRandomProjection
+from sklearn.tree._tree import TREE_LEAF
+from sklearn.tree.tree import DecisionTreeRegressor, DTYPE
+from sklearn.utils.random import check_random_state
+from sklearn.utils.validation import check_arrays
+from scipy import special
+
+from commonutils import check_sample_weight
+import commonutils
+
+
+class BinomialDeviance(MyLossFunction):
+    """Binomial deviance loss function for binary classification.
+
+    Binary classification is a special case; here, we only need to
+    fit one tree instead of ``n_classes`` trees.
+    """
+    def __init__(self):
+        # we only need to fit one tree for binary clf.
+        super(BinomialDeviance, self).__init__()
+
+    def __call__(self, y, pred, sample_weight=None):
+        """Compute the deviance (= 2 * negative log-likelihood). """
+        # numpy.logaddexp(0, v) == log(1.0 + exp(v))
+        pred = pred.ravel()
+        return -2.0 * numpy.average((y * pred) - numpy.logaddexp(0.0, pred), weights=sample_weight)
+
+    def fit(self, X, y, sample_weight=None):
+        pass
+
+    def negative_gradient(self, y, y_pred, sample_weight=None):
+        """Compute the residual (= negative gradient). """
+        sample_weight = check_sample_weight(y, sample_weight=sample_weight)
+        return (y - expit(y_pred.ravel())) * sample_weight
+
+    def _update_terminal_region(self, tree, terminal_regions, leaf, X, y,
+                                residual, pred, sample_weight):
+        """Make a single Newton-Raphson step. our node estimate is given by:
+            sum(y - prob) / sum(prob * (1 - prob))
+        """
+        terminal_region = numpy.where(terminal_regions == leaf)[0]
+        residual = residual.take(terminal_region, axis=0)
+        # y = y.take(terminal_region, axis=0)
+        prob = expit(pred.take(terminal_region, axis=0))
[…]

### 2014-08-21 IPythonWorkflow/metrics.py
makefile for the article

@@ -1,0 +1,1 @@
+__author__ = 'axelr'

### 2014-08-21 assets/paperdraft/paper_draft.tex
Corrected according to JINST styles

@@
-\documentclass[12pt]{article}
+\pdfoutput=1 % only if pdf/png/jpg images are used
+\documentclass{JINST}
+
+\title{New Approaches to Uniform Boosting}
+
+\author{
+Aleksandar Bukva, 
+Vladimir Gligorov$^d$,
+Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~ ,
+Andrey Ustuzhanin$^b$ and
+Mike Williams$^c$\\
+\llap{$^a$}Lomonosov Moscow State University, Moscow, Russia\\
+\llap{$^b$}Yandex LLC, Moscow, Russia\\
+\llap{$^c$}Massachusetts Institute of Technology, Cambridge, MA, United States \\
+\llap{$^d$}Organisation Européenne pour la Recherche Nucléaire (CERN), Geneva \\Switzerland 
+E-mail: \email{alex.rogozhnikov@yandex.ru}}
+
+\abstract{
+% Abstract must be short enough to appear in this page together with title, authors, their addresses and keywords.
+The use of multivariate classifiers has become commonplace in particle physics. 
+Typically, a series of classifiers is trained rather than just one to enhance the performance; this is known as boosting.
+
+In some applications of high energy physics (i.e., amplitude analyses) 
+classifiers should not only optimize some integrated figure of merit, but produce a uniform selection efficiency in a space of selected physical variables.
+Recently uBoost technique of boosting was proposed which addresses this issue.
+
+In this paper we introduce several approaches to measure the uniformity of efficiency and propose new boosting techniques.
+}
 
@@
 \usepackage{amsmath, amsthm, amsfonts}
-\usepackage[margin=0.5in]{geometry}
 
@@
 %-----------------------------------------------------------------
-\title{New Approaches to Uniform Boosting }
-\author{
-}
 
@@
 \maketitle
-
-% \abstract{Abstract goes here}
 
@@
 \subsection{Distribution Similarity Approach}
-\label{subsection:similarity}
+\label{sec:similarity}
 
@@
 
-Let's use the metrics introduces in the section \ref{subsection:similarity}:
+Let's use the metrics introduces in the section \ref{sec:similarity}:
 \[
@@
 
-\section{Acknowledgments}
-
 \section{Source code}
 
-The code for classifiers ans links to final notebooks with results in the article.
+The link on the repository ans links to final notebooks with results in the article.
 
-\begin{thebibliography}{99}
+\acknowledgments
 
-\bibitem{Cd94} Author, \emph{Title}, Journal/Editor, (year)
+\begin{thebibliography}{9}
+
+\bibitem{bib1}
+Authors,
+\emph{Title},
+\emph{J. Ref.} \textbf{vol} (year) page.
+
+\bibitem{bib2}
+A. Pietropaolo et al., \emph{DINS measurements on VESUVIO in the
+    resonance detector configuration: proton mean kinetic energy in
+    water}, \jinst{1}{2006}{P04001}.
+
+\bibitem{bib3}
+A.I. Harris,
+\emph{Spectroscopy with multichannel correlation radiometers},
+\href{http://dx.doi.org/10.1063/1.1898643}
+{\emph{Rev.\ Sci.\ Instrum.} {\bf 76} (2005) 054503}
+[\astroph{0504449}].
+
+\bibitem{bib4}
+G.F. Knoll, \emph{Radiation detection and measurements}, John Wiley
+    and Sons, Inc., New York 2000.
+
+\bibitem{bib5}
+V. Dangendorf, \emph{Time-resolved fast-neutron imaging with a
+pulse-counting image intensifier}, in proceedings of
+\emph{International workshop on fast neutron detectors and
+applications}, April, 3--6, 2006 University of Cape Town, South Africa
+\pos{PoS(FNDA2006)008}.
 

### 2014-08-23 IPythonWorkflow/grid_search.py
added printing of crash report on the node

@@ -279,7 +279,7 @@ def estimate_classifier(params_dict, base_estimator, X, y, folds, fold_checks,
                         score_function, sample_weight=None, label=1, scorer_needs_x=False):
     """This function is needed """
     try:
-        k_folder = StratifiedKFold(y=y, n_folds=folds)
+        k_folder = StratifiedKFold(y=y, n_folds=folds, shuffle=True)
         score = 0.
         for train_indices, test_indices in islice(k_folder, fold_checks):
             trainX, trainY = X.irow(train_indices), y[train_indices]
@@ -301,8 +301,8 @@ def estimate_classifier(params_dict, base_estimator, X, y, folds, fold_checks,
             score += score_function(testY, proba[:, label], **test_options)
 
         return score / fold_checks
-    except:
-        return -0.999
+    except Exception as e:
+        return e
 
 
 class GridOptimalSearchCV(BaseEstimator, ClassifierMixin):
@@ -430,8 +430,16 @@ class GridOptimalSearchCV(BaseEstimator, ClassifierMixin):
                     [self.scorer_needs_x] * portion)
                 assert len(result) == portion, "The length of result is very strange"
                 for state_indices, state_dict, score in zip(state_indices_array, state_dict_array, result):
+                    params = ", ".join([k + '=' + str(v) for k, v in state_dict.iteritems()])
+                    if isinstance(Exception, score):
+                        # returned exception
+                        message = 'Fail during training on the node \nException ' + str(score) +\
+                                  '\nParameters:' + params
+                        print(message)
+                        self._log(message)
+                        continue
                     self.generator.add_result(state_indices, score)
-                    self._log(score, ": ", ", ".join([k + '=' + str(v) for k, v in state_dict.iteritems()]))
+                    self._log(score, ": ", params)
 
                 self.evaluations_done += portion
                 print("%i evaluations done" % self.evaluations_done)

### 2014-08-24 IPythonWorkflow/grid_search.py
doen't catches the exceptions in single-thread

@@ -276,7 +276,7 @@ test_simple_optimizer()
 
 
 def estimate_classifier(params_dict, base_estimator, X, y, folds, fold_checks,
-                        score_function, sample_weight=None, label=1, scorer_needs_x=False):
+                        score_function, sample_weight=None, label=1, scorer_needs_x=False, catch_exceptions=True):
     """This function is needed """
     try:
         k_folder = StratifiedKFold(y=y, n_folds=folds, shuffle=True)
@@ -302,7 +302,10 @@ def estimate_classifier(params_dict, base_estimator, X, y, folds, fold_checks,
 
         return score / fold_checks
     except Exception as e:
-        return e
+        if catch_exceptions:
+            return e
+        else:
+            raise
 
 
 class GridOptimalSearchCV(BaseEstimator, ClassifierMixin):
@@ -409,7 +412,9 @@ class GridOptimalSearchCV(BaseEstimator, ClassifierMixin):
                                             X=X, y=y, folds=self.folds, fold_checks=self.fold_checks,
                                             score_function=self.score_function, sample_weight=sample_weight,
                                             label=self.label,
-                                            scorer_needs_x=self.scorer_needs_x)
+                                            scorer_needs_x=self.scorer_needs_x,
+                                            catch_exceptions=False)
+
                 self.generator.add_result(state_indices, value)
                 self.evaluations_done += 1
                 state_string = ", ".join([k + '=' + str(v) for k, v in state_dict.iteritems()])

### 2014-08-25 assets/paperdraft/paper_draft.tex
small fixes
splitted into several parts

--- 2014-08-19 assets/paperdraft/paper_draft.tex
@@
-\documentclass[12pt]{article}
+\pdfoutput=1 % only if pdf/png/jpg images are used
+\documentclass{JINST}
+
+\title{New Approaches to Uniform Boosting}
+
+\author{
+Aleksandar Bukva, 
+Vladimir Gligorov$^d$,
+Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~ ,
+Andrey Ustuzhanin$^b$ and
+Mike Williams$^c$\\
+\llap{$^a$}Lomonosov Moscow State University, Moscow, Russia\\
+\llap{$^b$}Yandex LLC, Moscow, Russia\\
+\llap{$^c$}Massachusetts Institute of Technology, Cambridge, MA, United States \\
+\llap{$^d$}Organisation Europ\'eenne pour la Recherche Nucl\'eaire (CERN), Geneva, Switzerland  \\
+E-mail: \email{alex.rogozhnikov@yandex.ru}}
+
+\abstract{
+\input{parts/abstract}
+}
 
@@
 
-% \abstract{Abstract goes here}
-
 \section{Introduction}
 
-Methods of machine learning are playing a important role in particles physics nowadays, in particular, multivariate analysis (MVA). 
-Different classifiers like boosted decision trees (BDTs) and artificial neural networks (ANNs) are often used as an important step in analysis selection criteria.
-
-BDTs are now even used in software triggers [link]. The main point of boosting technique is training many simple classifiers and building a composition of their outputs.
-Classifiers are trained one-after-another, their inputs are augmented in such a way that new classifier should target more at those events which were poorly classified by previous ones.
-The resulting classifier obtained by combining them often much more powerful, than individual ones.
-
-In practice, however, there are some restrictions that trained classifier should meet (apart from having good classification quality, which is usually measured by some integrated FOM).
-For example, in an amplitude analysis obtaining a uniform efficiency in a multivariate space of physics variates, i.e., variates that are of physical interest, is often times more important than any integrated FOM based on the total amount of signal and background.
-Such analyses often have many variates in which a uniform efficiency is desired.
-A uniform efficiency reduces systematic uncertainties and helps maintain sensitivity to all hypotheses being tested.
-
-The one approach that was already proposed is uBoost (uniform BOOSTing) --- a modification of AdaBoost algorithm [link]
-
-Why do we need such algorithms:
-
-\begin{enumerate}
-	\item The result of classification is stable to check different hypotheses
-	\item To get unbiased distribution of signal, this enables us to compute real mass of particle
-	\item Not to get fake peak, thus not get a false discovery
-\end{enumerate}
-
-The features, along which we want the prediction to be flat, will be referred to as 
-\textit{uniform variables}. In high energy physics these are usually masses.
+\input{parts/introduction}
 
@@
 
-In this section we come up with some approaches on how to measure uniformity of prediction. The typical way of 'checking' uniformity of prediction used by physicists 
-is fitting the distribution of the events that were classified as signal (or background) over the mass (or some other variable).
-
-This approach is hardly formalizable, and not automatable --- each time you should assume some kind of distribution. 
-Ideally we want to have some easy-to-use out-of-the-box metrics like FOMs in machine learning (like area under the ROC, f1 or ).
-
-\subsection{Ideal uniformity}
-
-We start from the simplest case --- when we are fully satisfied by predictions of our classifier. Remember that output of 
-classification is probabilities of each event being a signal and background event, only after we select some cut on probability we get classification.
-
-Ideal uniformity of signal prediction means that whichever cut we select, the efficiency (part of signal event that passed the cut) doesn't depend on uniform variables: 
-in every region of uniform variables space the part of signal events that passed the cut is the same.
-
-In practice, of course, this never happens.
-
-There is uniformity of predictions on signal and uniformity of efficiency on background, which can be defined one from another by swapping classes. 
-In what follows in this section we are writing about the efficiency on signal events.
-
-A good example of classifier that has close to ideal uniformity is classifier which returns a random probability in [0, 1] range. What a pity: in practice it's absolutely useless.
-
-% The most significant drawback of the upcoming metrics is they are ill-defined if there are not too many events of the named class.
-
-\subsection{Some Restrictions on Uniformity Metrics}
-
-There are some additional conditions which we expect metric to meet:
-\begin{enumerate}
-\item
-shouldn't depend much on the number of events (i.e., if we randomly select half of the events, the metrics should roughly be the same)
-\item
-shouldn't depend on the weights renormalization: if we multiply all the weight by some arbitrary number, it shouldn't change at all.
-\item 
-depends only on the order of predictions, not the exact values of probabilities.
-This is because we care about which events pass the cut and which don't, not about the exact values of predictions.
-
-Example: correlation of prediction and mass doesn't satisfy this restriction.
-\item
-parameter stability: if it uses bins, changing the number of bins shouldn't affect the metrics value much, if it uses $k$-nearest neighbors, it should be stable to small deviations of $k$.
-\end{enumerate}
-
-\subsection{Standard Deviation of Efficiency on Bins (SDE)}
-
-\def\bineff{\text{eff}_\text{bin}}
-\def\binweight{\text{weight}_\text{bin}}
-\def\globaleff{\text{eff}}
-\def\SDE{\text{SDE}}
-\def\bin{\text{bin}}
-
-Let's split the space of uniform features into bins, the uniformity in these terms is following: when we select some probability cut, 
-the part of signal events that passes the cut is equal in all bins. Assume we selected some cut, then we have global efficiency
-\[
-	\globaleff = \dfrac{
-		\text{total weight of signal events that passed the cut}}
-		{\text{total weight of signal events}}
-\]
-
-Efficiency in every bin is defined respectively, 
-\[
-	\bineff = \dfrac{
-		\text{weight of signal events in bin that passed the cut}}
-		{\text{weight of signal events in this bin}} 
-\]
-
-So, basically, what we want to have in our dreams:
-\[
-	\bineff = \text{global efficiency} \qquad \forall \; \text{bin}
-\]
-
-To measure how far we are from ideal situation we use standard deviation:
-\[
-	\sqrt{\sum_{\bin} \left( \bineff - \globaleff \right)^2  }
-\]
-
-What is bad in this formula that every bin has some impact in the result, which does not depend on how many events are there, so metrics becomes very unstable to deviations in bins with only few events.
-To cure this, we add weights to the bins (note that $\sum_\bin \binweight = 1$):
-\[
-	\binweight = \dfrac{\text{total weight of signal events in bin}}
-		{\text{total weight of signal events}},
-\]
-so we have SDE formula:
-\[
-	\SDE(\globaleff) = 
-	\sqrt{\sum_{\bin} \binweight \times \left(\bineff - \globaleff \right)^2}. 
-\] 
-In fact, the expression depends on the cut, but for cuts which produce equal efficiency, this is
-
-Finally we note that the weighted average of $\bineff$ is $\globaleff$:
-\[
-	\globaleff = < \bineff > =  \sum_{\bin} \binweight \times \bineff,
-\]
-and this is why the introduced metrics was named SDE --- this is a weighted standard deviations of array of bin efficiencies.
-
-But this is how we measure the non-uniformity for only one fixed cut, to measure the overall non-flatness, we take several global efficiencies (for instance, [0.5, 0.6, 0.7, 0.8, 0.9], because in practice usually we are interested in cuts with high global efficiency) and use
-\[
-	\SDE^2  =  \frac{1}{k} 
-	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
-		\text{SDE}^2(\globaleff)
-\]
-
-Some other power $p \neq 2$ can be used as well, but $p=2$ is considered as the default value: 
-\[
-	\SDE^p(\globaleff) = 
-	\sum_{\bin} \binweight \times \abs{\bineff - \globaleff}^p,
-\qquad
-	\SDE^p  =  \frac{1}{k} 
-	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
-		\text{SDE}^p(\globaleff).
-\]
-
-\subsection{Theil Index of Efficiency}
-\def\theil{\text{Theil}}
-
-One more measure uses Theil Index frequently used to measure economic inequality:
-\[
-	\theil = \frac{1}{N} \sum_i \frac{x_i}{<x>} \ln{\frac{x_i}{<x>}}, 
-		\qquad <x> = \frac{1}{N} \sum_i x_i
-\]
-In our case we have to alter formula a bit to take into account that different bins have different impact, thus the formula turns into
-\[
-	\theil(\globaleff) = \sum_\bin \binweight \; \frac{\bineff}{\globaleff} \; \ln{\frac{\bineff}{\globaleff}}
-\]
-
-TODO how to combine Theil for different global efficiencies?
-\[
-	\theil = ??? \text{from} \theil(\globaleff)
-\]
-
-\subsection{Distribution Similarity Approach}
-\label{subsection:similarity}
-
-Let's start from reformulation of what is uniform predictions in signal.
-First we split all signal events into some bins in uniform variables.
-There is some empirical distribution $F_\bin$ of predictions in each bin.
-Ideal uniformity means that all the distributions $F_\bin$ are equal and hence equal to the global distribution $F(x)$.
-
-To 'measure' non-flatness we can use some distribution distance, like Kolmogorov-Smirnov:
-\[
-	 \sum_{\bin} \binweight \max_x \abs{F_{\bin}(x) - F(x)},
-\]
-but Cram\'er--von Mises similarity is more informative (usually $p=2$ is used):
-\[
-	 \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x),
-\]
-
-The good point is we don't need to select some global efficiencies like in the other metrics.
-
-\subsection{Connection Between SDE and Distance Similarity Approach}
-
-SDE and DSA based on Cram\'er--von Mises similarity can be shown to have connection.
-Let's consider the SDE with global efficiencies $= [1/N, 2/N, \dots, N/N]$. In the limit $N \to \infty$
-\[
-	\lim_{N \to \infty} \SDE^2 = 
-	\lim_{N \to \infty} \frac{1}{N} \sum_{\globaleff}\SDE^2(\globaleff) = 
-	\int_0^1 \SDE^2(\globaleff) d\, \globaleff = 
-	\int_0^1 \sum_{\bin} \binweight \abs{\bineff - \globaleff}^2 d\, \globaleff
-\]
-
-From the other side, we can write the expression for similarity-based measure (for $p=2$) 
-\[
-	\sum_{\bin} \binweight \int \abs{F_\text{bin}(x) - F(x)}^2 dF(x) =
-	\int \sum_{\bin} \binweight \abs{F_\text{bin}(x) - F(x)}^2 dF(x) 
-\] The hard thing now is to believe this is literally the same and these two expressions are equal.
-
-\subsection{Knn-based modifications}
-
-\def\knni{\text{knn}(i)}
-\def\effknni{\text{eff}_{\knni}}
-\def\weightknni{\text{weight}_{\knni}}
-\def\Fknn{F_{\knni}}
-
-\def\knnSDE{\text{knnSDE}}
-
-Though operating with bins is usually both simple and very efficient, 
-in many cases it is hard to find optimal size of bins in the space of uniform variables (specifically in the case of more than two dimensions).
-One more situation when bins-based approach fails, is when we have too few events to obtain a good statistics at least in several bins.
-
-In these cases we can switch to $k$-nearest neighbors: for each signal event we find $k$ nearest signal events (including the event itself) in the space of uniform variables.
-Now we can compute the efficiency $\effknni$, empirical distribution $\Fknn$ of nearest neighbors.
-The weights for $\knni$ are proportional to the total weight of events in $\knni$:
-\[
-	\weightknni = \alpha \sum_{j \in \knni} w_j, \qquad \alpha^{-1} = \sum_i \sum_{j \in \knni} w_j,
-\]
-so again weights are normed to 1: $\sum_{i} \knni = 1$.
-
-Now we are ready to write knn version of SDE:
-\[
-	\knnSDE^2(\globaleff)
-		= \sum_{i \in \text{events}} \weightknni \abs{\effknni - \globaleff}^2
-\]
-\[
-	\knnSDE^2 = \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]}
-		\knnSDE^2(\globaleff),
-\]
-knn version of Theil index of Efficiency
-\[
-	\text{knnTheil}(\globaleff) = \sum_{i \in \text{events}} \weightknni \; \frac{\effknni}{\globaleff} \; \ln{\frac{\effknni}{\globaleff}}
-\]
-\[
-	\text{knnTheil} = ??? \text{knnTheil}(\globaleff)
-\]
-and knn version of similarity-based measure:
-
-\[
-	 \sum_{i \in \text{events}} \weightknni \int \abs{\Fknn(x) - F(x)}^p dF(x),
-\]
-
-$K$-nearest neighbors approach suffers from the other drawback: the impact of different events has very little connection with the weights, because some events are met in knn of other events much more frequently while the other.
-This effect can be suppressed by dividing initial weight of the event by the number of times it is met in knn.
-
-\subsection{Advantages and Disadvantages of Different Metrics}
-
-... TODO, here some plots with comparison, maybe timings ...
+\input{parts/metrics}
 
@@
 
-\subsection{Mean Ada Boost}
-
-This is a modification of AdaBoost algorithm. In AdaBoost one multiplies weights in such a way:
-\[
-	w_i = w_i \times \exp[-y_i \, \text{score}_i],
-\]
-to enlarge the weights of poorly classified events ($y_i$ is $+1$ for signal and $-1$ for background).
-
-But now we use the mean of prediction of $k$ nearest neighbors (of the same class)
-\[
-	w_i = w_i \times \exp[-y_i \, \dfrac{1}{k} \sum_{j \in \knni} \text{score}_j]
-\]
-Thus boosting focuses not on the events that were poorly classified, but on the regions with poor classification.
-
-\subsection{Gradient Boosting with AdaLoss Modification (knn-Ada)}
-
-\def\score{\text{score}}
-\def\knn{\text{knn}}
-\def\FL{\text{FL}}
-
-Gradient boosting on trees is widely used algorithm[link], it's built upon decision tree regressors with usage of some loss function.
-
-Let's start with examples. One of the popular losses used is AdaLoss:
-\[
-	\sum_{i \in \text{events}} w_i \times \exp [- \score_i \, y_i] 
-\]
-where $y_i$ is either +1 (for signal events) or -1 (for background events).
-Good classification supposes that signal events should have large positive scores, while background ones should have large negative.
-
-The predictions of separate regressors are simply summed up to form a score:
-\[
-	\score_i = \sum_{r \in \text{regressors}} \text{pred}_r(i),
-\]
-which can be 'translated' into probabilities by logistic function.
-
-So the goal of algorithm is now to minimize the loss function.
-At each stage it trains one more regressor, which should decrease the value of loss, the most vivid way is to train it on negative gradient of loss.
-In the case of AdaLoss this can be done pretty easy:
-\[
-	-\dfrac{\partial \, \text{AdaLoss}}{\partial \, \score_i} = w_i \, y_i \exp[- \score_i \, y_i],
-\]
-so it is positive for signal events, negative for background events and has larger modulus for the events which are poorly classified.
-
-After a new tree was built, it's output is altered: for each leaf we can compute such a value, that the result will give the smallest possible value for loss.
-
-...computation of optimal value goes here...
-
-This loss function can be easily modified to take in account not only the score for individual elements, but also 'finds' regions with lower-than-average quality.
-
-\[
-	\text{knnAdaLoss} = \sum_{i \in events} \exp[-y_i \times \sum_{j \in \text{knn}(i)} \score_j],
-\]
-where the index $j$ goes over the $k$ nearest neighbors of $i$th event, which belongs to the same class (signal or background).
-
-We can introduce a supplementary sparse matrix $A \in \RR^{N \times N}$ ($N$ is number of events), which is defined as 
-\[
-a_{ij} = \begin{cases} 
-1, & j \in \knn(i), \text{ events $i$ and $j$ belong to the same class} \\
-0, & \text{otherwise},
-\end{cases}
-\] so the loss can be written as
-\[
-	\text{knnAdaLoss} = \sum_i \exp [- y_i \sum_j a_{ij} \, \score_j ],
-\]
-it's negative gradient is easily computed:
-\[
-	-\dfrac{\partial \, \text{knnAdaLoss}} {\partial \, \score_k} = 
-	 y_k \, \sum_i a_{ik} \exp [- y_i \sum_j a_{ij} \, \score_j ],
-\]
-from this we can see that new algorithm will pay more attention to the events, which has poorly classified neighbors (and neighbors of neighbors).
-The named loss targets to obtain uniformity in both signal and background.
-
-And the post-correction can be handled by using second-order Newton-Raphson optimization step
-
-... TODO computation of optimal value goes here ...
-
-One can note, that the matrix $A$ doesn't need to be necessarily square.
-One can introduce $M$ groups of events (which may intersect), each group consists of several events with close uniform variables (and close events).
-Then one introduces $A \in \RR^{M \times N}$:
-\[
-	a_{mj} = \begin{cases}
-		1, & \text{event $j$ is in group $m$} \\
-		0, & \text{otherwise}
-	\end{cases}
-\]
-
-In particular, if we take $A$ to be identity matrix: $A = I$ (each event to be in it's own group), knnAdaLoss turns into a simple AdaLoss.
-
-\subsection{Gradient Boosting with Flatness Loss (FL)}
-
-Let's use the metrics introduces in the section \ref{subsection:similarity}:
-\[
-	\sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x),
-\]
-which is good as a measure, but due to the non-smoothness of $F(x)$ it's 
-gradient is singular, so we use instead
-\[
-	\FL = \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dx.
-\]
-
-So, the derivative looks like:
-\[
-	\dfrac{\partial} {\partial \, \score_i} \FL
-	= \sum_{\bin} \binweight \frac{\partial }{ \partial \, \score_i} 
-			\int \abs{F_\bin(x) - F(x)}^p dx
-\]
-Let $\bin(i)$ be a bin to which event $i$ belongs, then we can compute:
-\def\binIweight{\text{weight}_\text{\bin(i)}}
-
-% TODO sign
-
-\begin{multline*}
-	- \dfrac{\partial} {\partial \, \score_i} \FL = 
-		- \binIweight
-		\frac{\partial }{ \partial \, \score_i} 
-			\int \abs{F_{\bin(i)}(x) - F(x)}^p dx \cong \\
-	\cong \binIweight \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1} 
-		\sgn[F_{\bin(i)}(x) - F(x)]
-		\dfrac{w_i}{\binIweight}
-		\Bigg|_{x=\score_i} = \\
-	= 
-		w_i \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1}
-		\sgn[F_{\bin(i)}(x) - F(x)]
-		\Bigg|_{x=\score_i}
-\end{multline*}
-
-TODO give explanations here
-
-The next thing we need to point is FL doesn't take into account the quality of predictions. So what we use in practice is linear combination of FlatnessLoss and AdaLoss:
-\[
-	\text{loss} = \FL + \alpha \, \text{AdaLoss}
-\]
-
-First one penalizes the non-uniformity, second one --- poor predictions.
+\input{parts/approaches}
 
@@
 
-Here should be some description of $D \to hhh$ sample.
+\input{parts/dataset}
 
@@
 
+\input{parts/plots}
+
 \section{Timings}
 
-The main drawback of uBoost technique is it's high computational complexity: 
-while simple AdaBoost trains $M$ trees, uBoost builds $100 \times M$ trees (contribution of other operations usually can be neglected).
-
-Presented in this paper classifiers are building $M$ trees, though there is more complicated boosting, it takes at each iteration ($k$ is number of neighbors, $N$ is number of events in training sample)
-
-\begin{itemize}
-	\item meanAdaBoost: $O(k \times N)$
-	\item knnAdaLoss: $O(k \times N)$, for the arbitrary matrix $A$ it is 
-	$O( \text{\#nonzero elements in the matrix})$
-	\item FlatnessLoss: $O(N \ln N)$
-\end{itemize}
+\input{parts/timings}
 
@@
 
-\section{Acknowledgments}
+\input{parts/summary}
 
@@
 
-The code for classifiers ans links to final notebooks with results in the article.
+The link on the repository ans links to final notebooks with results in the article.
 
-\begin{thebibliography}{99}
+\acknowledgments
 
-\bibitem{Cd94} Author, \emph{Title}, Journal/Editor, (year)
+\begin{thebibliography}{9}
+
+% The examples of items:
+% \bibitem{bib1}
+% Authors,
+% \emph{Title},
+% \emph{J. Ref.} \textbf{vol} (year) page.
+
+% \bibitem{bib2}
+% A. Pietropaolo et al., \emph{DINS measurements on VESUVIO in the
+%     resonance detector configuration: proton mean kinetic energy in
+%     water}, \jinst{1}{2006}{P04001}.
+
+% \bibitem{bib3}
+% A.I. Harris,
+% \emph{Spectroscopy with multichannel correlation radiometers},
+% \href{http://dx.doi.org/10.1063/1.1898643}
+% {\emph{Rev.\ Sci.\ Instrum.} {\bf 76} (2005) 054503}
+% [\astroph{0504449}].
+
+% \bibitem{bib4}
+% G.F. Knoll, \emph{Radiation detection and measurements}, John Wiley
+%     and Sons, Inc., New York 2000.
+
+% \bibitem{bib5}
+% V. Dangendorf, \emph{Time-resolved fast-neutron imaging with a
+% pulse-counting image intensifier}, in proceedings of
+% \emph{International workshop on fast neutron detectors and
+% applications}, April, 3--6, 2006 University of Cape Town, South Africa
+% \pos{PoS(FNDA2006)008}.
 

### 2014-08-25 assets/paperdraft/parts/abstract.tex
splitted into several parts

% Abstract must be short enough to appear in this page together with title, authors, their addresses and keywords.
The use of multivariate classifiers has become commonplace in particle physics. 
Typically, a series of classifiers is trained rather than just one to enhance the performance; this is known as boosting.

In some applications of high energy physics (i.e., amplitude analyses) 
classifiers should not only optimize some integrated figure of merit, but produce a uniform selection efficiency in a space of selected physical variables.
Recently uBoost technique of boosting was proposed which addresses this issue.

In this paper we introduce several approaches to measure the uniformity of efficiency and propose new boosting techniques.

### 2014-08-25 assets/paperdraft/parts/approaches.tex
splitted into several parts

\subsection{Mean Ada Boost}

This is a modification of AdaBoost algorithm. In AdaBoost one multiplies weights in such a way:
\[
	w_i = w_i \times \exp[-y_i \, \text{score}_i],
\]
to enlarge the weights of poorly classified events ($y_i$ is $+1$ for signal and $-1$ for background).

But now we use the mean of prediction of $k$ nearest neighbors (of the same class)
\[
	w_i = w_i \times \exp[-y_i \, \dfrac{1}{k} \sum_{j \in \knni} \text{score}_j]
\]
Thus boosting focuses not on the events that were poorly classified, but on the regions with poor classification.

\subsection{Gradient Boosting with AdaLoss Modification (knn-Ada)}

\def\score{\text{score}}
\def\knn{\text{knn}}
\def\FL{\text{FL}}

Gradient boosting on trees is widely used algorithm[link], it's built upon decision tree regressors with usage of some loss function.

Let's start with examples. One of the popular losses used is AdaLoss:
\[
	\sum_{i \in \text{events}} w_i \times \exp [- \score_i \, y_i] 
\]
where $y_i$ is either +1 (for signal events) or -1 (for background events). Good classification supposes that signal events should have large positive scores, while background ones should have large negative.

The predictions of separate regressors are simply summed up to form a score:
\[
	\score_i = \sum_{r \in \text{regressors}} \text{pred}_r(i),
\]
which can be 'translated' into probabilities by logistic function.

So the goal of algorithm is now to minimize the loss function. At each stage it trains one more regressor, which should decrease the value of loss, the most vivid way is to train it on negative gradient of loss. In the case of AdaLoss this can be done pretty easy:
\[
	-\dfrac{\partial \, \text{AdaLoss}}{\partial \, \score_i} = w_i \, y_i \exp[- \score_i \, y_i],
\]
so it is positive for signal events, negative for background events and has larger modulus for the events which are poorly classified.

After a new tree was built, it's output is altered: for each leaf we can compute such a value, that the result will give the smallest possible value for loss.

...computation of optimal value goes here...

This loss function can be easily modified to take in account not only the score for individual elements, but also 'finds' regions with lower-than-average quality.

\[
	\text{knnAdaLoss} = \sum_{i \in events} \exp[-y_i \times \sum_{j \in \text{knn}(i)} \score_j],
\]
where the index $j$ goes over the $k$ nearest neighbors of $i$th event, which belongs to the same class (signal or background).

We can introduce a supplementary sparse matrix $A \in \RR^{N \times N}$ ($N$ is number of events), which is defined as 
\[
a_{ij} = \begin{cases} 
1, & j \in \knn(i), \text{ events $i$ and $j$ belong to the same class} \\
0, & \text{otherwise},
\end{cases}
\] so the loss can be written as
\[
	\text{knnAdaLoss} = \sum_i \exp [- y_i \sum_j a_{ij} \, \score_j ],
\]
it's negative gradient is easily computed:
\[
	-\dfrac{\partial \, \text{knnAdaLoss}} {\partial \, \score_k} = 
	 y_k \, \sum_i a_{ik} \exp [- y_i \sum_j a_{ij} \, \score_j ],
\]
from this we can see that new algorithm will pay more attention to the events, which has poorly classified neighbors (and neighbors of neighbors). The named loss targets to obtain uniformity in both signal and background.

And the post-correction can be handled by using second-order Newton-Raphson optimization step

... TODO computation of optimal value goes here ...

One can note, that the matrix $A$ doesn't need to be necessarily square. One can introduce $M$ groups of events (which may intersect), each group consists of several events with close uniform variables (and close events). Then one introduces $A \in \RR^{M \times N}$:
\[
	a_{mj} = \begin{cases}
		1, & \text{event $j$ is in group $m$} \\
		0, & \text{otherwise}
	\end{cases}
\]

In particular, if we take $A$ to be identity matrix: $A = I$ (each event to be in it's own group), knnAdaLoss turns into a simple AdaLoss.

\subsection{Gradient Boosting with Flatness Loss (FL)}

Let's use the metrics introduces in the section \ref{sec:similarity}:
\[
	\sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x),
\]
which is good as a measure, but due to the non-smoothness of $F(x)$ it's 
gradient is singular, so we use instead
\[
	\FL = \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dx.
\]

So, the derivative looks like:
\[
	\dfrac{\partial} {\partial \, \score_i} \FL
	= \sum_{\bin} \binweight \frac{\partial }{ \partial \, \score_i} 
			\int \abs{F_\bin(x) - F(x)}^p dx
\]
Let $\bin(i)$ be a bin to which event $i$ belongs, then we can compute:
\def\binIweight{\text{weight}_\text{\bin(i)}}

% TODO sign

\begin{multline*}
	- \dfrac{\partial} {\partial \, \score_i} \FL = 
		- \binIweight
		\frac{\partial }{ \partial \, \score_i} 
			\int \abs{F_{\bin(i)}(x) - F(x)}^p dx \cong \\
	\cong \binIweight \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1} 
		\sgn[F_{\bin(i)}(x) - F(x)]
		\dfrac{w_i}{\binIweight}
		\Bigg|_{x=\score_i} = \\
	= 
		w_i \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1}
		\sgn[F_{\bin(i)}(x) - F(x)]
		\Bigg|_{x=\score_i}
\end{multline*}

TODO give explanations here

The next thing we need to point is FL doesn't take into account the quality of predictions. So what we use in practice is linear combination of FlatnessLoss and AdaLoss:
\[
	\text{loss} = \FL + \alpha \, \text{AdaLoss}
\]

First one penalizes the non-uniformity, second one --- poor predictions.

### 2014-08-25 assets/paperdraft/parts/introduction.tex
splitted into several parts

Methods of machine learning are playing a important role in particles physics nowadays, in particular, multivariate analysis (MVA). 
Different classifiers like boosted decision trees (BDTs) and artificial neural networks (ANNs) are often used as an important step in analysis selection criteria.

BDTs are now even used in software triggers [link]. The main point of boosting technique is training many simple classifiers and building a composition of their outputs.
Classifiers are trained one-after-another, their inputs are augmented in such a way that new classifier should target more at those events which were poorly classified by previous ones. The resulting classifier obtained by combining them often much more powerful, than individual ones.

In practice, however, there are some restrictions that trained classifier should meet (apart from having good classification quality, which is usually measured by some integrated FOM). For example, in an amplitude analysis obtaining a uniform efficiency in a multivariate space of physics variates, i.e., variates that are of physical interest, is often times more important than any integrated FOM based on the total amount of signal and background. Such analyses often have many variates in which a uniform efficiency is desired. A uniform efficiency reduces systematic uncertainties and helps maintain sensitivity to all hypotheses being tested.

The one approach that was already proposed is uBoost (uniform BOOSTing) --- a modification of AdaBoost algorithm [link]

Why do we need such algorithms:

\begin{enumerate}
	\item The result of classification is stable to check different hypotheses
	\item To get unbiased distribution of signal, this enables us to compute real mass of particle
	\item Not to get fake peak, thus not get a false discovery
\end{enumerate}

The features, along which we want the prediction to be flat, will be referred to as 
\textit{uniform variables}. In high energy physics these are usually masses.

### 2014-08-26 IPythonWorkflow/grid_search.py
minor fix

@@ -436,7 +436,7 @@ class GridOptimalSearchCV(BaseEstimator, ClassifierMixin):
                 assert len(result) == portion, "The length of result is very strange"
                 for state_indices, state_dict, score in zip(state_indices_array, state_dict_array, result):
                     params = ", ".join([k + '=' + str(v) for k, v in state_dict.iteritems()])
-                    if isinstance(Exception, score):
+                    if isinstance(score, Exception):
                         # returned exception
                         message = 'Fail during training on the node \nException ' + str(score) +\
                                   '\nParameters:' + params

### 2014-08-26 IPythonWorkflow/grid_search.py
avoiding negative p while generation

@@ -214,6 +214,7 @@ class SimpleParameterOptimizer(AbstractParameterGenerator):
             new_state_indices = list(start_indices)
 
             p = 1. - len(self.queued_tasks_) / self.n_evaluations
+            p = numpy.clip(p, 0, 100)
             distance = self.random_state.binomial(self.dimensions_sum // 6, p) + 1
             for _ in range(distance):
                 axis = self.random_state.randint(len(self.dimensions))

### 2014-08-27 IPythonWorkflow/metrics.py
different metrics were implemented listed in the article, many thing moved from reports to metrics, some ready-to-use things are implemented

@@ -1,1 +1,691 @@
-__author__ = 'axelr'
+# About
+
+# this module contains different metrics of uniformity
+# and the metrics of quality as well (which support weights, actually)
+
+from __future__ import division
+from __future__ import print_function
+
+import warnings
+import numpy
+import pandas
+
+from sklearn.metrics import auc
+from sklearn.utils.validation import column_or_1d, check_arrays
+from commonutils import check_sample_weight, compute_cut_for_efficiency, computeSignalKnnIndices
+from numpy.random.mtrand import RandomState
+from scipy.stats import ks_2samp
+
+__author__ = 'Alex Rogozhnikov'
+
+__all__ = ['sde', 'cvm_flatness', 'cvm_2samp', 'theil']
+
+
+#region Utilities for tests
+
+def generate_test_dataset(n_samples, n_bins):
+    random = RandomState()
+    y = random.uniform(size=n_samples) > 0.5
+    pred = random.uniform(size=(n_samples, 2))
+    weights = random.exponential(size=(n_samples,))
+    bins = random.randint(0, n_bins, n_samples)
+    groups = bin_to_group_indices(bin_indices=bins, mask=(y == 1))
+    return y, pred, weights, bins, groups
+
+#endregion
+
+
+#region Quality metrics
+
+def check_metrics_arguments(y_true, y_pred, sample_weight, two_class=True, binary_pred=True):
+    sample_weight = check_sample_weight(y_true, sample_weight=sample_weight)
+    y_true = column_or_1d(y_true)
+    y_pred = column_or_1d(y_pred)
+    assert len(y_true) == len(y_pred), \
+        'The lengths of y_true and y_pred are different: %i and %i' % (len(y_true), len(y_pred))
+    if two_class:
+        assert numpy.in1d(y_true, [0, 1]).all(), 'The y_true array should contain only two labels: 0 and 1, ' \
+            'it contains:' + str(numpy.unique(y_true))
+    if binary_pred:
+        assert numpy.in1d(y_pred, [0, 1]).all(), 'The y_pred array should contain only two labels: 0 and 1, ' \
+            'it contains:' + str(numpy.unique(y_pred))
+    return y_true, y_pred, sample_weight
+
+
+def roc_curve(y_true, y_score, sample_weight=None):
+    """ The same as sklearn.metrics.roc_curve, but this one supports weights,
+    function will be removed when move to sklearn 0.15"""
+    y_true, y_score, sample_weight = \
[…]

### 2014-08-27 IPythonWorkflow/reports.py
some things moved to metrics, different uniformity curves implemented,  regions introduced + some refactoring

@@ -7,7 +7,6 @@
 from __future__ import print_function
 from __future__ import division
 from itertools import islice
-from numpy.random.mtrand import RandomState
 
 try:
     from collections import OrderedDict
@@ -23,8 +22,11 @@ from sklearn.utils.validation import check_arrays, column_or_1d
 from matplotlib import cm
 from scipy.stats import pearsonr
 
-from commonutils import compute_bdt_cut, Binner, roc_curve, roc_auc_score, check_sample_weight, build_normalizer, \
-    compute_cut_for_efficiency
+from commonutils import compute_bdt_cut, Binner, \
+    check_sample_weight, build_normalizer, computeSignalKnnIndices, map_on_cluster
+from metrics import roc_curve, roc_auc_score, compute_bin_indices, \
+    compute_msee_on_bins, compute_sde_on_bins, compute_sde_on_groups, compute_theil_on_bins, \
+    bin_based_cvm, compute_bin_efficiencies, efficiency_score, compute_bin_weights
 
 
 __author__ = 'Alex Rogozhnikov'
@@ -37,39 +39,18 @@ __author__ = 'Alex Rogozhnikov'
 # IsBackgroundAsSignal - background, but classified as signal
 # ... and so on. Cute, right?
 
-def efficiency_score(y_true, y_pred, sample_weight=None):
-    """Efficiency = right classified signal / everything that is really signal
-    Efficiency == recall, returns -0.1 when ill-defined"""
-    sample_weight = check_sample_weight(y_true, sample_weight=sample_weight)
-    assert len(y_true) == len(y_pred), "Different size of arrays"
-    isSignal = numpy.sum(y_true * sample_weight) - 1e-6
-    isSignalAsSignal = numpy.sum(y_true * y_pred * sample_weight) + 1e-7
-    return isSignalAsSignal / isSignal
-    # the same, but with notifications
-    # return recall_score(answer, prediction)
-
-
-def background_efficiency_score(y_true, y_pred, sample_weight=None):
-    """BackgroundEfficiency == right classified bg / everything that is really bg == fpr"""
-    return efficiency_score(1 - y_true, 1 - y_pred, sample_weight=sample_weight)
-
-
-def as_signal_score(y_true, y_pred, sample_weight=None):
-    """Part of is signal = classified as signal / total amount of events"""
-    sample_weight = check_sample_weight(y_true, sample_weight)
-    assert len(y_true) == len(y_pred), "Different size of arrays"
[…]

### 2014-08-27 IPythonWorkflow/binninguniforming.py
Binner moved to commonutils

@@ -1,5 +1,7 @@
 import numpy
-import commonutils
+from commonutils import check_sample_weight, weighted_percentile, \
+    build_normalizer, generate_sample
+from numpy.random.mtrand import RandomState
 from sklearn.base import BaseEstimator, ClassifierMixin, clone
 from sklearn.cross_validation import train_test_split
 
@@ -16,6 +18,76 @@ __author__ = 'Alex Rogozhnikov'
 # this module is concerned deprecated
 
 # TODO try doing the same without binning
+
+
+class Binner:
+    def __init__(self, values, n_bins, sample_weight=None):
+        """Binner is a class that helps to split the values into several bins.
+        Initially an array of values is given, which is then splitted into 'bins_number' equal parts,
+        and thus we are computing limits (boundaries of bins)."""
+        percentiles = [i / n_bins for i in range(1, n_bins)]
+        sample_weight = check_sample_weight(values, sample_weight=sample_weight)
+        self.limits = weighted_percentile(values, percentiles, sample_weight=sample_weight)
+
+    def get_bins(self, values):
+        return numpy.searchsorted(self.limits, values)
+
+    def set_limits(self, limits):
+        self.limits = limits
+
+    @property
+    def n_bins(self):
+        return len(self.limits) + 1
+
+    def split_into_bins(self, *arrays):
+        """Splits the data of parallel arrays into bins, the first array is binning variable"""
+        values = arrays[0]
+        numpy_arrays = []
+        for i, array in enumerate(arrays):
+            assert len(array) == len(values), "passed arrays have different lengths"
+            numpy_arrays.append(numpy.array(array))
+        bins = self.get_bins(values)
+        result = []
+        for bin in range(len(self.limits) + 1):
+            indices = bins == bin
+            result.append([array[indices] for array in numpy_arrays])
+        return result
+
+
+def test_binner():
+    """This function tests binner class"""
+    random = RandomState()
+    binner = Binner(random.permutation(30), 3)
+    assert numpy.all(binner.limits > [9, 19]), 'failed on the limits'
+    assert numpy.all(binner.limits < [10, 20]), 'failed on the limits'
+    bins = binner.get_bins([-1000, 1000, 0, 10, 20, 9.0, 10.1, 19.0, 20.1])
[…]

### 2014-08-27 IPythonWorkflow/binninguniforming.py
added imports from future

@@ -1,3 +1,6 @@
+from __future__ import division
+from __future__ import print_function
+
 import numpy
 from commonutils import check_sample_weight, weighted_percentile, \
     build_normalizer, generate_sample
@@ -203,7 +206,7 @@ def test_binning_uniforming():
 
     score = roc_auc_score(testY, proba[:, 1])
     if score < 0.8:
-        print 'Predictions are awful, roc = %.2f' % score
+        print('Predictions are awful, roc = %.2f' % score)
 
 
 test_binning_uniforming()

### 2014-08-27 IPythonWorkflow/commonutils.py
some refactoring, Binner moved outside

@@ -3,9 +3,14 @@
 # This file contains some helpful functions and classes
 # which are often used (by other modules)
 
-
 from __future__ import print_function
 from __future__ import division
+from sklearn.utils.validation import check_arrays
+
+try:
+    from collections import OrderedDict
+except ImportError:
+    from ordereddict import OrderedDict
 
 import math
 import io
@@ -16,7 +21,6 @@ from scipy.special import expit
 import sklearn.cross_validation
 from sklearn.metrics.pairwise import pairwise_distances
 from sklearn.neighbors.unsupervised import NearestNeighbors
-from sklearn.metrics import auc
 
 __author__ = "Alex Rogozhnikov"
 
@@ -34,17 +38,6 @@ def execute_notebook(filename):
             ip.run_cell(cell.input)
 
 
-def check_sample_weight(y_true, sample_weight):
-    """Checks the weights, returns normalized version """
-    if sample_weight is None:
-        return numpy.ones(len(y_true), dtype=numpy.float)
-    else:
-        sample_weight = numpy.array(sample_weight, dtype=numpy.float)
-        assert len(y_true) == len(sample_weight), \
-            "The lengths are different: {0} and {1}".format(len(y_true), len(sample_weight))
-        return sample_weight
-
-
 def map_on_cluster(ipc_profile, *args, **kw_args):
     """The same as map, but the first argument is ipc_profile
     :param str|None ipc_profile: the IPython cluster profile to use.
@@ -54,7 +47,63 @@ def map_on_cluster(ipc_profile, *args, **kw_args):
         return map(*args, **kw_args)
     else:
         from IPython.parallel import Client
+
         return Client(profile=ipc_profile).load_balanced_view().map_sync(*args, **kw_args)
+
+
+def sigmoid_function(x, width):
+    """ Sigmoid function is smoothing oh Heaviside function,
+    the less width, the closer we are to Heaviside function
+    :type x: array-like with floats, arbitrary shape
+    :type width: float, if width == 0, this is simply Heaviside function
+    """
+    assert width >= 0, 'the width should be non-negative'
+    if abs(width) > 0.0001:
+        return expit(numpy.clip(x / width, -500, 500))
+    else:
+        return (x > 0) * 1.0
+
+
+def generate_sample(n_samples, n_features, distance=2.0):
+    """Generates some test distribution,
[…]

### 2014-08-27 IPythonWorkflow/reports.py
removed correlation

@@ -22,11 +22,11 @@ from sklearn.utils.validation import check_arrays, column_or_1d
 from matplotlib import cm
 from scipy.stats import pearsonr
 
-from commonutils import compute_bdt_cut, Binner, \
+from commonutils import compute_bdt_cut, \
     check_sample_weight, build_normalizer, computeSignalKnnIndices, map_on_cluster
 from metrics import roc_curve, roc_auc_score, compute_bin_indices, \
     compute_msee_on_bins, compute_sde_on_bins, compute_sde_on_groups, compute_theil_on_bins, \
-    bin_based_cvm, compute_bin_efficiencies, efficiency_score, compute_bin_weights
+    bin_based_cvm, compute_bin_efficiencies, compute_bin_weights
 
 
 __author__ = 'Alex Rogozhnikov'
@@ -54,9 +54,8 @@ def train_classifier(name_classifier, X, y, sample_weight=None):
 
 
 class ClassifiersDict(OrderedDict):
-    """This class is a collection of classifiers, which will be trained simultaneously
-    and will be
-    """
+    """A collection of classifiers, which will be trained simultaneously
+    and after that will be compared"""
     def fit(self, X, y, sample_weight=None, ipc_profile=None):
         """Trains all classifiers on the same train data,
         if ipc_profile in not None, it is used as a name of IPython cluster to use for parallel computations"""
@@ -390,41 +389,6 @@ class Predictions(object):
                     pylab.show()
         return self
 
-    def correlation(self, var_name, stages=None, metrics=efficiency_score, n_bins=20, thresholds=None):
-        """ Plots the dependence of efficiency / sensitivity / whatever vs one of the variables
-        :type var_name: str, the name of variable
-        :type stages: list(int) | NoneType
-        :type metrics: function
-        :type n_bins: int, the number of bins
-        :type thresholds: list(float) | NoneType
-        :rtype: Predictions, returns self
-        """
-        thresholds = [0.2, 0.4, 0.5, 0.6, 0.8] if thresholds is None else thresholds
-        sample_weight = check_sample_weight(self.y, self.sample_weight)
-        for stage, predictions in pandas.DataFrame(self._get_stages(stages=stages)).iterrows():
-            self._strip_figure(len(predictions))
-            print('stage ' + str(stage))
-            for i, (name, proba) in enumerate(predictions.iteritems()):
-                pylab.subplot(1, len(predictions), i + 1)
-                correlation_values = numpy.ravel(self.X[var_name])
-                # Do we really need binner?
[…]

### 2014-08-27 IPythonWorkflow/metrics.py
readmes fixed in metrics

@@ -35,6 +35,23 @@ def generate_test_dataset(n_samples, n_bins):
 #endregion
 
 
+""" README on quality metrics
+
+Some notation used here
+IsSignal - is really signal
+AsSignal - classified as signal
+IsBackgroundAsSignal - background, but classified as signal
+... and so on. Cute, right?
+
+tpr = s = IsSAsS / isS
+fpr = b = IsBAsS / isB
+
+signal efficiency = tpr = s
+background efficiency = IsBasB / IsB = 1 - fpr
+background rejection = background efficiency
+
+"""
+
 #region Quality metrics
 
 def check_metrics_arguments(y_true, y_pred, sample_weight, two_class=True, binary_pred=True):
@@ -226,11 +243,13 @@ test_bins()
 README on flatness
 
 this metrics are unfortunately more complicated than usual ones
-and require the information about
+and require more information: not only predictions and classes,
+but also mass (or other variables along which we want to split data)
 
 Here we compute the different metrics of uniformity of predictions:
 
 SDE (formerly MSEE) - the standard deviation of efficiency
+Theil - Theil index of Efficiency (Theil index is used in economics)
 KS  - based on Kolmogorov-Smirnov distance between distributions
 CVM - based on Cramer-von Mises similarity between distributions
 
@@ -512,6 +531,7 @@ def compute_theil_on_bins(y_pred, mask, bin_indices, target_efficiencies, sample
 
 
 #endregion
+
 
 #region Similarity-based measures of flatness, KS, CvM
 

### 2014-08-27 IPythonWorkflow/reports.py
change of order + moved comments to metrics module

@@ -32,14 +32,6 @@ from metrics import roc_curve, roc_auc_score, compute_bin_indices, \
 __author__ = 'Alex Rogozhnikov'
 
 
-# Score functions
-# Some notation used here
-# IsSignal - is really signal
-# AsSignal - classified as signal
-# IsBackgroundAsSignal - background, but classified as signal
-# ... and so on. Cute, right?
-
-
 def train_classifier(name_classifier, X, y, sample_weight=None):
     """ Trains one classifier on a separate node in cluster,
     :param name_classifier: 2-tuple (name, classifier)
@@ -200,58 +192,59 @@ class Predictions(object):
 
     #endregion
 
-    #region MSE-related stuff (to be removed)
+    #region Quality-related methods
 
-    def _compute_bin_indices(self, var_names, n_bins=20, mask=None):
-        """Mask is used to show events that will be binned after"""
-        #TODO merge with next function
-        for var in var_names:
-            assert var in self.X.columns, "the variable %i is not in dataset" % var
-        mask = self._check_mask(mask)
-        bin_limits = []
-        for var_name in var_names:
-            var_data = self.X.loc[mask, var_name]
-            bin_limits.append(numpy.linspace(numpy.min(var_data), numpy.max(var_data), n_bins + 1)[1: -1])
-        return compute_bin_indices(self.X, var_names, bin_limits)
-
-    def _compute_bin_centers(self, var_names, n_bins=20, mask=None):
-        """Mask is used to show events that will be binned after"""
-        bin_centers = []
-        mask = self._check_mask(mask)
-        for var_name in var_names:
-            var_data = self.X.loc[mask, var_name]
-            bin_centers.append(numpy.linspace(numpy.min(var_data), numpy.max(var_data), 2 * n_bins + 1)[1::2])
-            assert len(bin_centers[-1]) == n_bins
-        return bin_centers
-
-    def _compute_staged_mse(self, var_names, target_efficiencies=None, step=3, n_bins=20, power=2., label=1):
-        target_efficiencies = self._check_efficiencies(target_efficiencies)
-        mask = self.y == label
-        bin_indices = self._compute_bin_indices(var_names, n_bins=n_bins, mask=mask)
-        compute_mse = lambda pred: \
-            compute_msee_on_bins(pred[:, label], mask, bin_indices, target_efficiencies=target_efficiencies,
-                                 power=power, sample_weight=self.sample_weight)
-        return self._map_on_staged_proba(compute_mse, step)
-
[…]

### 2014-08-28 IPythonWorkflow/metrics.py
added lazy function for theil + minor fixes

@@ -190,7 +190,7 @@ def compute_bin_indices(X, var_names, bin_limits=None, mask=None, n_bins=20):
     if bin_limits is None:
         bin_limits = []
         for var_name in var_names:
-            var_data = X.loc[mask, var_name]
+            var_data = X.loc[:, var_name] if mask is None else X.loc[mask, var_name]
             bin_limits.append(numpy.linspace(numpy.min(var_data), numpy.max(var_data), n_bins + 1)[1: -1])
 
     assert len(var_names) == len(bin_limits), "Different size of arrays"
@@ -340,6 +340,7 @@ def weighted_deviation(a, weights, power=2.):
 def compute_sde_on_bins(y_pred, mask, bin_indices, target_efficiencies, power=2., sample_weight=None):
     # skip check for a while
     # ignoring events from other classes
+    sample_weight = check_sample_weight(y_pred, sample_weight=sample_weight)
     y_pred = y_pred[mask]
     bin_indices = bin_indices[mask]
     sample_weight = sample_weight[mask]
@@ -412,7 +413,6 @@ def compute_msee_on_groups(y_pred, mask, groups_indices, target_efficiencies, sa
     cuts = compute_cut_for_efficiency(target_efficiencies, mask, y_pred=y_pred, sample_weight=sample_weight)
 
     efficiencies = [list() for eff in target_efficiencies]
-    groups_sizes = numpy.array([len(x) for x in groups_indices])
     groups_weights = numpy.array([numpy.sum(numpy.take(sample_weight, g)) for g in groups_indices])
     signal_weight = sample_weight[mask]
 
@@ -511,9 +511,10 @@ def theil(x, weights):
 
 def compute_theil_on_bins(y_pred, mask, bin_indices, target_efficiencies, sample_weight):
     warnings.warn('Theil on bins is in experimental version', UserWarning)
-    # ignoring events from other classes
+    y_pred = column_or_1d(y_pred)
     sample_weight = check_sample_weight(y_pred, sample_weight=sample_weight)
 
+    # ignoring events from other classes
     y_pred = y_pred[mask]
     bin_indices = bin_indices[mask]
     sample_weight = sample_weight[mask]
@@ -529,6 +530,34 @@ def compute_theil_on_bins(y_pred, mask, bin_indices, target_efficiencies, sample
     return result
 
 
[…]

### 2014-08-28 IPythonWorkflow/metrics.py
fixed mean in theil

@@ -527,7 +527,7 @@ def compute_theil_on_bins(y_pred, mask, bin_indices, target_efficiencies, sample
         bin_efficiencies = compute_bin_efficiencies(y_pred, bin_indices=bin_indices,
                                                     cut=cut, sample_weight=sample_weight)
         result += theil(bin_efficiencies, weights=bin_weights)
-    return result
+    return result / len(cuts)
 
 
 def compute_theil_on_groups(y_pred, mask, groups_indices, target_efficiencies, sample_weight):
@@ -540,7 +540,7 @@ def compute_theil_on_groups(y_pred, mask, groups_indices, target_efficiencies, s
     for cut in cuts:
         groups_efficiencies = compute_group_efficiencies(y_pred, groups_indices, cut, sample_weight=sample_weight)
         result += theil(groups_efficiencies, groups_weights)
-    return result
+    return result / len(cuts)
 
 
 def theil_flatness(y, proba, X, uniform_variables, sample_weight=None, label=1, knn=30):

### 2014-08-28 IPythonWorkflow/metrics.py
== replaced with np.allclose

@@ -43,11 +43,11 @@ AsSignal - classified as signal
 IsBackgroundAsSignal - background, but classified as signal
 ... and so on. Cute, right?
 
-tpr = s = IsSAsS / isS
-fpr = b = IsBAsS / isB
+tpr = s = isSasS / isS
+fpr = b = isBasS / isB
 
 signal efficiency = tpr = s
-background efficiency = IsBasB / IsB = 1 - fpr
+background efficiency = isBasB / isB = 1 - fpr
 background rejection = background efficiency
 
 """
@@ -105,14 +105,6 @@ def roc_auc_score(y_true, y_score, sample_weight=None):
     return auc(fpr, tpr, reorder=True)
 
 
-def sensitivity(y_true, y_score, sample_weight=None):
-    """ Returns s / sqrt{s+b} """
-    y_true, y_score, sample_weight = \
-        check_metrics_arguments(y_true, y_score, sample_weight=sample_weight, two_class=True, binary_pred=True)
-    s, b = compute_sb(y_true, y_score, sample_weight=sample_weight)
-    return s / numpy.sqrt(s + b)
-
-
 def efficiency_score(y_true, y_pred, sample_weight=None):
     """Efficiency = right classified signal / everything that is really signal
     Efficiency == recall, returns -0.1 when ill-defined"""
@@ -126,7 +118,7 @@ def efficiency_score(y_true, y_pred, sample_weight=None):
 
 
 def background_efficiency_score(y_true, y_pred, sample_weight=None):
-    """BackgroundEfficiency == right classified bg / everything that is really bg == fpr"""
+    """BackgroundEfficiency == isBasB / isB == 1 - fpr"""
     return efficiency_score(1 - y_true, 1 - y_pred, sample_weight=sample_weight)
 
 
@@ -135,6 +127,14 @@ def as_signal_score(y_true, y_pred, sample_weight=None):
     sample_weight = check_sample_weight(y_true, sample_weight)
     assert len(y_true) == len(y_pred), "Different size of arrays"
     return numpy.sum(y_pred * sample_weight) / numpy.sum(sample_weight)
+
+
+def sensitivity(y_true, y_score, sample_weight=None):
+    """ Returns s / sqrt{s+b} """
+    y_true, y_score, sample_weight = \
+        check_metrics_arguments(y_true, y_score, sample_weight=sample_weight, two_class=True, binary_pred=True)
[…]

### 2014-08-28 IPythonWorkflow/metrics.py
kolmogorov-smirnov added, speed-up of Cramer-von Mises

@@ -7,21 +7,26 @@ from __future__ import division
 from __future__ import print_function
 
 import warnings
+
 import numpy
 import pandas
-
 from sklearn.metrics import auc
 from sklearn.utils.validation import column_or_1d, check_arrays
-from commonutils import check_sample_weight, compute_cut_for_efficiency, computeSignalKnnIndices
 from numpy.random.mtrand import RandomState
 from scipy.stats import ks_2samp
+
+from commonutils import check_sample_weight, compute_cut_for_efficiency, computeSignalKnnIndices
+
 
 __author__ = 'Alex Rogozhnikov'
 
 __all__ = ['sde', 'cvm_flatness', 'cvm_2samp', 'theil']
 
 
-#region Utilities for tests
+#region Utilities
+
+def compute_F(weights):
+    return numpy.cumsum(weights) - 0.5 * weights
 
 def generate_test_dataset(n_samples, n_bins):
     random = RandomState()
@@ -565,24 +570,84 @@ test_theil()
 
 #region Similarity-based measures of flatness, KS, CvM
 
-def bin_based_KS(y_pred, sample_weight, bin_indices):
-    """Kolmogorov-Smirnov"""
-    # TODO use KS criterion with weights
-    assert len(y_pred) == len(sample_weight) == len(bin_indices)
-    global_distribution = numpy.sort(y_pred)
+def _prepare_data(data, weights):
+    """Prepares the distribution to be used later in KS and CvM"""
+    weights = weights / numpy.sum(weights)
+    prepared_data = numpy.unique(data)
+    indices = numpy.searchsorted(prepared_data, data)
+    prepared_weights = numpy.bincount(indices, weights=weights)
+    F = compute_F(prepared_weights)
+    return prepared_data, prepared_weights, F
+
+
+def _ks_2samp_fast(prepared_data1, data2, prepared_weights1, weights2, F1):
+    """Pay attention - prepared data should not only be sorted,
+    but equal items should be merged (by summing weights),
+    data2 should not have elements larger then max(prepared_data1) """
+    indices = numpy.searchsorted(prepared_data1, data2)
+    weights2 /= numpy.sum(weights2)
+    prepared_weights2 = numpy.bincount(indices, weights=weights2, minlength=len(prepared_data1))
+    F2 = compute_F(prepared_weights2)
+    return numpy.max(numpy.abs(F1 - F2))
+
+
+def test_ks2samp_fast(size=1000):
[…]

### 2014-08-29 IPythonWorkflow/uniformgradientboosting.py
starting to write new gradient boosting

@@ -1,15 +1,15 @@
 from __future__ import print_function
 from __future__ import division
 
-from collections import defaultdict, OrderedDict
-from itertools import izip
+from collections import defaultdict
+import copy
 import numbers
 from time import time
-import itertools
-import math
+
+import pandas
 import scipy.sparse as sparse
 import sklearn
-from sklearn.base import BaseEstimator
+from sklearn.base import BaseEstimator, ClassifierMixin
 from sklearn.ensemble import GradientBoostingClassifier as GBClassifier
 from sklearn.ensemble._gradient_boosting import _random_sample_mask
 from sklearn.ensemble.forest import RandomForestRegressor
@@ -19,11 +19,14 @@ import numpy
 from sklearn.metrics.pairwise import pairwise_distances
 from sklearn.neighbors.unsupervised import NearestNeighbors
 from sklearn.tree._tree import DTYPE
+from sklearn.tree.tree import DecisionTreeRegressor
 from sklearn.utils.random import check_random_state
 from sklearn.utils.validation import check_arrays, column_or_1d
+
 from commonutils import generate_sample, check_sample_weight
 import commonutils
 import reports
+import metrics
 
 __author__ = 'Alex Rogozhnikov'
 
@@ -516,7 +519,7 @@ def exp_margin(margin):
 
 class FlatnessLossFunction(LossFunction, BaseEstimator):
     def __init__(self, uniform_variables, bins=10, uniform_label=1, power=2., ada_coefficient=1.,
-                 allow_wrong_signs=True, keep_debug_info=False):
+                 allow_wrong_signs=True, median=False, keep_debug_info=False):
         """
         This loss function contains separately penalty for non-flatness and ada_coefficient.
         The penalty for non-flatness is using bins.
@@ -538,6 +541,7 @@ class FlatnessLossFunction(LossFunction, BaseEstimator):
         self.ada_coefficient = ada_coefficient
         self.allow_wrong_signs = allow_wrong_signs
         self.keep_debug_info = keep_debug_info
+        self.median = median
         LossFunction.__init__(self, 1)
 
     def fit(self, X, y, sample_weight=None):
@@ -585,7 +589,7 @@ class FlatnessLossFunction(LossFunction, BaseEstimator):
             for axis_limits in bin_limits:
                 bin_limits2.append(axis_limits[1 + shift:-1:2])
             bin_indices = reports.compute_bin_indices(X, self.uniform_variables, bin_limits2)
-            result += reports.bin_to_group_indices(bin_indices, mask=mask)
[…]

### 2014-08-29 IPythonWorkflow/metrics.py
removed MSE (now it is named SDE)

@@ -22,11 +22,17 @@ __author__ = 'Alex Rogozhnikov'
 
 __all__ = ['sde', 'cvm_flatness', 'cvm_2samp', 'theil']
 
+#TODO simpler interfaces
+#TODO uniformity in usage of masks
 
 #region Utilities
 
-def compute_F(weights):
-    return numpy.cumsum(weights) - 0.5 * weights
+
+def compute_cdf(ordered_weights):
+    """Computes cumulative distribution function (CDF) by ordered weights,
+    be sure that sum(ordered_weights) == 1
+    """
+    return numpy.cumsum(ordered_weights) - 0.5 * ordered_weights
 
 def generate_test_dataset(n_samples, n_bins):
     random = RandomState()
@@ -58,6 +64,7 @@ background rejection = background efficiency
 """
 
 #region Quality metrics
+
 
 def check_metrics_arguments(y_true, y_pred, sample_weight, two_class=True, binary_pred=True):
     sample_weight = check_sample_weight(y_true, sample_weight=sample_weight)
@@ -188,7 +195,7 @@ def compute_bin_indices(X, var_names, bin_limits=None, mask=None, n_bins=20):
     the indices are unique numbers of bin from zero to \prod_j (len(bin_limits[j])+1)
     Example:
         var_names = ["M2AB", "M2AC"]
-        bin_limits = [numpy.linspace(0, 1, 20), numpy.linspace(0, 1, 20)]
+        bin_limits = [numpy.linspace(0, 1, 21), numpy.linspace(0, 1, 21)]
 
     If bin_limits is not provided, they are computed using mask and n_bins
     """
@@ -232,7 +239,7 @@ def test_bin_to_group_indices(size=100, bins=10):
 test_bin_to_group_indices()
 
 
-def test_bins(size=500, n_bins = 10):
+def test_bins(size=500, n_bins=10):
     columns = ['var1', 'var2']
     df = pandas.DataFrame(numpy.random.random((size, 2)), columns=columns)
     x_limits = numpy.linspace(0, 1, n_bins + 1)[1:-1]
@@ -323,7 +330,8 @@ def compute_group_efficiencies(y_score, groups_indices, cut, sample_weight=None)
 
     if isinstance(groups_indices, numpy.ndarray) and numpy.ndim(groups_indices) == 2:
         # this speedup is specially for knn
[…]

### 2014-08-29 IPythonWorkflow/reports.py

@@ -25,8 +25,8 @@ from scipy.stats import pearsonr
 from commonutils import compute_bdt_cut, \
     check_sample_weight, build_normalizer, computeSignalKnnIndices, map_on_cluster
 from metrics import roc_curve, roc_auc_score, compute_bin_indices, \
-    compute_msee_on_bins, compute_sde_on_bins, compute_sde_on_groups, compute_theil_on_bins, \
-    bin_based_cvm, compute_bin_efficiencies, compute_bin_weights
+    compute_sde_on_bins, compute_sde_on_groups, compute_theil_on_bins, \
+    bin_based_cvm, compute_bin_efficiencies, compute_bin_weights, bin_based_ks
 
 
 __author__ = 'Alex Rogozhnikov'
@@ -217,7 +217,7 @@ class Predictions(object):
             for i, (clf_name, predict_proba) in enumerate(proba_on_stage.iteritems(), 1):
                 pylab.subplot(1, len(proba_on_stage), i)
                 for label in numpy.unique(self.y):
-                    pylab.hist(predict_proba[self.y == label, label], histtype=histtype, bins=bins, label=label)
+                    pylab.hist(predict_proba[self.y == label, label], histtype=histtype, bins=bins, label=str(label))
                 pylab.title('Predictions of %s at stage %s' % (clf_name, str(stage_name)))
                 if show_legend:
                     pylab.legend()
@@ -249,6 +249,29 @@ class Predictions(object):
     #endregion
 
     #region Uniformity-related methods
+
+    def _compute_bin_indices(self, var_names, n_bins=20, mask=None):
+        """Mask is used to show events that will be binned after"""
+        #TODO merge with next function
+        for var in var_names:
+            assert var in self.X.columns, "the variable %i is not in dataset" % var
+        mask = self._check_mask(mask)
+        bin_limits = []
+        for var_name in var_names:
+            var_data = self.X.loc[mask, var_name]
+            bin_limits.append(numpy.linspace(numpy.min(var_data), numpy.max(var_data), n_bins + 1)[1: -1])
+        return compute_bin_indices(self.X, var_names, bin_limits)
+
+    def _compute_bin_centers(self, var_names, n_bins=20, mask=None):
+        """Mask is used to show events that will be binned after"""
+        bin_centers = []
+        mask = self._check_mask(mask)
+        for var_name in var_names:
+            var_data = self.X.loc[mask, var_name]
[…]

### 2014-08-29 assets/paperdraft/parts/metrics.tex
typo fixed

In this section we come up with some approaches on how to measure uniformity of prediction. The typical way of 'checking' uniformity of prediction used by physicists 
is fitting the distribution of the events that were classified as signal (or background) over the mass (or some other variable).

This approach is hardly formalizable, and not automatable --- each time you should assume some kind of distribution. 
Ideally we want to have some easy-to-use out-of-the-box metrics like FOMs in machine learning (like area under the ROC, f1 or ).

\subsection{Ideal uniformity}

We start from the simplest case --- when we are fully satisfied by predictions of our classifier. Remember that output of 
classification is probabilities of each event being a signal and background event, only after we select some cut on probability we get classification.

Ideal uniformity of signal prediction means that whichever cut we select, the efficiency (part of signal event that passed the cut) doesn't depend on uniform variables: 
in every region of uniform variables space the part of signal events that passed the cut is the same.

In practice, of course, this never happens.

There is uniformity of predictions on signal and uniformity of efficiency on background, which can be defined one from another by swapping classes. 
In what follows in this section we are writing about the efficiency on signal events.

A good example of classifier that has close to ideal uniformity is classifier which returns a random probability in [0, 1] range. What a pity: in practice it's absolutely useless.

% The most significant drawback of the upcoming metrics is they are ill-defined if there are not too many events of the named class.

\subsection{Some Restrictions on Uniformity Metrics}

There are some additional conditions which we expect metric to meet:
\begin{enumerate}
\item
shouldn't depend much on the number of events (i.e., if we randomly select half of the events, the metrics should roughly be the same)
\item
shouldn't depend on the weights renormalization: if we multiply all the weight by some arbitrary number, it shouldn't change at all.
\item 
depends only on the order of predictions, not the exact values of probabilities.
This is because we care about which events pass the cut and which don't, not about the exact values of predictions.

Example: correlation of prediction and mass doesn't satisfy this restriction.
\item
parameter stability: if it uses bins, changing the number of bins shouldn't affect the metrics value much, if it uses $k$-nearest neighbors, it should be stable to small deviations of $k$.
\end{enumerate}

\subsection{Standard Deviation of Efficiency on Bins (SDE)}

\def\bineff{\text{eff}_\text{bin}}
\def\binweight{\text{weight}_\text{bin}}
\def\globaleff{\text{eff}}
\def\SDE{\text{SDE}}
\def\bin{\text{bin}}

Let's split the space of uniform features into bins, the uniformity in these terms is following: when we select some probability cut, 
the part of signal events that passes the cut is equal in all bins. Assume we selected some cut, then we have global efficiency
\[
	\globaleff = \dfrac{
		\text{total weight of signal events that passed the cut}}
		{\text{total weight of signal events}}
\]

Efficiency in every bin is defined respectively, 
\[
	\bineff = \dfrac{
		\text{weight of signal events in bin that passed the cut}}
		{\text{weight of signal events in this bin}} 
\]

So, basically, what we want to have in our dreams:
\[
	\bineff = \text{global efficiency} \qquad \forall \; \text{bin}
\]

To measure how far we are from ideal situation we use standard deviation:
\[
	\sqrt{\sum_{\bin} \left( \bineff - \globaleff \right)^2  }
\]

What is bad in this formula that every bin has some impact in the result, which does not depend on how many events are there, so metrics becomes very unstable to deviations in bins with only few events. To cure this, we add weights to the bins (note that $\sum_\bin \binweight = 1$):
\[
	\binweight = \dfrac{\text{total weight of signal events in bin}}
		{\text{total weight of signal events}},
\]
so we have SDE formula:
\[
	\SDE(\globaleff) = 
	\sqrt{\sum_{\bin} \binweight \times \left(\bineff - \globaleff \right)^2}. 
\] 
In fact, the expression depends on the cut, but for cuts which produce equal efficiency, this is

Finally we note that the weighted average of $\bineff$ is $\globaleff$:
\[
	\globaleff = < \bineff > =  \sum_{\bin} \binweight \times \bineff,
\]
and this is why the introduced metrics was named SDE --- this is a weighted standard deviations of array of bin efficiencies.

But this is how we measure the non-uniformity for only one fixed cut, to measure the overall non-flatness, we take several global efficiencies (for instance, [0.5, 0.6, 0.7, 0.8, 0.9], because in practice usually we are interested in cuts with high global efficiency) and use 
\[
	\SDE^2  =  \frac{1}{k} 
	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
		\text{SDE}^2(\globaleff)
\]

Some other power $p \neq 2$ can be used as well, but $p=2$ is considered as the default value: 
\[
	\SDE^p(\globaleff) = 
	\sum_{\bin} \binweight \times \abs{\bineff - \globaleff}^p,
\qquad
	\SDE^p  =  \frac{1}{k} 
	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
		\text{SDE}^p(\globaleff).
\]

\subsection{Theil Index of Efficiency}
\def\theil{\text{Theil}}

One more measure uses Theil Index frequently used to measure economic inequality:
\[
	\theil = \frac{1}{N} \sum_i \frac{x_i}{<x>} \ln{\frac{x_i}{<x>}}, 
		\qquad <x> = \frac{1}{N} \sum_i x_i
\]
In our case we have to alter formula a bit to take into account that different bins have different impact, thus the formula turns into
\[
	\theil(\globaleff) = \sum_\bin \binweight \; \frac{\bineff}{\globaleff} \; \ln{\frac{\bineff}{\globaleff}}
\]

TODO how to combine Theil for different global efficiencies?
\[
	\theil = ??? \text{from} \theil(\globaleff)
\]

\subsection{Distribution Similarity Approach}
\label{sec:similarity}

Let's start from reformulation of what is uniform predictions in signal. First we split all signal events into some bins in uniform variables. There is some empirical distribution $F_\bin$ of predictions in each bin. Ideal uniformity means that all the distributions $F_\bin$ are equal and hence equal to the global distribution $F(x)$.

To 'measure' non-flatness we can use some distribution distance, like Kolmogorov-Smirnov:
\[
	 \sum_{\bin} \binweight \max_x \abs{F_{\bin}(x) - F(x)},
\]
but Cram\'er--von Mises similarity is more informative (usually $p=2$ is used):
\[
	 \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x),
\]

The good point is we don't need to select some global efficiencies like in the other metrics.

\subsection{Connection Between SDE and Distribution Similarity Approach}

SDE and DSA based on Cram\'er--von Mises similarity can be shown to have connection.
Let's consider the SDE with global efficiencies $= [1/N, 2/N, \dots, N/N]$. In the limit $N \to \infty$
\[
	\lim_{N \to \infty} \SDE^2 = 
	\lim_{N \to \infty} \frac{1}{N} \sum_{\globaleff}\SDE^2(\globaleff) = 
	\int_0^1 \SDE^2(\globaleff) d\, \globaleff = 
	\int_0^1 \sum_{\bin} \binweight \abs{\bineff - \globaleff}^2 d\, \globaleff
\]

From the other side, we can write the expression for similarity-based measure (for $p=2$) 
\[
	\sum_{\bin} \binweight \int \abs{F_\text{bin}(x) - F(x)}^2 dF(x) =
	\int \sum_{\bin} \binweight \abs{F_\text{bin}(x) - F(x)}^2 dF(x) 
\] The hard thing now is to believe this is literally the same and these two expressions are equal.

\subsection{Knn-based modifications}

\def\knni{\text{knn}(i)}
\def\effknni{\text{eff}_{\knni}}
\def\weightknni{\text{weight}_{\knni}}
\def\Fknn{F_{\knni}}

\def\knnSDE{\text{knnSDE}}

Though operating with bins is usually both simple and very efficient, 
in many cases it is hard to find optimal size of bins in the space of uniform variables (specifically in the case of more than two dimensions).
One more situation when bins-based approach fails, is when we have too few events to obtain a good statistics at least in several bins.

In these cases we can switch to $k$-nearest neighbors: for each signal event we find $k$ nearest signal events (including the event itself) in the space of uniform variables. Now we can compute the efficiency $\effknni$, empirical distribution $\Fknn$ of nearest neighbors. 
The weights for $\knni$ are proportional to the total weight of events in $\knni$:
\[
	\weightknni = \alpha \sum_{j \in \knni} w_j, \qquad \alpha^{-1} = \sum_i \sum_{j \in \knni} w_j,
\]
so again weights are normed to 1: $\sum_{i} \knni = 1$.

Now we are ready to write knn version of SDE:
\[
	\knnSDE^2(\globaleff)
		= \sum_{i \in \text{events}} \weightknni \abs{\effknni - \globaleff}^2
\]
\[
	\knnSDE^2 = \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]}
		\knnSDE^2(\globaleff),
\]
knn version of Theil index of Efficiency
\[
	\text{knnTheil}(\globaleff) = \sum_{i \in \text{events}} \weightknni \; \frac{\effknni}{\globaleff} \; \ln{\frac{\effknni}{\globaleff}}
\]
\[
	\text{knnTheil} = ??? \text{knnTheil}(\globaleff)
\]
and knn version of similarity-based measure:

\[
	 \sum_{i \in \text{events}} \weightknni \int \abs{\Fknn(x) - F(x)}^p dF(x),
\]

$K$-nearest neighbors approach suffers from the other drawback: the impact of different events has very little connection with the weights, because some events are met in knn of other events much more frequently while the other.
This effect can be suppressed by dividing initial weight of the event by the number of times it is met in knn.

\subsection{Advantages and Disadvantages of Different Metrics}

... TODO, here some plots with comparison, maybe timings ...

### 2014-08-29 IPythonWorkflow/meanadaboost.py
different voting was added to meanadaboost

@@ -19,7 +19,7 @@ from commonutils import check_sample_weight, computeKnnIndicesOfSameClass
 __author__ = 'Alex Rogozhnikov'
 
 # TODO add some tests and assertions here
-# TODO based on predict algorithm (SAMME)?
+# TODO algorithm based on predict (SAMME)?
 # TODO multiclass classification?
 
 
@@ -31,16 +31,16 @@ class MeanAdaBoostClassifier(BaseEstimator, ClassifierMixin):
                  learning_rate=1.,
                  n_neighbours=10,
                  uniform_label=1,
-                 separate_reweighting=False,
-                 train_variables=None):
+                 train_variables=None,
+                 voting='mean'):
         self.uniform_variables = uniform_variables
         self.base_estimator = base_estimator
         self.n_estimators = n_estimators
         self.learning_rate = learning_rate
         self.n_neighbours = n_neighbours
         self.uniform_label = uniform_label
-        self.separate_reweighting = separate_reweighting
         self.train_variables = train_variables
+        self.voting = voting
 
     def fit(self, X, y, sample_weight=None):
         label = self.uniform_label
@@ -64,24 +64,29 @@ class MeanAdaBoostClassifier(BaseEstimator, ClassifierMixin):
         self.estimators = []
 
         for stage in range(self.n_estimators):
+            knn_scores = numpy.take(cumulative_score, knn_indices)
+            if self.voting == 'mean':
+                voted_score = numpy.mean(knn_scores, axis=1)
+            elif self.voting == 'median':
+                voted_score = numpy.median(knn_scores, axis=1)
+            elif self.voting == 'percentile':
+                voted_score = numpy.percentile(knn_scores, numpy.random.random(), axis=1)
+            else: # self.voting is callable
+                voted_score = self.voting(knn_scores)
+
+            weight = sample_weight * numpy.exp(- y_signed * voted_score)
+            weight = self.normalize_weights(y=y, sample_weight=weight)
+
             classifier = sklearn.clone(self.base_estimator)
-            classifier.fit(X, y, sample_weight=sample_weight)
-            score = self.learning_rate * self.compute_score(classifier, X=X)
-            cumulative_score += score
-            sample_weight *= numpy.exp(- y_signed * numpy.take(score, knn_indices).mean(axis=1))
-            sample_weight = self.normalize_weights(y=y, sample_weight=sample_weight)
[…]

### 2014-08-29 IPythonWorkflow/metrics.py
minor fix to theil metrics

@@ -435,7 +435,7 @@ def theil(x, weights):
     """Theil index of array with regularization"""
     assert numpy.all(x >= 0)
     x_mean = numpy.average(x, weights=weights)
-    normed = x / x_mean
+    normed = x / (x_mean + 1e-100)
     normed[normed < 1e-10] = 1e-10
     return numpy.average(normed * numpy.log(normed), weights=weights)
 

### 2014-09-08 IPythonWorkflow/ugradientboosting.py
fixed computations of out-of-all-bins events

@@ -17,7 +17,7 @@ from sklearn.utils.validation import check_arrays, column_or_1d
 
 from commonutils import check_sample_weight, computeSignalKnnIndices
 import commonutils
-from metrics import compute_group_weights, compute_divided_weight
+from metrics import compute_group_weights
 import metrics
 
 
@@ -281,10 +281,10 @@ class AbstractFlatnessLossFunction(AbstractLossFunction):
         for label in self.uniform_label:
             self.group_indices[label] = self.compute_groups_indices(X, y, label=label)
             self.group_weights[label] = compute_group_weights(self.group_indices[label], sample_weight=sample_weight)
-            for group in self.group_indices:
+            for group in self.group_indices[label]:
                 occurences[group] += 1
 
-        out_of_bins = (occurences == 0) & (~numpy.in1d(y, self.uniform_label))
+        out_of_bins = (occurences == 0) & numpy.in1d(y, self.uniform_label)
         if numpy.mean(out_of_bins) > 0.01:
             warnings.warn("%i events out of all bins " % numpy.sum(out_of_bins), UserWarning)
 
@@ -301,30 +301,8 @@ class AbstractFlatnessLossFunction(AbstractLossFunction):
         raise NotImplementedError()
 
     def __call__(self, pred):
-        # computing the common distribution of signal
-        # taking only signal by now
-        # this is approximate computation!
-
-        #TODO reimplement, this is wrong implementation
+        #TODO implement
         return 0
-        loss = 0
-
-        for label in self.uniform_label:
-            label_mask = self.y == label
-            sorted_pred = numpy.sort(pred[label_mask])
-
-            for group_weight, group_indices in zip(self.group_weights[label], self.group_indices[label]):
-                probs_in_bin = numpy.take(pred, group_indices)
-                probs_in_bin = numpy.sort(probs_in_bin)
-                positions = numpy.searchsorted(sorted_pred, probs_in_bin)
-                global_effs = positions / float(len(sorted_pred))
-                local_effs = (numpy.arange(0, len(probs_in_bin)) + 0.5) / len(probs_in_bin)
-                bin_loss = numpy.sum((global_effs - local_effs) ** self.power)
-                loss += bin_loss * group_weight
-
-        # Ada loss now
-        loss += self.ada_coefficient * numpy.sum(numpy.exp(-self.y_signed * pred))
-        return loss
 
     def negative_gradient(self, y_pred):
         y_pred = numpy.ravel(y_pred)

### 2014-09-09 assets/paperdraft/paper_draft.tex
Mike edits

@@
 
-\title{New Approaches to Uniform Boosting}
+\title{New approaches for boosting to uniformity}
 
 \author{
-Aleksandar Bukva, 
+Aleksandar Bukva$^{\rm ADD}$, 
 Vladimir Gligorov$^d$,
@@
 % AMS packages:
-\usepackage{amsmath, amsthm, amsfonts}
+\usepackage{amsmath, amsthm, amsfonts, xspace}
 
@@
 
+\def \x {\ensuremath{\vec{x}}\xspace}
+\def \y {\ensuremath{\vec{y}}\xspace}
+\def\knn{\text{knn}}
+\def\knni{\text{knn}(i)}
+
 \begin{document}
@@
 
-\section{Uniformity Measurement}
+%\section{Uniformity Measurement}
+%\input{parts/metrics}
+%\section{Approaches Proposed}
+%\input{parts/approaches}
 
-\input{parts/metrics}
+\section{Uniformity Boosting Methods}
+\input{parts/methods}
 
-\section{Approaches Proposed}
-
-\input{parts/approaches}
-
-\section{Dataset}
+\section{Example Analysis}
 
 \input{parts/dataset}
-
-\section{Plots}
-
 \input{parts/plots}
@@
 
+%\textcolor{red}{xxx}
+
 The link on the repository ans links to final notebooks with results in the article.
@@
 
-% The examples of items:
-% \bibitem{bib1}
-% Authors,
-% \emph{Title},
-% \emph{J. Ref.} \textbf{vol} (year) page.
+\bibitem{ref:lhcbhlt} R. Aaij {\em et al.} [LHCb Trigger Group], {\em The LHCb trigger and its performance}, JINST {\bf 8}, P04022 (2013). [arXiv:1211.3055]
 
-% \bibitem{bib2}
-% A. Pietropaolo et al., \emph{DINS measurements on VESUVIO in the
-%     resonance detector configuration: proton mean kinetic energy in
-%     water}, \jinst{1}{2006}{P04001}.
+\bibitem{ref:bbdt}  V. Gligorov and M. Williams, {\em Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree}, JINST {\bf 8}, P02013 (2013). [arXiv:1210.6861]
 
-% \bibitem{bib3}
-% A.I. Harris,
-% \emph{Spectroscopy with multichannel correlation radiometers},
-% \href{http://dx.doi.org/10.1063/1.1898643}
-% {\emph{Rev.\ Sci.\ Instrum.} {\bf 76} (2005) 054503}
-% [\astroph{0504449}].
-
-% \bibitem{bib4}
-% G.F. Knoll, \emph{Radiation detection and measurements}, John Wiley
-%     and Sons, Inc., New York 2000.
-
-% \bibitem{bib5}
-% V. Dangendorf, \emph{Time-resolved fast-neutron imaging with a
-% pulse-counting image intensifier}, in proceedings of
-% \emph{International workshop on fast neutron detectors and
-% applications}, April, 3--6, 2006 University of Cape Town, South Africa
-% \pos{PoS(FNDA2006)008}.
+\bibitem{ref:uboost} J.
+Stevens and M.
+Williams, {\em uBoost: A boosting method for producing uniform selection efficiencies from multivariate classifiers}, JINST {\bf 8}, P12013 (2013) [arXiv:1305.7248]
 
@@
 
+\appendix
+\section{Other Flatness Loss Criteria}
+
+ADD THEIL, ETC DESCRIPTIONS AND HOW TO IMPLEMENT IN THE ALGORITHM HERE.
+
 \end{document}

### 2014-09-09 assets/paperdraft/ugb.tex
Mike edits

The variates used in the BDT are denoted by \x, while the variates in which uniformity is desired are denoted by \y.  Some (perhaps all) of the \x variantes will be {\em biasing} in \y, {\em i.e.} they provide discriminating power between signal and background that varies in \y.  A uniform BDT selection efficiency can be obtained by removing all such variates; however, this will also reduce the power of the BDT.  The goal of boosting algorithms presented in this paper is to {\em balance} the biases to produce the optimal uniform selection.

Traditional boosting works by assigning training events more weight based on classification erros made by previous members of the series.  For example, in gradient boosting~\cite{ref:gradboost} the analyst chooses a differentiable loss function with the goal of building a classifier that minimizes the loss.  
A popular choice of loss function is the AdaBoost~\cite{ref:adaboost} loss function defined as follows:
\begin{equation}
L_{\rm ada} = \sum\limits_{i=1}^{n}w_i \times {\rm exp}\left[-\gamma_i s_i\right], 
\end{equation}
where $w$ is the weight of each event\footnote{We use the convention that $\sum w_i = 1$ for both signal and background events at each stage of training.}, $\gamma = +1(-1)$ for signal(background) events and $s$ is the so-called score of each event which is obtained as the sum of scores of all of the previous classifiers in the series.   
At each stage in the gradient boosting process, a {\em regressor} (a decision tree in our case) is trained whose purpose is to decrease the loss.  This is accomplished using the gradient decent method and the {\em pseudo-residuals}
\begin{equation}
  -\frac{\partial L_{\rm ada}}{\partial s_i} = w_i \gamma_i  {\rm exp}\left[-\gamma_i s_i\right],
\end{equation}
which are positive(negative) for signal(background) events and have larger modulai for poorly classified events.

The gradient-boosting algorithm is general in that it only requires the analyst specify a loss function and its gradient.  The AdaBoost loss function considers each event individually, but can easily be modified to take into account {\em non-local} properties of the classifier as follows:
\begin{equation}
  \label{eq:loss_general}
  L_{\rm general} = \sum\limits_{i=1}^{n} w_i {\rm exp}\left[-\gamma_i \sum\limits_{j} a_{ij} s_j \right], 
\end{equation} 
where $A$ is a matrix\footnote{A natural choice is a square $n\times n$ matrix, but this is not required.}.  For the case where $A$ is the identity matrix, the AdaBoost loss function is recovered.  
Other choices of $A$ will induce non-local effects, {\em e.g.}, consider the sparse matrix 
\begin{equation}
  a_{ij}^{\rm knn} = \begin{cases} 
    1, & j \in \knni, \text{ events $i$ and $j$ belong to the same class} \\
    0, & \text{otherwise},
\end{cases}
\end{equation}
where $\knn(i)$ denotes the set of $k$-nearest-neighbor events to event $i$.   The loss function obtained from Eq.~\ref{eq:loss_general} using $A^{\rm knn}$, denoted by $L_{\rm knn}$, accounts for the score of each event's $k$ nearest neighbors and not just each event individually.
The pseudo-residuals are then
\begin{equation}
  -\frac{\partial L_{\rm knn}}{\partial s_k} = -\sum\limits_i w_i \gamma_i a_{ik} {\rm exp}\left[-\gamma_i \sum_j a_{ij} s_j \right].
\end{equation}
One can see that the direction of the gradient will be influenced the most by events whose $k$-nearest-neighbor events are classified poorly.

Another approach is to include in the definition of the loss function some uniformity metric.  Consider first the case where the data has been binned in \y.  If the distribution of classifier responses in each bin, $f_b(s)$, is the same as the global response distribution, $f(s)$, then any cut made on the response will produce a uniform selection efficiency in \y.  Therefore, performing a one-dimensional goodness-of-fit test of the hypothesis $f_b \equiv f$ in each bin provides an assessment of the selection uniformity.  
For example, one could perform the Kolmogorov-Smirnov test in each bin and define a loss function as follows:
\begin{equation}
  L_{\rm flat(KS)} = \sum\limits_{b} w_b {\rm max}|F_b(s)-F(s)|,
\end{equation}
where $F_{(b)}(s)$ denotes the cummulative distribution of $f_{(b)}(s)$ and $w_b = \sum \delta({\rm bin}_i - b) w_i$, {\em i.e.} $w_b$ is the sum of the weights of the signal events in the bin.

The gradient of the Kolmogorov-Smirnov loss function is zero for events with responses greater than the value of $s$ at which ${\rm max}|F_b(s)-F(s)|$ occurs.  Thus, it is not suitable for gradient boosting due to its instability.  Instead, we use the following loss function:
\begin{equation}
  L_{\rm flat} = \sum\limits_{b} w_b \sum\limits_{i \in b}|F_b(s_i)-F(s_i)|,
\end{equation}
whose pseudo-residuals are
\begin{equation}
 -\frac{\partial L_{\rm flat}}{\partial s_k} = - w_b {\rm sign}\left[F_b(s_k)-F(s_k)\right] \left( w_k/w_b - w_k \right).  
\end{equation}
This so-called flatness loss penalizes non-uniformity but does not consider the quality of the classification.  Therefore, the full loss function used is 
\begin{equation}
\label{eq:L_ada_flat}
  L_{{\rm ada}+{\rm flat}} = L_{\rm flat} + \alpha L_{\rm ada},
\end{equation}  
where $\alpha$ is a real-valued parameter that is typically chosen to be small.  The first term in Eq.~\ref{eq:L_ada_flat} penalizes non-uniformity, while that second term penalizes poor classification.

ADD KNN HERE ONCE ALGORITHM IS TESTED.

### 2014-09-17 assets/paperdraft/paper_draft.tex

@@
 \author{
-Aleksandar Bukva$^{\rm ADD}$, 
+Aleksandar Bukva$^e$, 
 Vladimir Gligorov$^d$,
@@
 \llap{$^d$}Organisation Europ\'eenne pour la Recherche Nucl\'eaire (CERN), Geneva, Switzerland  \\
+\llap{$^e$}Faculty of Physics, Belgrade, Serbia \\
 E-mail: \email{alex.rogozhnikov@yandex.ru}}
@@
 % AMS packages:
+
 \usepackage{amsmath, amsthm, amsfonts, xspace}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{float}
 
@@
 
-\def \x {\ensuremath{\vec{x}}\xspace}
-\def \y {\ensuremath{\vec{y}}\xspace}
+\def\x {\ensuremath{\vec{x}}\xspace}
+\def\y {\ensuremath{\vec{y}}\xspace}
 \def\knn{\text{knn}}
@@
 
-%\section{Uniformity Measurement}
-%\input{parts/metrics}
-%\section{Approaches Proposed}
-%\input{parts/approaches}
-
 \section{Uniformity Boosting Methods}
-\input{parts/methods}
+\input{ugb}
 
@@
 
-ADD THEIL, ETC DESCRIPTIONS AND HOW TO IMPLEMENT IN THE ALGORITHM HERE.
+Clean all this up...
+
+%\section{Uniformity Measurement}
+\input{parts/metrics}
+%\section{Approaches Proposed}
+\input{parts/approaches}
 

### 2014-09-17 assets/paperdraft/parts/metrics.tex

@@
 \subsection{Advantages and Disadvantages of Different Metrics}
+Let us compare two metrics that have shown most appropriate for our problem, SDE and Theil. We have some linearly spaced array of mass of particles
+from interval $[0,1]$, some constant $\alpha$ linearly sapced in interval $[.1,1]$. The predictions are correlated with mass via beta distribution.
+We have two distributions that are different in a way that they are symmetrical.
+SDE should show no changes if we flip the distribution, while the Theil should make difference between pit and peak.
+First distribution with a peak is given by:
 
-... TODO, here some plots with comparison, maybe timings ...
+\[
+	p\text{[i]}=\textbf{beta}(\alpha \cdot \bar m, \text{abs}(m\text[i]-\bar m))
+\]
+
+while the second one has minus sign in front.
+\begin{figure}[H]
+\centering
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/distribution1.png}
+			\caption{Mass vs prediction}
+		\end{subfigure}
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/efficiencies1.png}
+			\caption{Efficiencies}
+		\end{subfigure}
+		\caption{Peak distribution}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+	\includegraphics[width=8cm]{graphs/measures1.png}
+	\caption{SDE and Theil}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/distribution2.png}
+			\caption{Mass vs prediction}
+		\end{subfigure}
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/efficiencies2.png}
+			\caption{Efficiencies}
+		\end{subfigure}
+		\caption{Pit distribution}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+	\includegraphics[width=8cm]{graphs/measures2.png}
+	\caption{SDE and Theil}
+\end{figure}
+
+From these figures we can see that SDE has the same value and it is not able to make difference between these distributions.
+In the second case Theil has lower value which indicates that distribution is flatter.
+In the combination with SDE it can be used to detarmine the type of a distribution.
+Lower value of SDE can indicate that distribution is flat or has some narrow peaks or pits, and by using Theil we can distinguish between these two cases.

### 2014-09-17 assets/paperdraft/ugb.tex

@@
 
-Traditional boosting works by assigning training events more weight based on classification erros made by previous members of the series.
-For example, in gradient boosting~\cite{ref:gradboost} the analyst chooses a differentiable loss function with the goal of building a classifier that minimizes the loss.
-A popular choice of loss function is the AdaBoost~\cite{ref:adaboost} loss function defined as follows:
+One category of boosting works by assigning training events more weight based on classification erros made by previous members of the series.
+For example, the AdaBoost~\cite{ref:adaboost} algorithm updates the weight of event $i$, $w_i$, according to
 \begin{equation}
-L_{\rm ada} = \sum\limits_{i=1}^{n}w_i \times {\rm exp}\left[-\gamma_i s_i\right], 
+w_i^{\prime} = w_i \times {\rm exp}\left[-\gamma_i s_i\right], 
 \end{equation}
-where $w$ is the weight of each event\footnote{We use the convention that $\sum w_i = 1$ for both signal and background events at each stage of training.}, $\gamma = +1(-1)$ for signal(background) events and $s$ is the so-called score of each event which is obtained as the sum of scores of all of the previous classifiers in the series.
+where $\gamma = +1(-1)$ for signal(background) events and $s$ is the so-called score of each event which is obtained as the sum of scores of all of the previous classifiers in the series.  
+The uBoost technique, described in detail in Ref.~\cite{cit:uboost}, alters the event-weight updating procedure to achieve uniformity in the signal-selection efficiency.
+
+Another approach to obtain uniformity, introduced in this paper, involves defining a more general expression of the AdaBoost criteria:
+\begin{equation}
+ w_i^{\prime} = w_i \times {\rm exp}\left[-\gamma_i \sum_j a_{ij}\right],
+\end{equation}
+where $a_{ij}$ are the elements of some matrix $A$\footnote{A natural choice is a square $n\times n$ matrix, but this is not required.}.
+For the case where $A$ is the identity matrix, the AdaBoost weighting procedure is recovered.
+Other choices of $A$ will induce non-local effects, {\em e.g.}, consider the sparse matrix $A_{\rm knn}$ given by
+\begin{equation}
+  a_{ij}^{\rm knn} = \begin{cases} 
+    \frac{1}{k}, & j \in \knni, \text{ events $i$ and $j$ belong to the same class} \\
+    0, & \text{otherwise},
+\end{cases}
+\end{equation}
+where $\knn(i)$ denotes the set of $k$-nearest-neighbor events to event $i$.
+This procedure for updating the event weights, which we refer to as kNNAdaBoost, accounts for the score of each event's $k$ nearest neighbors and not just each event individually.
+
+The gradient boosting~\cite{ref:gradboost} (GB) algorithm category requires the analyst to choose a differentiable loss function with the goal of building a classifier that minimizes the loss.  
+A popular choice of loss function is the AdaLoss~\cite{ref:adaboost} function 
+\begin{equation}
+L_{\rm ada} = \sum\limits_{i=1}^{n} {\rm exp}\left[-\gamma_i s_i\right]. 
+\end{equation}
 At each stage in the gradient boosting process, a {\em regressor} (a decision tree in our case) is trained whose purpose is to decrease the loss.
@@
 \begin{equation}
-  -\frac{\partial L_{\rm ada}}{\partial s_i} = w_i \gamma_i  {\rm exp}\left[-\gamma_i s_i\right],
+  -\frac{\partial L_{\rm ada}}{\partial s_i} = \gamma_i  {\rm exp}\left[-\gamma_i s_i\right],
 \end{equation}
@@
 The gradient-boosting algorithm is general in that it only requires the analyst specify a loss function and its gradient.
-The AdaBoost loss function considers each event individually, but can easily be modified to take into account {\em non-local} properties of the classifier as follows:
+The AdaLoss function considers each event individually, but can easily be modified to take into account non-local properties of the classifier as follows:
 \begin{equation}
   \label{eq:loss_general}
-  L_{\rm general} = \sum\limits_{i=1}^{n} w_i {\rm exp}\left[-\gamma_i \sum\limits_{j} a_{ij} s_j \right], 
+  L_{\rm general} = \sum\limits_{i=1}^{n} {\rm exp}\left[-\gamma_i \sum\limits_{j} a_{ij} s_j \right]. 
 \end{equation} 
-where $A$ is a matrix\footnote{A natural choice is a square $n\times n$ matrix, but this is not required.}.  For the case where $A$ is the identity matrix, the AdaBoost loss function is recovered.  
-Other choices of $A$ will induce non-local effects, {\em e.g.}, consider the sparse matrix 
+For example, the loss function obtained from Eq.~\ref{eq:loss_general} using $A_{\rm knn}$, which we refer to as kNNAdaLoss and denote $L_{\rm knn}$, accounts for the score of each event's $k$ nearest neighbors and not just each event individually.
+The pseudo-residuals of $L_{\rm knn}$ are 
 \begin{equation}
-  a_{ij}^{\rm knn} = \begin{cases} 
-    1, & j \in \knni, \text{ events $i$ and $j$ belong to the same class} \\
-    0, & \text{otherwise},
-\end{cases}
-\end{equation}
-where $\knn(i)$ denotes the set of $k$-nearest-neighbor events to event $i$.
-The loss function obtained from Eq.~\ref{eq:loss_general} using $A^{\rm knn}$, denoted by $L_{\rm knn}$, accounts for the score of each event's $k$ nearest neighbors and not just each event individually.
-The pseudo-residuals are then
-\begin{equation}
-  -\frac{\partial L_{\rm knn}}{\partial s_k} = -\sum\limits_i w_i \gamma_i a_{ik} {\rm exp}\left[-\gamma_i \sum_j a_{ij} s_j \right].
+  -\frac{\partial L_{\rm knn}}{\partial s_k} = -\sum\limits_i \gamma_i a_{ik}^{\rm knn} {\rm exp}\left[-\gamma_i \sum_j a_{ij}^{\rm knn} s_j \right].
 \end{equation}
@@
 Another approach is to include in the definition of the loss function some uniformity metric.
-Consider first the case where the data has been binned in \y.
+Consider first the case where the data have been binned in \y.
 If the distribution of classifier responses in each bin, $f_b(s)$, is the same as the global response distribution, $f(s)$, then any cut made on the response will produce a uniform selection efficiency in \y.
@@
 \end{equation}
-where $F_{(b)}(s)$ denotes the cummulative distribution of $f_{(b)}(s)$ and $w_b = \sum \delta({\rm bin}_i - b) w_i$, {\em i.e.} $w_b$ is the sum of the weights of the signal events in the bin.
+where $F_{(b)}(s)$ denotes the cummulative distribution of $f_{(b)}(s)$ and $w_b = \sum \delta({\rm bin}_i - b)/n_{\rm signal}$, {\em i.e.} $w_b$ is the fraction of signal events in the bin\footnote{If weighted events are used, then the fractional sum of weights should be used for $w_b$.}.
 
@@
 Thus, it is not suitable for gradient boosting due to its instability.
-Instead, we use the following loss function:
+Instead, we use the following {\em flatness} loss function:
 \begin{equation}
-  L_{\rm flat} = \sum\limits_{b} w_b \sum\limits_{i \in b}|F_b(s_i)-F(s_i)|,
+  L_{\rm flat} = \sum\limits_{b} w_b \int |F_b(s)-F(s)|^2 ds,
 \end{equation}
@@
 \begin{equation}
- -\frac{\partial L_{\rm flat}}{\partial s_k} = - w_b {\rm sign}\left[F_b(s_k)-F(s_k)\right] \left( w_k/w_b - w_k \right).  
+ -\frac{\partial L_{\rm flat}}{\partial s_k} \approx - 2\left[F_b(s_k)-F(s_k)\right]/n_{\rm signal}.
 \end{equation}
@@
 
-ADD KNN HERE ONCE ALGORITHM IS TESTED.
+The loss function given in Eq.~\ref{eq:L_ada_flat} can also be constructed without binning the data using k-nearest-neighbor events.
+The cummulative distribution $F_{\rm knn}(s)$ is easily obtained and the bin weight, $w_b$, is replaced by ???

### 2014-09-18 assets/paperdraft/parts/dataset.tex
First draft

\section{Datasets}
This study uses simulated event samples produced using the official LHCb simulation framework.
The software used for the generation of the events is described in LHCb publications as follows :

\begin{quote}
In the simulation, $pp$ collisions are generated using
PYTHIA~\cite{Sjostrand:2006za} 
with a specific LHCb configuration~\cite{LHCb-PROC-2010-056}.  Decays of hadronic particles
are described by EvtGen~\cite{Lange:2001uf}, in which final state
radiation is generated using PHOTOS~\cite{Golonka:2005pn}. The
interaction of the generated particles with the detector and its
response are implemented using the GEANT
toolkit~\cite{Allison:2006ve, Agostinelli:2002hh} as described in
Ref.~\cite{LHCb-PROC-2011-006}.
\end{quote}

All simulated event samples are generated inside the LHCb detector acceptance.
The signal used in this analysis consists of $D_s^\pm\to\pi^+\pi^-\pi^\pm$ decays, simulated
using the $\textrm{D}\_\textrm{DALITZ}$ model of EvtGen to simulate the intermediate resonances which contribute to the
three pion final state. The background candidates are three pion combinations reconstructed in
simulated samples of $c\bar{c}$ and $b\bar{b}$ events, where the charm and bottom quark decays are
inclusively modelled by EvtGen. The simulated events contain ``truth'' information which identifies them as
signal or background, and which identifies the physical origin of the three pion combinations reconstructed
in the $c\bar{c}$ and $b\bar{b}$ simulated samples.

### 2014-09-18 assets/paperdraft/paper_draft.tex
Add references

--- 2014-08-25 assets/paperdraft/paper_draft.tex
@@
 \author{
-Aleksandar Bukva, 
+Aleksandar Bukva$^e$, 
 Vladimir Gligorov$^d$,
@@
 \usepackage{amsmath, amsthm, amsfonts}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{float}
 
@@
 
+These results were obtained using events generated with the official LHCb simulation, and we are grateful to the LHCb collaboration for this privilege. 
+We particularly acknowledge the work of the LHCb Simulation and Core Computing teams in tuning the simulation software and managing the productions of simulated events.
+
 \begin{thebibliography}{9}
+
+\bibitem{Sjostrand:2006za}
+{"Sj\"{o}strand, Torbj\"{o}rn and Mrenna, Stephen and Skands, Peter"},
+\emph{PYTHIA 6.4 physics and manual},
+\href{http://dx.doi.org/10.1088/1126-6708/2006/05/026},
+{\emph{JHEP} {\bf 05} (2006) 026},
+[\hepph{0603175}].
+
+\bibitem{LHCb-PROC-2010-056}
+{"Belyaev, I. and others"},
+\emph{Handling of the generation of primary events in GAUSS, the LHCb simulation framework},
+\href{http://dx.doi.org/10.1109/NSSMIC.2010.5873949},
+{\emph{NSS/MIC} {\bf IEEE} (2010) 1155}.
+
+\bibitem{Lange:2001uf}
+{"Lange, D. J."},
+\emph{The EvtGen particle decay simulation package},
+\href{http://dx.doi.org/10.1016/S0168-9002(01)00089-4},
+{\emph{NIM} {\bf A462} (2001) 152-155}.
+
+\bibitem{Golonka:2005pn}
+{"Golonka, Piotr and Was, Zbigniew"},
+\emph{PHOTOS Monte Carlo: a precision tool for QED corrections in $Z$ and $W$ decays},
+\href{http://dx.doi.org/10.1140/epjc/s2005-02396-4},
+{\emph{EPJ} {\bf C45} (2006) 97-107},
+[\hepph{0506026}].
+
+\bibitem{Allison:2006ve}
+{"Allison, John and Amako, K. and Apostolakis, J. and Araujo, H. and Dubois, P.A. and others"},
+\emph{Geant4 developments and applications},
+\href{http://dx.doi.org/10.1109/TNS.2006.869826},
+{\emph{IEEE TNS} {\bf 53} (2006) 270}.
+
+\bibitem{Agostinelli:2002hh}
+{"Agostinelli, S. and others"},
+\emph{Geant4: a simulation toolkit},
+\href{http://dx.doi.org/10.1016/S0168-9002(03)01368-8},
+{\emph{NIM} {\bf A506} (2003) 250}.
+
+\bibitem{LHCb-PROC-2011-006}
+{"Clemencic, M and others"},
+\emph{The LHCb simulation application, GAUSS : design, evolution and experience},
+\href{http://dx.doi.org/10.1088/1742-6596/331/3/032023},
+{\emph{J. Phys. Conf. Ser.} {\bf 331} (2011) 032023}.
 

### 2014-09-18 assets/paperdraft/paper_draft.tex

@@
 
-\title{New Approaches to Uniform Boosting}
+\title{New approaches for boosting to uniformity}
 
@@
 % AMS packages:
-\usepackage{amsmath, amsthm, amsfonts}
+
+\usepackage{amsmath, amsthm, amsfonts, xspace}
 \usepackage{caption}
@@
 
+\def\x {\ensuremath{\vec{x}}\xspace}
+\def\y {\ensuremath{\vec{y}}\xspace}
+\def\knn{\text{knn}}
+\def\knni{\text{knn}(i)}
+
 \begin{document}
@@
 
-\section{Uniformity Measurement}
+\section{Uniformity Boosting Methods}
+\input{ugb}
 
-\input{parts/metrics}
-
-\section{Approaches Proposed}
-
-\input{parts/approaches}
-
-\section{Dataset}
+\section{Example Analysis}
 
 \input{parts/dataset}
-
-\section{Plots}
-
 \input{parts/plots}
@@
 
+\bibitem{ref:lhcbhlt} 
+R. Aaij {\em et al.} [LHCb Trigger Group], 
+{\em The LHCb trigger and its performance}, 
+JINST {\bf 8}, P04022 (2013). 
+[arXiv:1211.3055]
+
+\bibitem{ref:bbdt}  
+V. Gligorov and M. Williams, 
+{\em Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree}, 
+JINST {\bf 8}, P02013 (2013). 
+[arXiv:1210.6861]
+
+\bibitem{ref:uboost} 
+J. Stevens and M. Williams,
+{\em uBoost: A boosting method for producing uniform selection efficiencies from multivariate classifiers}, 
+JINST {\bf 8}, P12013 (2013). 
+[arXiv:1305.7248]
+
 \bibitem{Sjostrand:2006za}
@@
 \href{http://dx.doi.org/10.1088/1126-6708/2006/05/026},
-{\emph{JHEP} {\bf 05} (2006) 026},
-[\hepph{0603175}].
+{\emph{JHEP} {\bf 05} (2006) 026}.
+[\hepph{0603175}]
 
@@
 \href{http://dx.doi.org/10.1140/epjc/s2005-02396-4},
-{\emph{EPJ} {\bf C45} (2006) 97-107},
-[\hepph{0506026}].
+{\emph{EPJ} {\bf C45} (2006) 97-107}.
+[\hepph{0506026}]
 
@@
 
+\appendix
+\section{Other Flatness Loss Criteria}
+
+Clean all this up...
+
+%\section{Uniformity Measurement}
+\input{parts/metrics}
+%\section{Approaches Proposed}
+\input{parts/approaches}
+
 \end{document}

### 2014-09-19 assets/paperdraft/paper_draft.tex
ignore temp files, author list update

--- 2014-09-17 assets/paperdraft/paper_draft.tex
@@
 Vladimir Gligorov$^d$,
-Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~ ,
-Andrey Ustuzhanin$^b$ and
+Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~,\newline
+Andrey Ustyuzhanin$^{b,f}$ and
 Mike Williams$^c$\\
 \llap{$^a$}Lomonosov Moscow State University, Moscow, Russia\\
-\llap{$^b$}Yandex LLC, Moscow, Russia\\
+\llap{$^b$}Yandex, Moscow, Russia\\
 \llap{$^c$}Massachusetts Institute of Technology, Cambridge, MA, United States \\
@@
 \llap{$^e$}Faculty of Physics, Belgrade, Serbia \\
+\llap{$^f$}Moscow Institute of Physics and Technology, Moscow, Russia\\
 E-mail: \email{alex.rogozhnikov@yandex.ru}}

### 2014-09-21 assets/paperdraft/parts/metrics.tex
Major textual changes to the section on uniformity metrics.

@@
-In this section we come up with some approaches on how to measure uniformity of prediction. The typical way of 'checking' uniformity of prediction used by physicists 
-is fitting the distribution of the events that were classified as signal (or background) over the mass (or some other variable).
+\section{Measures of uniformity}
+In this section we discuss different methods for measuring the uniformity of prediction. 
+One typical way of 'checking' uniformity of prediction used by physicists 
+is fitting the distribution of the events that were classified as signal (or background) over the 
+feature for which you wish to check uniformity.
+This approach requires assumptions about the shape of the distribution,
+which makes quantitative comparisons of different classifiers difficult.
+Our aim here is to explore uniformity figures of merit which make comparing classifiers easier,
+analogously to how the area under the ROC curve can be used to compare absolute classifier performance.
+%is hardly formalizable, and not automatable --- each time you should assume some kind of distribution. 
+%Ideally we want to have some easy-to-use out-of-the-box metrics like FOMs in machine learning (like area under the ROC, f1 or ).
 
-This approach is hardly formalizable, and not automatable --- each time you should assume some kind of distribution. 
-Ideally we want to have some easy-to-use out-of-the-box metrics like FOMs in machine learning (like area under the ROC, f1 or ).
+%We start from the simplest case --- when we are fully satisfied by predictions of our classifier. 
+The output of event classification is the probability of each event being signal or background,
+and it is only after we apply a cut on this probability that events are classified.
+An ideal uniformity of signal prediction can then be defined for a given ``uniform feature'' of interest.
+It means that whichever cut we select,
+the efficiency for a signal event to pass the cut doesn't depend on the uniform feature.
+Uniformity for background can be defined in the same manner, but for simplicity,
+% (uniform variables) 
+%in every region of the uniform variables space the part of signal events that passed the cut is the same.
+%In practice, of course, this never happens.
+%There is uniformity of predictions on signal and uniformity of efficiency on background, which can be defined one from another by swapping classes. 
+in what follows we will only discuss the uniformity of efficiency for signal events.
 
-\subsection{Ideal uniformity}
-
-We start from the simplest case --- when we are fully satisfied by predictions of our classifier. Remember that output of 
-classification is probabilities of each event being a signal and background event, only after we select some cut on probability we get classification.
-
-Ideal uniformity of signal prediction means that whichever cut we select, the efficiency (part of signal event that passed the cut) doesn't depend on uniform variables: 
-in every region of uniform variables space the part of signal events that passed the cut is the same.
-
-In practice, of course, this never happens.
-
-There is uniformity of predictions on signal and uniformity of efficiency on background, which can be defined one from another by swapping classes. 
-In what follows in this section we are writing about the efficiency on signal events.
-
-A good example of classifier that has close to ideal uniformity is classifier which returns a random probability in [0, 1] range. What a pity: in practice it's absolutely useless.
-
+A trivial example of a classifier that has ideal uniformity is a classifier which returns a random classification probability,
+but such a classifier is of course not very useful. One can try to design a uniform classifier with respect to
+a given feature by not using this feature, or any correlated features, in the classification; in practice, however,
+this approach also tends to lead to poorly performing classifiers. The approach which we take in this paper is 
+to explicitly let the classifier learn how to balance non-uniformities coming from different features in such a way
+as to generate a classification which is uniform on average. It is then important to be able to accurately measure
+the uniformity of classification. 
+%What a pity: in practice it's absolutely useless.
 % The most significant drawback of the upcoming metrics is they are ill-defined if there are not too many events of the named class.
 
-\subsection{Some Restrictions on Uniformity Metrics}
-
-There are some additional conditions which we expect metric to meet:
+Before proceeding, it is useful to define some desirable properties of uniformity metrics
 \begin{enumerate}
 \item
-shouldn't depend much on the number of events (i.e., if we randomly select half of the events, the metrics should roughly be the same)
+The metric shouldn't depend strongly on the number of events used to test uniformity;
+%(i.e., if we randomly select half of the events, the metrics should roughly be the same)
 \item
-shouldn't depend on the weights renormalization: if we multiply all the weight by some arbitrary number, it shouldn't change at all.
+The metric shouldn't depend on the normalization of the event weights: if we multiply all the weights by some arbitrary number, it shouldn't change at all;
 \item 
-depends only on the order of predictions, not the exact values of probabilities.
+The metric should depend only on the order of predictions, not the exact values of probabilities.
 This is because we care about which events pass the cut and which don't, not about the exact values of predictions.
-
-Example: correlation of prediction and mass doesn't satisfy this restriction.
+For example: correlation of prediction and mass doesn't satisfy this restriction.
 \item
-parameter stability: if it uses bins, changing the number of bins shouldn't affect the metrics value much, if it uses $k$-nearest neighbors, it should be stable to small deviations of $k$.
+The metric should be stable against any of its own free paramters : if it uses bins, changing the number of bins shouldn't affect the result,
+if it uses $k$-nearest neighbors, it should be stable against different values of $k$.
 \end{enumerate}
+In what follows we will consider different metrics which satisfy these criteria, and then compare their performance in
+some test cases.
 
@@
 
-Let's split the space of uniform features into bins, the uniformity in these terms is following: when we select some probability cut, 
-the part of signal events that passes the cut is equal in all bins. Assume we selected some cut, then we have global efficiency
+If the space of uniform features is split into bins, it is possible to define the global efficiency
+%  when we select some probability cut, 
+%the part of signal events that passes the cut is equal in all bins. Assume we selected some cut, then we have global efficiency
 \[
@@
 		\text{total weight of signal events that passed the cut}}
-		{\text{total weight of signal events}}
+		{\text{total weight of signal events}},
 \]
 
-Efficiency in every bin is defined respectively, 
+as well as the efficiency in every bin, 
 \[
@@
 		\text{weight of signal events in bin that passed the cut}}
-		{\text{weight of signal events in this bin}} 
+		{\text{weight of signal events in this bin}}.
+\]
+%So, basically, what we want to have in our dreams:
+%\[
+%	\bineff = \text{global efficiency} \qquad \forall \; \text{bin}
+%\]
+One measure of non-uniformity is the standard deviation of bin efficiencies from the global efficiency:
+\[
+	\sqrt{\sum_{\bin} \left( \bineff - \globaleff \right)^2  }.
 \]
 
-So, basically, what we want to have in our dreams:
-\[
-	\bineff = \text{global efficiency} \qquad \forall \; \text{bin}
-\]
-
-To measure how far we are from ideal situation we use standard deviation:
-\[
-	\sqrt{\sum_{\bin} \left( \bineff - \globaleff \right)^2  }
-\]
-
-What is bad in this formula that every bin has some impact in the result, which does not depend on how many events are there, so metrics becomes very unstable to deviations in bins with only few events.
-To cure this, we add weights to the bins (note that $\sum_\bin \binweight = 1$):
+%What is bad in this formula that every bin has some impact in the result, which does not depend on how many events are there, so metrics becomes very unstable to deviations in bins with only few events.
+To make the metric more stable against fluctuations in bins which contain very few events, we add weights to the bins (note that $\sum_\bin \binweight = 1$):
 \[
@@
 \]
-so we have SDE formula:
+giving the weighted standard deviation (SDE) formula
 \[
@@
 \] 
-In fact, the expression depends on the cut, but for cuts which produce equal efficiency, this is
+%In fact, the expression depends on the cut, but for cuts which produce equal efficiency, this is
 
-Finally we note that the weighted average of $\bineff$ is $\globaleff$:
-\[
-	\globaleff = < \bineff > =  \sum_{\bin} \binweight \times \bineff,
-\]
-and this is why the introduced metrics was named SDE --- this is a weighted standard deviations of array of bin efficiencies.
+%Finally we note that the weighted average of $\bineff$ is $\globaleff$:
+%\[
+%	\globaleff = < \bineff > =  \sum_{\bin} \binweight \times \bineff,
+%\]
+%and this is why the introduced metrics was named SDE --- this is a weighted standard deviations of array of bin efficiencies.
 
-But this is how we measure the non-uniformity for only one fixed cut, to measure the overall non-flatness, we take several global efficiencies (for instance, [0.5, 0.6, 0.7, 0.8, 0.9], because in practice usually we are interested in cuts with high global efficiency) and use
+%But this is how we measure the non-uniformity for only one fixed cut, 
+This formula is valid for any given cut value. To measure the overall non-flatness of the selection, we
+take several global efficiencies and use
+%(for instance, [0.5, 0.6, 0.7, 0.8, 0.9], because in practice usually we are interested in cuts with high global efficiency) and use 
 \[
@@
 \]
-
-Some other power $p \neq 2$ can be used as well, but $p=2$ is considered as the default value: 
-\[
-	\SDE^p(\globaleff) = 
-	\sum_{\bin} \binweight \times \abs{\bineff - \globaleff}^p,
-\qquad
-	\SDE^p  =  \frac{1}{k} 
-	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
-		\text{SDE}^p(\globaleff).
-\]
+Another power $p \neq 2$ can be used as well, but $p=2$ is considered as the default value.
+%\[
+%	\SDE^p(\globaleff) = 
+%	\sum_{\bin} \binweight \times \abs{\bineff - \globaleff}^p,
+%\qquad
+%	\SDE^p  =  \frac{1}{k} 
+%	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
+%		\text{SDE}^p(\globaleff).
+%\]
+%
+%
 
@@
 
-One more measure uses Theil Index frequently used to measure economic inequality:
+The Theil Index is frequently used to measure economic inequality:
 \[
@@
 \[
-	\theil(\globaleff) = \sum_\bin \binweight \; \frac{\bineff}{\globaleff} \; \ln{\frac{\bineff}{\globaleff}}
+	\theil(\globaleff) = \sum_\bin \binweight \; \frac{\bineff}{\globaleff} \; \ln{\frac{\bineff}{\globaleff}}.
 \]
@@
 TODO how to combine Theil for different global efficiencies?
+VAVA : Suggest to combine a Theil of Theils, i.e. treat each global efficiency as a bin?
 \[
@@
 
-Let's start from reformulation of what is uniform predictions in signal.
+%Let's start from reformulation of what is uniform predictions in signal.
 First we split all signal events into some bins in uniform variables.
 There is some empirical distribution $F_\bin$ of predictions in each bin.
+Instead of measuring uniformity in terms of binned efficiencies, it is possible to consider the distribution of
+the binned classifier predictions, $F_\bin$, directly.
 Ideal uniformity means that all the distributions $F_\bin$ are equal and hence equal to the global distribution $F(x)$. 
-
 To 'measure' non-flatness we can use some distribution distance, like Kolmogorov-Smirnov:
@@
 \]
-
-The good point is we don't need to select some global efficiencies like in the other metrics.
-
-\subsection{Connection Between SDE and Distribution Similarity Approach}
-
-SDE and DSA based on Cram\'er--von Mises similarity can be shown to have connection.
-Let's consider the SDE with global efficiencies $= [1/N, 2/N, \dots, N/N]$. In the limit $N \to \infty$
-\[
-	\lim_{N \to \infty} \SDE^2 = 
-	\lim_{N \to \infty} \frac{1}{N} \sum_{\globaleff}\SDE^2(\globaleff) = 
-	\int_0^1 \SDE^2(\globaleff) d\, \globaleff = 
-	\int_0^1 \sum_{\bin} \binweight \abs{\bineff - \globaleff}^2 d\, \globaleff
-\]
-
-From the other side, we can write the expression for similarity-based measure (for $p=2$) 
-\[
-	\sum_{\bin} \binweight \int \abs{F_\text{bin}(x) - F(x)}^2 dF(x) =
-	\int \sum_{\bin} \binweight \abs{F_\text{bin}(x) - F(x)}^2 dF(x) 
-\] The hard thing now is to believe this is literally the same and these two expressions are equal.
+in particular because Kolmogorov-Smirnov measures are too sensitive to local non-uniformities.
+The advantage of this method is that we don't need to select some global efficiencies like in the previous metrics.
+%\subsection{Connection Between SDE and Distribution Similarity Approach}
+%Moreover, SDE and DSA based on Cram\'er--von Mises similarity can be shown to be connected.
+%Let's consider the SDE with global efficiencies $= [1/N, 2/N, \dots, N/N]$. In the limit $N \to \infty$
+%\[
+%	\lim_{N \to \infty} \SDE^2 = 
+%	\lim_{N \to \infty} \frac{1}{N} \sum_{\globaleff}\SDE^2(\globaleff) = 
+%	\int_0^1 \SDE^2(\globaleff) d\, \globaleff = 
+%	\int_0^1 \sum_{\bin} \binweight \abs{\bineff - \globaleff}^2 d\, \globaleff
+%\]
+%
+%From the other side, we can write the expression for similarity-based measure (for $p=2$) 
+%\[
+%	\sum_{\bin} \binweight \int \abs{F_\text{bin}(x) - F(x)}^2 dF(x) =
+%	\int \sum_{\bin} \binweight \abs{F_\text{bin}(x) - F(x)}^2 dF(x) 
+%\] The hard thing now is to believe this is literally the same and these two expressions are equal.
+%
 
@@
 Though operating with bins is usually both simple and very efficient, 
-in many cases it is hard to find optimal size of bins in the space of uniform variables (specifically in the case of more than two dimensions).
-One more situation when bins-based approach fails, is when we have too few events to obtain a good statistics at least in several bins.
-
-In these cases we can switch to $k$-nearest neighbors: for each signal event we find $k$ nearest signal events (including the event itself) in the space of uniform variables.
-Now we can compute the efficiency $\effknni$, empirical distribution $\Fknn$ of nearest neighbors.
+in many cases it is hard to find the optimal size of bins in the space of uniform features (specifically in the case of more than two dimensions).
+As mentioned earlier, problems can also arise due to bins with very low populations.
+%One more situation when bins-based approach fails, is when we have too few events to obtain a good statistics at least in several bins.
+In these cases we can switch to $k$-nearest neighbors: for each signal event we find $k$ nearest signal events (including the event itself)
+in the space of uniform features. Now we can compute the efficiency $\effknni$, from the empirical distribution $\Fknn$ of nearest neighbors. 
 The weights for $\knni$ are proportional to the total weight of events in $\knni$:
@@
 
-Now we are ready to write knn version of SDE:
+It is then possible to write the knn versions of SDE
 \[
 	\knnSDE^2(\globaleff)
-		= \sum_{i \in \text{events}} \weightknni \abs{\effknni - \globaleff}^2
+		= \sum_{i \in \text{events}} \weightknni \abs{\effknni - \globaleff}^2,
 \]
@@
 \]
-knn version of Theil index of Efficiency
+the Theil index of Efficiency
 \[
-	\text{knnTheil}(\globaleff) = \sum_{i \in \text{events}} \weightknni \; \frac{\effknni}{\globaleff} \; \ln{\frac{\effknni}{\globaleff}}
+	\text{knnTheil}(\globaleff) = \sum_{i \in \text{events}} \weightknni \; \frac{\effknni}{\globaleff} \; \ln{\frac{\effknni}{\globaleff}},
 \]
@@
 \]
-and knn version of similarity-based measure:
+and the similarity-based measure:
 
 \[
-	 \sum_{i \in \text{events}} \weightknni \int \abs{\Fknn(x) - F(x)}^p dF(x),
+	 \sum_{i \in \text{events}} \weightknni \int \abs{\Fknn(x) - F(x)}^p dF(x).
 \]
 
-$K$-nearest neighbors approach suffers from the other drawback: the impact of different events has very little connection with the weights, because some events are met in knn of other events much more frequently while the other.
-This effect can be suppressed by dividing initial weight of the event by the number of times it is met in knn.
+The knn approach suffers from a drawback: the impact of different events has very little connection with the weights,
+because some events are selected as nearest neighbours much more frequently than others.
+This effect can be suppressed by dividing the initial weight of the event by the number of times it is selected 
+as a nearest neighbour.
 
 \subsection{Advantages and Disadvantages of Different Metrics}
-Let us compare two metrics that have shown most appropriate for our problem, SDE and Theil. We have some linearly spaced array of mass of particles
+Let us compare two metrics that have proven most appropriate for our problem, SDE and Theil. We have some linearly spaced array of mass of particles
 from interval $[0,1]$, some constant $\alpha$ linearly sapced in interval $[.1,1]$. The predictions are correlated with mass via beta distribution.

### 2014-09-22 assets/paperdraft/parts/approaches.tex
Major textual changes to the section on uniformity metrics.
nothing

@@
 
-% TODO sign
-
 \begin{multline*}
@@
 
-TODO give explanations here
+Explanation:
+\begin{multline*}
+	- \dfrac{\partial} {\partial \, \score_i} \FL = 
+		- \binIweight \frac{\partial }{ \partial \, \score_i} \int \abs{F_{\bin(i)}(x) - F(x)}^p dx = \\
+	= \binIweight \abs{F_{\bin(i)}(x) - F(x)}^p \bigg|_{x=\score_i-}^{x=\score_i+} = \\
+	= \{ \text{assuming that $F_{\bin(i)}(x)$, $F(x)$ changed small } \} = \\
+	= \binIweight p \abs{F_{\bin(i)}(x) - F(x)}^{p-1} \sgn(F_{\bin(i)}(x) - F(x)) \left[F(x-) - F(x+) + F_{\bin(i)}(x+) - F_{\bin(i)}(x-)\right] \bigg|_{x=\score_i}= \\
+	= \binIweight p \abs{F_{\bin(i)}(x) - F(x)}^{p-1} \sgn(F_{\bin(i)}(x) - F(x)) \left[- \frac{w_i}{\sum_j w_j} + \frac{w_i}{\binIweight} \right] \bigg|_{x=\score_i}= \\
+	= \{ \text{ignoring $\frac{w_i}{\sum_j w_j}$}  \}
+	\cong \binIweight \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1} 
+	% 	\sgn[F_{\bin(i)}(x) - F(x)]
+	% 	\dfrac{w_i}{\binIweight}
+	% 	\Bigg|_{x=\score_i} = \\
+	% = 
+	% 	w_i \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1}
+	% 	\sgn[F_{\bin(i)}(x) - F(x)]
+	% 	\Bigg|_{x=\score_i}
+\end{multline*}
 

### 2014-09-22 IPythonWorkflow/higgs-submission.py

@@ -1,0 +1,42 @@
+from __future__ import print_function
+__author__ = 'axelr'
+
+import numpy
+import gradient_boosting as gb
+import pandas
+import transformations
+import cPickle as pickle
+
+trainX, trainY, trainW = gb.get_higgs_data()
+testX = pandas.read_csv('/Users/axelr/ipython/datasets/higgs/test.csv', index_col='EventId')
+
+
+clf = gb.ReweightingGB(loss=gb.AdaLossFunction()).set_params(
+    loss__signal_curvature=0.7, learning_rate=0.004, min_samples_leaf=125,
+    n_estimators=5000, max_features=14,
+    update_tree=True, max_depth=12, subsample=0.15, sig_weight=0.1,
+    weights_in_loss=False, update_on='all', recount_step=10000, smearing=0.)
+
+clf.fit(transformations.extend_data(trainX), trainY, trainW)
+p_old = clf.predict_proba(transformations.extend_data(testX))[:, 1]
+with open('submission-classifier2.pkl', 'wb') as output:
+    pickle.dump(clf, output)
+
+with open('submission-classifier2.pkl', 'rb') as input:
+    clf = pickle.load(input)
+
+p = clf.predict_proba(transformations.extend_data(testX))[:, 1]
+
+assert numpy.all(p == p_old)
+
+threshold = 0.5
+
+labels = numpy.array(['b'] * len(p))
+labels[p >= threshold] = 's'
+result_csv = pandas.DataFrame()
+result_csv['EventId'] = numpy.array(testX.index)
+result_csv['RankOrder'] = numpy.argsort(numpy.argsort(p)) + 1
+result_csv['Class'] = labels
+result_csv.to_csv('submission_2.csv', sep=',',  index=False)
+
+print('signal_events: ' + str(numpy.sum(p > threshold)) + ' from ' + str(len(p)))

### 2014-09-22 assets/paperdraft/ugb.tex

@@
 \begin{equation}
-  -\frac{\partial L_{\rm knn}}{\partial s_k} = -\sum\limits_i \gamma_i a_{ik}^{\rm knn} {\rm exp}\left[-\gamma_i \sum_j a_{ij}^{\rm knn} s_j \right].
+  -\frac{\partial L_{\rm knn}}{\partial s_k} = \sum\limits_i \gamma_i a_{ik}^{\rm knn} {\rm exp}\left[-\gamma_i \sum_j a_{ij}^{\rm knn} s_j \right].
 \end{equation}
@@
 The loss function given in Eq.~\ref{eq:L_ada_flat} can also be constructed without binning the data using k-nearest-neighbor events.
-The cummulative distribution $F_{\rm knn}(s)$ is easily obtained and the bin weight, $w_b$, is replaced by ???
+The cummulative distribution $F_{\rm knn}(s)$ is easily obtained and the bin weight, $w_b$, is replaced by a k-nearest-neighbor weight, $w_{\rm knn}$.
+First, each event is weighted by the inverse of the number of times it is included in the k-nearest-neighbor sample of another event.
+Then, $w_{\rm knn}$ is the sum of such weights in a k-nearest-neighbor sample divided by the total sum of such weights in the full sample.
+This procedure is followed to offset the fact that some events are found in more k-nearest-neighbor samples than other events.

### 2014-09-23 assets/paperdraft/parts/metrics.tex
I've added Dto3h metrics result, and merged theil and SDE graphs for my distributions.

@@
 \subsection{Advantages and Disadvantages of Different Metrics}
+\subsubsection{Theil and SDE}
 Let us compare two metrics that have proven most appropriate for our problem, SDE and Theil. We have some linearly spaced array of mass of particles
@@
 \centering
-	\includegraphics[width=8cm]{graphs/measures1.png}
-	\caption{SDE and Theil}
-\end{figure}
-
-\begin{figure}[H]
-\centering
 		\begin{subfigure}[b]{0.95\textwidth}
@@
 \centering
-	\includegraphics[width=8cm]{graphs/measures2.png}
+	\includegraphics[width=0.8\textwidth]{graphs/measures_compared.png}
 	\caption{SDE and Theil}
@@
 Lower value of SDE can indicate that distribution is flat or has some narrow peaks or pits, and by using Theil we can distinguish between these two cases.
+
+\subsubsection{$D \to hhh$}
+When our classifiers where applied to data set $D \to hhh$ we got uniformity measures in a graph below:
+\begin{figure}[H]
+\centering
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/dto3h_measures_sig.png}
+			\caption{Metrics for $D \to hhh$ background}
+		\end{subfigure}
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/dto3h_measures.png}
+			\caption{Metrics for $D \to hhh$ signal}
+		\end{subfigure}
+		\caption{Metrics for $D \to hhh$ data set}
+\end{figure}
+
+From these graphs we can see the all metrics have shown similar results, and that there is no significant difference on real data sets.
+SDE and Theil can be used on limit cases to more precisely determine nature of a distribution.

### 2014-09-23 IPythonWorkflow/binninguniforming.py
deleted binning uniforming

@@ -1,212 +1,0 @@
-from __future__ import division
-from __future__ import print_function
-
-import numpy
-from commonutils import check_sample_weight, weighted_percentile, \
-    build_normalizer, generate_sample
-from numpy.random.mtrand import RandomState
-from sklearn.base import BaseEstimator, ClassifierMixin, clone
-from sklearn.cross_validation import train_test_split
-
-
-__author__ = 'Alex Rogozhnikov'
-# About
-
-# How does it work: it splits the mass space into bins,
-# for each bin a separate basic classifier is learnt,
-# when getting prediction_proba, the predictions of basic classifier
-# are altered by some correcting monotonic function
-
-# WARNING
-# this module is concerned deprecated
-
-# TODO try doing the same without binning
-
-
-class Binner:
-    def __init__(self, values, n_bins, sample_weight=None):
-        """Binner is a class that helps to split the values into several bins.
-        Initially an array of values is given, which is then splitted into 'bins_number' equal parts,
-        and thus we are computing limits (boundaries of bins)."""
-        percentiles = [i / n_bins for i in range(1, n_bins)]
-        sample_weight = check_sample_weight(values, sample_weight=sample_weight)
-        self.limits = weighted_percentile(values, percentiles, sample_weight=sample_weight)
-
-    def get_bins(self, values):
-        return numpy.searchsorted(self.limits, values)
-
-    def set_limits(self, limits):
-        self.limits = limits
-
-    @property
-    def n_bins(self):
-        return len(self.limits) + 1
-
-    def split_into_bins(self, *arrays):
-        """Splits the data of parallel arrays into bins, the first array is binning variable"""
-        values = arrays[0]
-        numpy_arrays = []
-        for i, array in enumerate(arrays):
-            assert len(array) == len(values), "passed arrays have different lengths"
-            numpy_arrays.append(numpy.array(array))
-        bins = self.get_bins(values)
-        result = []
-        for bin in range(len(self.limits) + 1):
-            indices = bins == bin
-            result.append([array[indices] for array in numpy_arrays])
-        return result
-
-
-def test_binner():
-    """This function tests binner class"""
-    random = RandomState()
-    binner = Binner(random.permutation(30), 3)
[…]

### 2014-09-23 IPythonWorkflow/higgs-submission.py
removed trash file

@@ -1,42 +1,0 @@
-from __future__ import print_function
-__author__ = 'axelr'
-
-import numpy
-import gradient_boosting as gb
-import pandas
-import transformations
-import cPickle as pickle
-
-trainX, trainY, trainW = gb.get_higgs_data()
-testX = pandas.read_csv('/Users/axelr/ipython/datasets/higgs/test.csv', index_col='EventId')
-
-
-clf = gb.ReweightingGB(loss=gb.AdaLossFunction()).set_params(
-    loss__signal_curvature=0.7, learning_rate=0.004, min_samples_leaf=125,
-    n_estimators=5000, max_features=14,
-    update_tree=True, max_depth=12, subsample=0.15, sig_weight=0.1,
-    weights_in_loss=False, update_on='all', recount_step=10000, smearing=0.)
-
-clf.fit(transformations.extend_data(trainX), trainY, trainW)
-p_old = clf.predict_proba(transformations.extend_data(testX))[:, 1]
-with open('submission-classifier2.pkl', 'wb') as output:
-    pickle.dump(clf, output)
-
-with open('submission-classifier2.pkl', 'rb') as input:
-    clf = pickle.load(input)
-
-p = clf.predict_proba(transformations.extend_data(testX))[:, 1]
-
-assert numpy.all(p == p_old)
-
-threshold = 0.5
-
-labels = numpy.array(['b'] * len(p))
-labels[p >= threshold] = 's'
-result_csv = pandas.DataFrame()
-result_csv['EventId'] = numpy.array(testX.index)
-result_csv['RankOrder'] = numpy.argsort(numpy.argsort(p)) + 1
-result_csv['Class'] = labels
-result_csv.to_csv('submission_2.csv', sep=',',  index=False)
-
-print('signal_events: ' + str(numpy.sum(p > threshold)) + ' from ' + str(len(p)))

### 2014-09-23 IPythonWorkflow/meanadaboost.py
added max-boost for triggers

@@ -14,6 +14,7 @@ from commonutils import check_sample_weight, computeKnnIndicesOfSameClass
 
 # In this module some very simple modifications of AdaBoost
 # is presented. In weights boosting it uses mean of neighbours scores,
+import commonutils
 
 
 __author__ = 'Alex Rogozhnikov'
@@ -81,7 +82,7 @@ class MeanAdaBoostClassifier(BaseEstimator, ClassifierMixin):
             elif self.voting == 'matrix':
                 voted_score = A.dot(cumulative_score)
             else: # self.voting is callable
-                voted_score = self.voting(knn_scores)
+                voted_score = self.voting(cumulative_score, knn_scores)
 
             weight = sample_weight * numpy.exp(- y_signed * voted_score)
             weight = self.normalize_weights(y=y, sample_weight=weight)
@@ -136,5 +137,20 @@ class MeanAdaBoostClassifier(BaseEstimator, ClassifierMixin):
     def staged_predict_proba(self, X):
         for score in self.staged_predict_score(X):
             yield self.score_to_proba(score=score)
+
+
+def generate_max_voter(event_indices):
+    """
+    This voter is prepared specially for using in triggers.
+    Voting returns max(svr predictions),
+    :param event_indices: array, each element is
+
+    """
+    groups = commonutils.indices_of_values(event_indices)
+    def voter(cumulative_score, knn_scores):
+        result = numpy.zeros(len(cumulative_score))
+        for group in groups:
+            result[group] = numpy.max(cumulative_score[group])
+    return voter
 
 

### 2014-09-23 IPythonWorkflow/meanadaboost.py
hotfix

@@ -149,7 +149,7 @@ def generate_max_voter(event_indices):
     groups = commonutils.indices_of_values(event_indices)
     def voter(cumulative_score, knn_scores):
         result = numpy.zeros(len(cumulative_score))
-        for group in groups:
+        for key, group in groups:
             result[group] = numpy.max(cumulative_score[group])
     return voter
 

### 2014-09-23 IPythonWorkflow/meanadaboost.py
hotfix

@@ -151,6 +151,7 @@ def generate_max_voter(event_indices):
         result = numpy.zeros(len(cumulative_score))
         for key, group in groups:
             result[group] = numpy.max(cumulative_score[group])
+        return result
     return voter
 
 

### 2014-09-26 IPythonWorkflow/reports.py
added return data to  sde etc

@@ -189,6 +189,7 @@ class Predictions(object):
         for name, values in result.iteritems():
             pylab.plot(values.keys(), values, label=name)
         pylab.xlabel('stage')
+        return result
 
     #endregion
 
@@ -272,7 +273,8 @@ class Predictions(object):
             assert len(bin_centers[-1]) == n_bins
         return bin_centers
 
-    def sde_curves(self, uniform_variables, target_efficiencies=None, n_bins=20, step=3, power=2., label=1):
+    def sde_curves(self, uniform_variables, target_efficiencies=None, n_bins=20, step=3, power=2., label=1,
+                   return_data=False):
         mask = self.y == label
         bin_indices = self._compute_bin_indices(uniform_variables, n_bins=n_bins, mask=mask)
         target_efficiencies = self._check_efficiencies(target_efficiencies)
@@ -281,14 +283,15 @@ class Predictions(object):
             return compute_sde_on_bins(pred[:, label], mask=mask, bin_indices=bin_indices,
                                        target_efficiencies=target_efficiencies, power=power,
                                        sample_weight=self.checked_sample_weight)
-        result = pandas.DataFrame(self._map_on_staged_proba(compute_sde, step=step))
+        result = self._plot_curves(compute_sde, step=step)
         pylab.xlabel("stage"), pylab.ylabel("SDE")
-        for name, sde_values in result.iteritems():
-            pylab.plot(numpy.array(sde_values.index), numpy.array(sde_values), label=name)
         pylab.ylim(0, pylab.ylim()[1] * 1.15)
         pylab.legend(loc='upper center', bbox_to_anchor=(0.5, 1.00), ncol=3, fancybox=True, shadow=True)
+        if return_data:
+            return result
 
-    def sde_knn_curves(self, uniform_variables, target_efficiencies=None, knn=30, step=3, power=2, label=1):
+    def sde_knn_curves(self, uniform_variables, target_efficiencies=None, knn=30, step=3, power=2, label=1,
+                       return_data=True):
         """Warning: this functions is very slow, specially on large datasets"""
         mask = self.y == label
         knn_indices = computeSignalKnnIndices(uniform_variables, self.X, is_signal=mask, n_neighbors=knn)
@@ -299,14 +302,15 @@ class Predictions(object):
             return compute_sde_on_groups(pred[:, label], mask, groups_indices=knn_indices,
[…]

### 2014-09-29 assets/paperdraft/parts/metrics.tex
Make uniformity plots in paper style

--- 2014-09-21 assets/paperdraft/parts/metrics.tex
@@
 \subsection{Advantages and Disadvantages of Different Metrics}
+\subsubsection{Theil and SDE}
 Let us compare two metrics that have proven most appropriate for our problem, SDE and Theil. We have some linearly spaced array of mass of particles
-from interval $[0,1]$, some constant $\alpha$ linearly sapced in interval $[.1,1]$. The predictions are correlated with mass via beta distribution.
+from interval $[0,1]$, some constant $\alpha$ linearly spaced in interval $[.1,1]$. The predictions are correlated with mass via beta distribution.
 We have two distributions that are different in a way that they are symmetrical.
@@
 
-while the second one has minus sign in front.
+while the second one has minus sign in front. Figures~\ref{fig:peakdists}~and~\ref{fig:pitdists} show the distribution of the predictions
+and classifier efficiency as a function of mass for the peak and pit distributions, respectively. 
 \begin{figure}[H]
+\label{fig:peakdists}
 \centering
 		\begin{subfigure}[b]{0.95\textwidth}
-			\includegraphics[width=\textwidth]{graphs/distribution1.png}
-			\caption{Mass vs prediction}
+			\includegraphics[width=\textwidth]{graphs/PeakDistributions.png}
+			\caption{Mass vs prediction.}
 		\end{subfigure}
 		\begin{subfigure}[b]{0.95\textwidth}
-			\includegraphics[width=\textwidth]{graphs/efficiencies1.png}
-			\caption{Efficiencies}
+			\includegraphics[width=\textwidth]{graphs/PeakEffs.png}
+			\caption{Mass vs efficiency. The different coloured lines correspond to
+different global classifier efficiencies, as explained in the legend on the leftmost subplot.}
 		\end{subfigure}
-		\caption{Peak distribution}
+		\caption{Peak response distribution and efficiencies as a function of mass. }
+\end{figure}
+
+\begin{figure}[H]
+\label{fig:pitdists}
+\centering
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/PitDistributions.png}
+			\caption{Mass vs prediction.}
+		\end{subfigure}
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/PitEffs.png}
+			\caption{Mass vs efficiency. The different coloured lines correspond to
+different global classifier efficiencies, as explained in the legend on the leftmost subplot.}
+		\end{subfigure}
+		\caption{Pit response distribution and efficiencies as a function of mass.}
 \end{figure}
@@
 \centering
-	\includegraphics[width=8cm]{graphs/measures1.png}
-	\caption{SDE and Theil}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-		\begin{subfigure}[b]{0.95\textwidth}
-			\includegraphics[width=\textwidth]{graphs/distribution2.png}
-			\caption{Mass vs prediction}
-		\end{subfigure}
-		\begin{subfigure}[b]{0.95\textwidth}
-			\includegraphics[width=\textwidth]{graphs/efficiencies2.png}
-			\caption{Efficiencies}
-		\end{subfigure}
-		\caption{Pit distribution}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-	\includegraphics[width=8cm]{graphs/measures2.png}
-	\caption{SDE and Theil}
+	\includegraphics[width=0.5\textwidth]{graphs/TheilVsMSE.png}
+	\caption{Comparison of SDE and Theil uniformity measures for different values of the constant $\alpha$ for the peak and pit distributions.
+The Theil can clearly distinguish between these two while the SDE measure cannot.}
 \end{figure}
@@
 Lower value of SDE can indicate that distribution is flat or has some narrow peaks or pits, and by using Theil we can distinguish between these two cases.
+
+\subsubsection{$D \to hhh$}
+When our classifiers where applied to data set $D \to hhh$ we got uniformity measures for the SDE and Theil
+metrics shown in Fig.~\ref{fig:d2hhhsdetheil}.
+\begin{figure}[H]
+\label{fig:d2hhhsdetheil}
+\centering
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/D23hSignalEffs.png}
+			\caption{Metrics for $D \to hhh$ signal.}
+		\end{subfigure}
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/D23hBgEffs.png}
+			\caption{Metrics for $D \to hhh$ background.}
+		\end{subfigure}
+		\caption{Uniformity metrics for the $D \to hhh$ data set.
+The (top) signal and (bottom) background uniformities are plotted as a function of the training stage of a given classifier, listed in the legends on the leftmost plots.}
+\end{figure}
+
+Both metrics shown similar results, and that there is no significant difference on real data sets.
+The SDE and Theil metrics can be used together in limit cases to more precisely determine nature of a distribution.

### 2014-09-29 assets/paperdraft/parts/metrics.tex
updated plots

@@
 There is some empirical distribution $F_\bin$ of predictions in each bin.
+\begin{figure}[H]
+\label{fig:dsavisualization}
+\centering
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/bins_uniform_distribution.pdf}
+			\caption{Predictions are uniform in mass, the distribution in the bin (yellow) is close to the global(blue).}
+		\end{subfigure}
+		\begin{subfigure}[b]{0.95\textwidth}
+			\includegraphics[width=\textwidth]{graphs/bins_nonuniform_distribution.pdf}
+			\caption{Distribution with peak in the middle, he distributions are quite different now.}
+		\end{subfigure}
+		% \caption{Peak response distribution and efficiencies as a function of mass. }
+\end{figure}
 Instead of measuring uniformity in terms of binned efficiencies, it is possible to consider the distribution of
@@
 Ideal uniformity means that all the distributions $F_\bin$ are equal and hence equal to the global distribution $F(x)$. 
+This is demonstrated on figure \ref{
 To 'measure' non-flatness we can use some distribution distance, like Kolmogorov-Smirnov:

### 2014-09-30 assets/paperdraft/parts/timings.tex
fixes in several places

The main drawback of uBoost technique is it's high computational complexity: 
while simple AdaBoost trains $M$ trees, uBoost builds $100 \times M$ trees (contribution of other operations usually can be neglected).

Presented in this paper classifiers are building $M$ trees, though there is more complicated boosting, it takes at each iteration ($k$ is number of neighbors, $N$ is number of events in training sample)

\begin{itemize}
	\item meanAdaBoost: $O(k \times N)$
	\item knnAdaLoss: $O(k \times N)$, for the arbitrary matrix $A$ it is 
	$O( \text{\#nonzero elements in the matrix})$
	\item FlatnessLoss on bins: $O(N \ln N)$
	\item FlatnessLoss on knn: $O(N \ln N + N k \ln k) $ 
\end{itemize}

### 2014-10-01 assets/paperdraft/graphs/ploteffsinpaperstyle.py
add files for plots

@@ -1,0 +1,86 @@
+import os,sys
+import ROOT
+from ROOT import *
+from math import *
+from array import *
+
+basedir = './data/raspodela2/efikasnosti/'
+marker = 20
+size = 0.5
+colour = 1
+
+filestoplot = { '#alpha=0.1':{'dirname':'0','subfiles' : {}},
+                '#alpha=0.28':{'dirname':'1','subfiles' : {}},
+                '#alpha=0.47':{'dirname':'2','subfiles' : {}},
+                '#alpha=0.65':{'dirname':'3','subfiles' : {}},
+                '#alpha=0.84':{'dirname':'4','subfiles' : {}}}
+
+suffix = 'eff'
+
+# Need a double loop here
+for entry in filestoplot :
+  i = 0 
+  for subfile in ['0','2','4','6','8'] :
+    if i == 4 : i = 5
+    thisfile = open(basedir+filestoplot[entry]['dirname']+'/'+filestoplot[entry]['dirname']+"_"+subfile+suffix,'r')
+
+    filestoplot[entry]['subfiles'][subfile] = {}
+    filestoplot[entry]['subfiles'][subfile]['x'] = array('f',[0])
+    filestoplot[entry]['subfiles'][subfile]['y'] = array('f',[0])
+    num = 0
+    for line in thisfile :
+      filestoplot[entry]['subfiles'][subfile]['x'].append(float(line.split()[0]))
+      filestoplot[entry]['subfiles'][subfile]['y'].append(float(line.split()[1].strip('\n')))
+      num += 1
+
+    filestoplot[entry]['subfiles'][subfile]['graph'] = TGraph(num, filestoplot[entry]['subfiles'][subfile]['x'],
+                                                                   filestoplot[entry]['subfiles'][subfile]['y'])
+    filestoplot[entry]['subfiles'][subfile]['graph'].SetLineColor(colour+i)
+    filestoplot[entry]['subfiles'][subfile]['graph'].SetMarkerColor(colour+i)
+    filestoplot[entry]['subfiles'][subfile]['graph'].SetMarkerSize(size)
+    filestoplot[entry]['subfiles'][subfile]['graph'].SetMarkerStyle(marker+i)
[…]

### 2014-10-01 assets/paperdraft/parts/metrics.tex

--- 2014-09-29 assets/paperdraft/parts/metrics.tex
@@
 \]
-
-TODO how to combine Theil for different global efficiencies?
-VAVA : Suggest to combine a Theil of Theils, i.e. treat each global efficiency as a bin?
+To measure the overall non-flatness, we
+average values for several global efficiencies:
 \[
-	\theil = ??? \text{from} \theil(\globaleff)
+	\theil  =  \frac{1}{k} 
+	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
+		\theil(\globaleff)
 \]
@@
 There is some empirical distribution $F_\bin$ of predictions in each bin.
+\begin{figure}
+		\centering
+		\begin{subfigure}[b]{0.8\textwidth}
+			\includegraphics[width=\textwidth]{graphs/bins_uniform_distribution.pdf}
+			\caption{Predictions are uniform in mass, the distribution of predictions in the bin (yellow) is close to the global (blue).
+			Yellow rectangle shows the events in the bin over mass.}
+		\end{subfigure}
+		\begin{subfigure}[b]{0.8\textwidth}
+			\includegraphics[width=\textwidth]{graphs/bins_nonuniform_distribution.pdf}
+			\caption{Distribution with peak in the middle, the distribution in the bin is quite different from global.}
+		\end{subfigure}
+		\caption{Demonstration of distribution similarity approach. \label{fig:dsavisualization}}
+\end{figure}
 Instead of measuring uniformity in terms of binned efficiencies, it is possible to consider the distribution of
@@
 Ideal uniformity means that all the distributions $F_\bin$ are equal and hence equal to the global distribution $F(x)$. 
+This is demonstrated on figure \ref{fig:dsavisualization}.
 To 'measure' non-flatness we can use some distribution distance, like Kolmogorov-Smirnov:
@@
 \]
-so again weights are normed to 1: $\sum_{i} \knni = 1$.
+so again weights are normed to 1: $\sum_{i} \weightknni = 1$.
 
@@
 \[
-	\knnSDE^2 = \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]}
-		\knnSDE^2(\globaleff),
+	\knnSDE^2 = \frac{1}{k} \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]} \knnSDE^2(\globaleff),
 \]
@@
 \[
-	\text{knnTheil} = ??? \text{knnTheil}(\globaleff)
+	\text{knnTheil} = \frac{1}{k} \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]} \text{knnTheil}(\globaleff),
 \]
 and the similarity-based measure:
-
 \[
@@
 \begin{figure}[H]
-\label{fig:peakdists}
 \centering
@@
 		\end{subfigure}
-		\caption{Peak response distribution and efficiencies as a function of mass. }
+		\caption{Peak response distribution and efficiencies as a function of mass. \label{fig:peakdists}}
 \end{figure}
@@
 \begin{figure}[H]
-\label{fig:pitdists}
 \centering
@@
 		\end{subfigure}
-		\caption{Pit response distribution and efficiencies as a function of mass.}
+		\caption{Pit response distribution and efficiencies as a function of mass. \label{fig:pitdists}}
 \end{figure}
@@
 \begin{figure}[H]
-\label{fig:d2hhhsdetheil}
 \centering
@@
 		\caption{Uniformity metrics for the $D \to hhh$ data set.
-The (top) signal and (bottom) background uniformities are plotted as a function of the training stage of a given classifier, listed in the legends on the leftmost plots.}
+The (top) signal and (bottom) background uniformities are plotted as a function of the training stage of a given classifier, listed in the legends on the leftmost plots.
+\label{fig:d2hhhsdetheil}}
 \end{figure}

### 2014-10-01 assets/paperdraft/ugb.tex

@@
 \begin{equation}
-w_i^{\prime} = w_i \times {\rm exp}\left[-\gamma_i s_i\right], 
+w_i^{\prime} = w_i \times {\rm exp}\left[-\gamma_i p_i\right], 
 \end{equation}
-where $\gamma = +1(-1)$ for signal(background) events and $s$ is the so-called score of each event which is obtained as the sum of scores of all of the previous classifiers in the series.  
+where $\gamma = +1(-1)$ for signal(background) events and $p$ is the prediction for each event produced by last classifier in the series.  
 The uBoost technique, described in detail in Ref.~\cite{cit:uboost}, alters the event-weight updating procedure to achieve uniformity in the signal-selection efficiency.
@@
 \begin{equation}
- w_i^{\prime} = w_i \times {\rm exp}\left[-\gamma_i \sum_j a_{ij}\right],
+ w_i^{\prime} = w_i \times {\rm exp}\left[-\gamma_i \sum_j a_{ij} p_j\right],
 \end{equation}
@@
 \end{equation}
+So-called scores $s$ are obtained for each event as sum of predictions of all regressots in the series
 At each stage in the gradient boosting process, a {\em regressor} (a decision tree in our case) is trained whose purpose is to decrease the loss.
@@
 \end{equation}
-where $F_{(b)}(s)$ denotes the cummulative distribution of $f_{(b)}(s)$ and $w_b = \sum \delta({\rm bin}_i - b)/n_{\rm signal}$, {\em i.e.} $w_b$ is the fraction of signal events in the bin\footnote{If weighted events are used, then the fractional sum of weights should be used for $w_b$.}.
+where $F_{(b)}(s)$ denotes the cumulative distribution of $f_{(b)}(s)$ and $w_b = \sum \delta({\rm bin}_i - b)/n_{\rm signal}$, {\em i.e.} $w_b$ is the fraction of signal events in the bin\footnote{If weighted events are used, then the fractional sum of weights should be used for $w_b$.}.
 
-The gradient of the Kolmogorov-Smirnov loss function is zero for events with responses greater than the value of $s$ at which ${\rm max}|F_b(s)-F(s)|$ occurs.
+The gradient of the Kolmogorov-Smirnov loss function is zero for events with responses not equal to than the value of $s$ at which ${\rm max}|F_b(s)-F(s)|$ occurs.
 Thus, it is not suitable for gradient boosting due to its instability.
@@
 \end{equation}
-whose pseudo-residuals are
+whose pseudo-residuals are ($b$ is bin containing $k$th event)
 \begin{equation}
- -\frac{\partial L_{\rm flat}}{\partial s_k} \approx - 2\left[F_b(s_k)-F(s_k)\right]/n_{\rm signal}.
+ -\frac{\partial L_{\rm flat}}{\partial s_k} \approx - 2 w_b \left[F_b(s_k)-F(s_k)\right] .
 \end{equation}
@@
 The loss function given in Eq.~\ref{eq:L_ada_flat} can also be constructed without binning the data using k-nearest-neighbor events.  
-The cummulative distribution $F_{\rm knn}(s)$ is easily obtained and the bin weight, $w_b$, is replaced by a k-nearest-neighbor weight, $w_{\rm knn}$.
+The cumulative distribution $F_{\rm knn}(s)$ is easily obtained and the bin weight, $w_b$, is replaced by a k-nearest-neighbor weight, $w_{\rm knn}$.
 First, each event is weighted by the inverse of the number of times it is included in the k-nearest-neighbor sample of another event.

### 2014-10-01 assets/paperdraft/parts/metrics.tex
minor fixes

@@
 \item
-The metric should be stable against any of its own free paramters : if it uses bins, changing the number of bins shouldn't affect the result,
+The metric should be stable against any of its own free parameters: if it uses bins, changing the number of bins shouldn't affect the result,
 if it uses $k$-nearest neighbors, it should be stable against different values of $k$.
@@
 There is some empirical distribution $F_\bin$ of predictions in each bin.
-\begin{figure}
+\begin{figure}[h]
 		\centering
@@
 and classifier efficiency as a function of mass for the peak and pit distributions, respectively. 
-\begin{figure}[H]
+\begin{figure}[h]
 \centering
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
 \centering
@@
 metrics shown in Fig.~\ref{fig:d2hhhsdetheil}.
-\begin{figure}[H]
+\begin{figure}[h]
 \centering

### 2014-10-01 assets/paperdraft/parts/approaches.tex

@@
+\section{Boosting approaches}
 \subsection{Mean Ada Boost}
@@
 
-After a new tree was built, it's output is altered: for each leaf we can compute such a value, that the result will give the smallest possible value for loss.
-
-...computation of optimal value goes here...
-
 This loss function can be easily modified to take in account not only the score for individual elements, but also 'finds' regions with lower-than-average quality.
@@
 The named loss targets to obtain uniformity in both signal and background.
-
-And the post-correction can be handled by using second-order Newton-Raphson optimization step
-
-... TODO computation of optimal value goes here ...
 

### 2014-10-01 assets/paperdraft/parts/introduction.tex

@@
-Methods of machine learning are playing a important role in particles physics nowadays, in particular, multivariate analysis (MVA). 
+Methods of machine learning are playing an important role in particles physics nowadays, in particular, multivariate analysis (MVA). 
 Different classifiers like boosted decision trees (BDTs) and artificial neural networks (ANNs) are often used as an important step in analysis selection criteria.
@@
 The features, along which we want the prediction to be flat, will be referred to as 
-\textit{uniform variables}. In high energy physics these are usually masses.
+``uniform variables''. In high energy physics these are usually masses.

### 2014-10-02 assets/paperdraft/parts/metrics.tex
minor fixes in text

@@
 \def\knnSDE{\text{knnSDE}}
+\newcommand*\mean[1]{\overline{#1}}
 
@@
 As mentioned earlier, problems can also arise due to bins with very low populations.
-%One more situation when bins-based approach fails, is when we have too few events to obtain a good statistics at least in several bins.
+
 In these cases we can switch to $k$-nearest neighbors: for each signal event we find $k$ nearest signal events (including the event itself)
@@
 \]
-the Theil index of Efficiency
+the Theil index of efficiency
 \[
@@
 \subsubsection{Theil and SDE}
-Let us compare two metrics that have proven most appropriate for our problem, SDE and Theil. We have some linearly spaced array of mass of particles
-from interval $[0,1]$, some constant $\alpha$ linearly spaced in interval $[.1,1]$. The predictions are correlated with mass via beta distribution.
-We have two distributions that are different in a way that they are symmetrical.
-SDE should show no changes if we flip the distribution, while the Theil should make difference between pit and peak.
-First distribution with a peak is given by:
-
+Let us compare two metrics that have proven most appropriate for our problem, SDE and Theil.
+We have some masses distributed uniformly in $[0,1]$, some constant $\alpha$ from interval $[.1,1]$.
+The predictions are correlated with mass via beta distribution.
+We have two distributions that are different in a way that they are symmetrical. SDE should show no changes if we flip the distribution, while the Theil should make difference between pit and peak. 
+First distribution with a peak is obtained by generating a prediction for each event according to it mass:
 \[
-	p\text{[i]}=\textbf{beta}(\alpha \cdot \bar m, \text{abs}(m\text[i]-\bar m))
+	p \thicksim \text{Beta}(\alpha \mean{m}, \abs{m - \mean{m}})
 \]
-
 while the second one has minus sign in front. Figures~\ref{fig:peakdists}~and~\ref{fig:pitdists} show the distribution of the predictions
@@
 
-Both metrics shown similar results, and that there is no significant difference on real data sets.
+Both metrics shown similar results, and that there is no significant difference on real data sets (as well as CvM).
 The SDE and Theil metrics can be used together in limit cases to more precisely determine nature of a distribution.

### 2014-10-02 assets/paperdraft/paper_draft.tex
Major textual changes to the section on uniformity metrics.

@@
 Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~,\newline
-Andrey Ustyuzhanin$^{b,f}$ and
+Andrey Ustyuzhanin$^{b,f,g}$ and
 Mike Williams$^c$\\
@@
 \llap{$^f$}Moscow Institute of Physics and Technology, Moscow, Russia\\
+\llap{$^g$}Imperial College, London, UK\\
 E-mail: \email{alex.rogozhnikov@yandex.ru}}
@@
 \emph{PYTHIA 6.4 physics and manual},
-\href{http://dx.doi.org/10.1088/1126-6708/2006/05/026},
+\href{http://dx.doi.org/10.1088/1126-6708/2006/05/026}
 {\emph{JHEP} {\bf 05} (2006) 026}.
@@
 \emph{Handling of the generation of primary events in GAUSS, the LHCb simulation framework},
-\href{http://dx.doi.org/10.1109/NSSMIC.2010.5873949},
+\href{http://dx.doi.org/10.1109/NSSMIC.2010.5873949}
 {\emph{NSS/MIC} {\bf IEEE} (2010) 1155}.
@@
 \emph{The EvtGen particle decay simulation package},
-\href{http://dx.doi.org/10.1016/S0168-9002(01)00089-4},
+\href{http://dx.doi.org/10.1016/S0168-9002(01)00089-4}
 {\emph{NIM} {\bf A462} (2001) 152-155}.
@@
 \emph{PHOTOS Monte Carlo: a precision tool for QED corrections in $Z$ and $W$ decays},
-\href{http://dx.doi.org/10.1140/epjc/s2005-02396-4},
+\href{http://dx.doi.org/10.1140/epjc/s2005-02396-4}
 {\emph{EPJ} {\bf C45} (2006) 97-107}.
@@
 \emph{Geant4 developments and applications},
-\href{http://dx.doi.org/10.1109/TNS.2006.869826},
+\href{http://dx.doi.org/10.1109/TNS.2006.869826}
 {\emph{IEEE TNS} {\bf 53} (2006) 270}.
@@
 \emph{Geant4: a simulation toolkit},
-\href{http://dx.doi.org/10.1016/S0168-9002(03)01368-8},
+\href{http://dx.doi.org/10.1016/S0168-9002(03)01368-8}
 {\emph{NIM} {\bf A506} (2003) 250}.
@@
 \emph{The LHCb simulation application, GAUSS : design, evolution and experience},
-\href{http://dx.doi.org/10.1088/1742-6596/331/3/032023},
+\href{http://dx.doi.org/10.1088/1742-6596/331/3/032023}
 {\emph{J. Phys. Conf. Ser.} {\bf 331} (2011) 032023}.
+
+\bibitem{ref:FS1997}
+{Y. Freund, and R. Schapire},
+\emph{A Decision-Theoretic Generalization of On-Line Learning and an Application to Boosting}, 
+{\emph{Journal of Computer and System Sciences} {\bf 55(1)} (August 1997) 119-139}.
+
+\bibitem{ref:F1999}
+{Jerome H. Friedman},
+\emph{Greedy Function Approximation: A Gradient Boosting Machine},
+\href{http://statweb.stanford.edu/~jhf/ftp/trebst.pdf}
+{\emph{Annals of Statistics} {\bf 29} (2000) 1189-1232}.
 
@@
 \appendix
-\section{Other Flatness Loss Criteria}
-
-Clean all this up...
-
 %\section{Uniformity Measurement}

### 2014-10-02 assets/paperdraft/parts/introduction.tex
bibliography update
minor fixes in text

@@
 
-The one approach that was already proposed is uBoost (uniform BOOSTing) --- a modification of AdaBoost algorithm [link]
+The one approach that was already proposed is uBoost (uniform BOOSTing) --- a modification of AdaBoost algorithm~\cite{ref:FS1997}
 
@@
 The features, along which we want the prediction to be flat, will be referred to as 
-``uniform variables''. In high energy physics these are usually masses.
+\textit{uniform variables}. In high energy physics these are usually masses.

### 2014-10-02 assets/paperdraft/parts/dataset.tex
Mike edits

@@
-\section{Datasets}
+%\section{Datasets}
+
+The example analysis studied here involves a so-called Daltiz-plot analysis.
+In such analyses, the distribution of events in a 2-D space is typically fit to extract some information of physical interest.
+The regions of the Daltiz-plot that tend to have the highest sensitivity to the desired information are the {\em edges}.
+Unfortunately, the edge regions also typically have the most background contamination and the least discrimination against background.
+Therefore, traditional classifier-based selections tend to produce selections for Dalitz-plot analyses with lower efficiency near the edges.
+
 This study uses simulated event samples produced using the official LHCb simulation framework.
@@
 in the $c\bar{c}$ and $b\bar{b}$ simulated samples.
+
+\begin{figure}[] 
+  \centering 
+  \includegraphics[width=0.49\textwidth]{DP_sig.pdf}
+  \includegraphics[width=0.49\textwidth]{DP_bkgd.pdf}
+  \caption{\label{fig:dalitz} Dalitz-plot distributions for (left) signal and (right) background for the $D_s^\pm\to\pi^+\pi^-\pi^\pm$.  The three pions are labeled here as 1, 2 and 3.}
+\end{figure}
+
+Figure~\ref{fig:dalitz} shows the Dalitz-plot distributions for signal and background events.
+These samples are split into training and testing samples and then various BDTs are trained.
+Figure~\ref{fig:dalitz_results} shows the efficiency obtained for each classifier {\em vs} distance from the a corner of the Dalitz-plot.
+The AdaBoost algorithm, as expected, produces a much lower efficiency in the interesting corner regions.
+The kNNAdaBoost algorithm does not improve upon the AdaBoost result much.
+This is likely due to the fact that while kNNAdaBoost uses non-local kNN information, it does not utilize global information.
+The UGBFL (binned and unbinned kNN) and uBoost algorithms each produce an efficiency which is statistically consistent with uniform across the Dalitz plot.
+
+\begin{figure}[] 
+  \centering 
+  \includegraphics[width=0.49\textwidth]{ROC_DP.pdf}
+  \includegraphics[width=0.49\textwidth]{ROC_DP_Alpha.pdf}
+  \caption{\label{fig:dalitz_rocs} Efficiency {\em vs} distance to a corner of the Dalitz-plot.  An arbitrary working point of 50\% integrated efficiency is displayed.}
+\end{figure}
+
+\begin{figure}[] 
+  \centering 
+  \includegraphics[width=0.49\textwidth]{DP_compare.pdf}
+  \caption{\label{fig:dalitz_results} Efficiency {\em vs} distance to a corner of the Dalitz-plot.  An arbitrary working point of 50\% integrated efficiency is displayed.}
+\end{figure}
+
+ADD DISCUSSION ON ROC CURVE HERE ...
+
+\begin{figure}[] 
+  \centering 
+  \includegraphics[width=0.49\textwidth]{MD_eff.pdf}
+  \caption{\label{fig:dalitz_results} Background mis-identification {\em vs} $D$ candidate mass for (black circles) AdaBoost, (cyan up triangles) kNNAdaBoost, (red filled squares) UGBFL(bin), (red open squares) UGBFL(knn), and (blue down triangles) uBoost.
+An arbitrary working point of 10\% background mis-identification in the training region $1.75 < m(D) < 1.85$~GeV is displayed.}
+\end{figure}

### 2014-10-02 assets/paperdraft/parts/introduction.tex
Mike edits

--- 2014-10-01 assets/paperdraft/parts/introduction.tex
@@
-Methods of machine learning are playing an important role in particles physics nowadays, in particular, multivariate analysis (MVA). 
-Different classifiers like boosted decision trees (BDTs) and artificial neural networks (ANNs) are often used as an important step in analysis selection criteria.
+Methods of machine learning play an important role in modern particles physics. 
+Multivariate classifiers, {\em e.g.}, boosted decision trees (BDTs) and artificial neural networks (ANNs), are now commonly used in analysis selection criteria.
+BDTs are now even used in software triggers~\cite{ref:lhcbhlt,ref:bbdt}.
+To enhance the performance, a series of classifiers is typically trained; this is a technique known as boosting.
+Boosting  involves training many simple classifiers and then building a single composite classifier from their responses.
+The classifiers are trained in series with the inputs of each member being augmented based on the performance of its predecessors.
+This augmentation is designed such that each new classifier targets more those events which were poorly classified by previous members of the series.
+The classifier obtained by combining all members of the series is typically much more powerful than any of the individual members.
 
-BDTs are now even used in software triggers [link]. The main point of boosting technique is training many simple classifiers and building a composition of their outputs.
-Classifiers are trained one-after-another, their inputs are augmented in such a way that new classifier should target more at those events which were poorly classified by previous ones.
-The resulting classifier obtained by combining them often much more powerful, than individual ones.
-
-In practice, however, there are some restrictions that trained classifier should meet (apart from having good classification quality, which is usually measured by some integrated FOM).
-For example, in an amplitude analysis obtaining a uniform efficiency in a multivariate space of physics variates, i.e., variates that are of physical interest, is often times more important than any integrated FOM based on the total amount of signal and background.
-Such analyses often have many variates in which a uniform efficiency is desired.
-A uniform efficiency reduces systematic uncertainties and helps maintain sensitivity to all hypotheses being tested.
-
-The one approach that was already proposed is uBoost (uniform BOOSTing) --- a modification of AdaBoost algorithm [link]
-
-Why do we need such algorithms:
-
-\begin{enumerate}
-	\item The result of classification is stable to check different hypotheses
-	\item To get unbiased distribution of signal, this enables us to compute real mass of particle
-	\item Not to get fake peak, thus not get a false discovery
-\end{enumerate}
-
-The features, along which we want the prediction to be flat, will be referred to as 
-``uniform variables''. In high energy physics these are usually masses.
+In particle physics, the most common usage of BDTs is in classifying candidates as signal or background.
+The BDT is determined by optimizing some figure of merit (FOM), {\em e.g.}, the signal purity or approximate signal significance.
+This approach is optimal for a counting experiment; however, in many cases the BDT-based selection obtained in this way is not optimal.
+For example, in a Dalitz-plot (or any angular or amplitude analysis) analysis, obtaining a selection efficiency on signal candidates that is uniform across the Dalitz-plot is more important than any integrated FOM.
+Similarly, when measuring a mean particle lifetime, obtaining an efficiency that is uniform in lifetime is what is desired.
+In both cases, obtaining a uniform selection efficiency greatly reduces the systematic uncertainties involved in the measurement.
+When searching for a new particle, an analyst may want a uniform efficiency in mass for selecting background candidates so that the BDT-based selection does not generate a fake signal peak.
+Furthermore, the analyst may also desire a uniform selection efficiency of signal candidates in mass (or other variates) since the new particle mass is not known.
+In such cases, the BDT is often trained on simulated data generated with several values of mass (lifetime, {\em etc.}).
+A uniform selection efficiency in mass ensures that the BDT is sensitive to the full range of masses involved in the search.

### 2014-10-04 assets/paperdraft/parts/metrics.tex
deleted multiplication by 10 in SDE

@@
 \subsubsection{Theil and SDE}
-Let us compare two metrics that have proven most appropriate for our problem, SDE and Theil.
-We have some masses distributed uniformly in $[0,1]$, some constant $\alpha$ from interval $[.1,1]$.
+Let us compare two metrics that have proven most appropriate for our problem, SDE and Theil. We have some masses distributed uniformly in $[0,1]$, some constant $\alpha$ from interval $[0,1]$. 
 The predictions are correlated with mass via beta distribution.
-We have two distributions that are different in a way that they are symmetrical. SDE should show no changes if we flip the distribution, while the Theil should make difference between pit and peak. 
-First distribution with a peak is obtained by generating a prediction for each event according to it mass:
+First distribution with a peak is obtained by generating a prediction for each event according to its mass:
 \[
-	p \thicksim \text{Beta}(\alpha \mean{m}, \abs{m - \mean{m}})
+	p_{\rm peak} \thicksim \text{Beta}(\alpha \mean{m}, \abs{m - \mean{m}})
 \]
-while the second one has minus sign in front. Figures~\ref{fig:peakdists}~and~\ref{fig:pitdists} show the distribution of the predictions
+The second distribution is obtained by flipping:
+\[
+	p_{\rm pit} \thicksim 1 - \text{Beta}(\alpha \mean{m}, \abs{m - \mean{m}})
+\]
+SDE should show no changes if we flip the distribution, while the Theil should make difference between pit and peak.
+
+Figures~\ref{fig:peakdists} and~\ref{fig:pitdists} show the distribution of the predictions
 and classifier efficiency as a function of mass for the peak and pit distributions, respectively. 
@@
 \centering
+		% maybe we can just ignore this?
 		\begin{subfigure}[b]{0.95\textwidth}
@@
 
-From these figures we can see that SDE has the same value and it is not able to make difference between these distributions.
-In the second case Theil has lower value which indicates that distribution is flatter.
-In the combination with SDE it can be used to detarmine the type of a distribution.
-Lower value of SDE can indicate that distribution is flat or has some narrow peaks or pits, and by using Theil we can distinguish between these two cases.
+From these figures we can see that SDE doesn't make any difference between these distributions, while Theil has lower value in the second case 
+which indicates that distribution is flatter.
+This example demonstrates that Theil has larger penalty for distributions with narrow peaks rather than with narrow pits in the distribution of efficiecies.
 
@@
 \centering
+		% TODO use metrics on signal for signal, use metrics on bck for bck.
 		\begin{subfigure}[b]{0.95\textwidth}
@@
 Both metrics shown similar results, and that there is no significant difference on real data sets (as well as CvM).
-The SDE and Theil metrics can be used together in limit cases to more precisely determine nature of a distribution.

### 2014-10-04 assets/paperdraft/parts/dataset.tex

@@
 These samples are split into training and testing samples and then various BDTs are trained.
-Figure~\ref{fig:dalitz_results} shows the efficiency obtained for each classifier {\em vs} distance from the a corner of the Dalitz-plot.
-The AdaBoost algorithm, as expected, produces a much lower efficiency in the interesting corner regions.
-The kNNAdaBoost algorithm does not improve upon the AdaBoost result much.
-This is likely due to the fact that while kNNAdaBoost uses non-local kNN information, it does not utilize global information.
-The UGBFL (binned and unbinned kNN) and uBoost algorithms each produce an efficiency which is statistically consistent with uniform across the Dalitz plot.
+For the BDTs designed to produce uniform selections, the $\vec{y}$ variates are the Dalitz masses with the choice of uniform selection efficiency on signal candidates in the Dalitz-plot.
+Figure~\ref{fig:dalitz_rocs} shows the ROC curves obtained for the various classifiers studied in this paper.
+For the UGBFL algorithms, there is a choice to be made for the value $\alpha$ which defines the relative weight of the flatness loss {\em vs} AdaLoss.
+As expected, increasing $\alpha$, which increases the weight of AdaLoss, drives the ROC curve to be similar to AdaBoost.
+Analysts will need to choose how much ROC performance to sacrifice to gain uniformity in the selection efficiency.
+In general, the ROC curves for the uniform-driven BDTs are not too different from AdaBoost.
 
@@
   \includegraphics[width=0.49\textwidth]{ROC_DP_Alpha.pdf}
-  \caption{\label{fig:dalitz_rocs} Efficiency {\em vs} distance to a corner of the Dalitz-plot.  An arbitrary working point of 50\% integrated efficiency is displayed.}
+  \caption{\label{fig:dalitz_rocs} (left) ROC curves for classifier algorithms studied in this paper.  (right) ROC curves for UGBFL(bin) for differnet values of $\alpha$.}
 \end{figure}
+
+Figure~\ref{fig:dalitz_results} shows the efficiency obtained for each classifier {\em vs} distance from the a corner of the Dalitz-plot\footnote{The Dalitz-plot is essentially a triangle with three corners.
+Our defnition of this distance is ${\rm MIN}\left[(m(D_s)-m(\pi))^2 - m_{ij}^2\right]$, where $ij$ is 12, 13 and 23.}.
+The AdaBoost algorithm, as expected, produces a much lower efficiency in the interesting corner regions.
+The kNNAdaBoost algorithm does not improve upon the AdaBoost result much.
+This is likely due to the fact that while kNNAdaBoost uses non-local kNN information, it does not utilize global information.
+The UGBFL (binned and unbinned kNN) and uBoost algorithms each produce an efficiency which is statistically consistent with uniform across the Dalitz plot.  
+As stated above, the analyst is free to optimize the choice of $\alpha$ for UGBFL by defining a metric that involves signal efficiency, background rejection and uniformity, {\em e.g.}, using uniformity metrics discussed in detail in Appendix~A.
 
@@
 
-ADD DISCUSSION ON ROC CURVE HERE ...
+As a separate study using the same data samples, consider the case where one has simulated signal events and uses data from a {\em nearby} region, a so-called sideband, for background.
+This is a common situation in particle-physics analyses.
+Figure~\ref{fig:md_results} shows the training samples used.
+A major problem can arise in these situations as typically input variates to the BDT are correlated with the parent particle mass.
+Therefore, the BDT may learn to reject the background in the training using the fact that the mass of the background and signal candidates is different.
+This is just an artifact of how the background sample is obtained and will not be true for background candidates {\em under} the signal peak.
+Figure~\ref{fig:md_results} shows the background mis-indentifcation rate {\em vs} $D$ candidate mass.
+AdaBoost has clearly learned to use this mis-match in signal and background candidate masses in the training.
+The background in the region of the signal is about three times higher than one would expect from looking only at the sideband data.
 
@@
   \centering 
+  \includegraphics[width=0.49\textwidth]{mD_train.pdf}
   \includegraphics[width=0.49\textwidth]{MD_eff.pdf}
-  \caption{\label{fig:dalitz_results} Background mis-identification {\em vs} $D$ candidate mass for (black circles) AdaBoost, (cyan up triangles) kNNAdaBoost, (red filled squares) UGBFL(bin), (red open squares) UGBFL(knn), and (blue down triangles) uBoost.
-An arbitrary working point of 10\% background mis-identification in the training region $1.75 < m(D) < 1.85$~GeV is displayed.}
+  \caption{\label{fig:md_results} (left) Signal and (filled) background samples used in training.
+(right) Background mis-identification {\em vs} $D$ candidate mass for an arbitrary working point of 10\% background mis-identification in the training region $1.75 < m(D) < 1.85$~GeV is displayed.}
 \end{figure}
+
+Figure~\ref{fig:md_results} also shows the background mis-indentifcation rate {\em vs} $D$ candidate mass for the various uniform classifiers where $y = m(D)$ and the choice is for uniformity in the background efficiency\footnote{The algorithms in this paper can easily be made uniform on signal, background or both.}.
+The uBoost algorithm does better than AdaBoost here but is still not optimal.
+The way that uBoost achieves uniformity is not such that it can be trusted to work outside the region of training.
+The algorithms presented in this paper each does well in achieving similar performance in the training and signal regions.
+Consider, {\em e.g.}, the UGBFL approach to achieving uniform selection efficiency.
+In this case the training drives the BDT response itself to have the same PDF everywhere in the region $1.75 < m(D) < 1.85$~GeV (the training region).
+This does not gaurantee that the BDT response is truly independent of $m(D)$ outside the training region, but does strongly suppress learning to use $m(D)$ and in this example results in the desired behavior.
+Finally, if both high and low $m(D)$ sidebands had been used, it is possible for a BDT to create a fake peak near the signal peak location.
+The use of UGBFL greatly reduces the chances and possible size of such an effect.

### 2014-10-04 assets/paperdraft/parts/timings.tex

@@
-The main drawback of uBoost technique is it's high computational complexity: 
-while simple AdaBoost trains $M$ trees, uBoost builds $100 \times M$ trees (contribution of other operations usually can be neglected).
-
-Presented in this paper classifiers are building $M$ trees, though there is more complicated boosting, it takes at each iteration ($k$ is number of neighbors, $N$ is number of events in training sample)
-
+One drawback of the uBoost technique is that it has a high degree of computational complexity: 
+while AdaBoost trains $M$ trees (a user-defined number), uBoost builds $100 \times M$ trees.  
+The algorithms presented in this paper only build $M$ trees; however, the boosting involves some more complicated algorithms.  Training each of the $M$ trees scales as follows for $N$ training events:
 \begin{itemize}
-	\item meanAdaBoost: $O(k \times N)$
-	\item knnAdaLoss: $O(k \times N)$, for the arbitrary matrix $A$ it is 
-	$O( \text{\#nonzero elements in the matrix})$
-	\item FlatnessLoss on bins: $O(N \ln N)$
-	\item FlatnessLoss on knn: $O(N \ln N + N k \ln k) $ 
+       \item kNNAdaBoost: $O(k \times N)$;
+	\item UGBkNNknn: $O(k \times N)$ for $A_{\rm knn}$, and  
+	$O( \text{\#nonzero elements in the matrix})$ for arbitrary matrix $A$;
+	\item UGBFL(bin): $O(N \ln N)$;
+	\item UGBFL(kNN): $O(N \ln N + N k \ln k) $.
 \end{itemize}
+In the example analysis studied in this paper, we find that the training time for these new algorithms is within a factor of two the same as AdaBoost.
+The CPU-resource usage of these new algorithms is not prohibative.

### 2014-10-07 IPythonWorkflow/commonutils.py
fix of test for bad arithmetics

@@ -223,7 +223,7 @@ def test_build_normalizer(checks=10):
     weighted_normalizer = build_normalizer(predictions, sample_weight=predictions)
     result = weighted_normalizer(predictions)
     assert numpy.all(result[numpy.argsort(predictions)] == sorted(result))
-    assert numpy.all(result >= 0) and numpy.all(result <= 1)
+    assert numpy.all(result >= 0) and numpy.all(result <= 1 + 1e-7)
     predictions = numpy.sort(predictions)
     result = weighted_normalizer(predictions)
     result2 = numpy.cumsum(predictions) / numpy.sum(predictions)

### 2014-10-07 IPythonWorkflow/metrics.py
fix for test of roc curve

@@ -164,7 +164,7 @@ def test_roc_curve(size=100):
     fpr1, tpr1, thr1 = sklearn.metrics.roc_curve(y, pred)
     fpr2, tpr2, thr2 = roc_curve(y, pred)
     # this is insufficient test really
-    assert auc(fpr1, tpr1) == auc(fpr2, tpr2)
+    assert numpy.allclose(auc(fpr1, tpr1), auc(fpr2, tpr2))
 
 test_roc_curve(100)
 
@@ -515,6 +515,12 @@ def _ks_2samp_fast(prepared_data1, data2, prepared_weights1, weights2, F1):
     prepared_weights2 = numpy.bincount(indices, weights=weights2, minlength=len(prepared_data1))
     F2 = compute_cdf(prepared_weights2)
     return numpy.max(numpy.abs(F1 - F2))
+
+def ks_2samp_weighted(data1, data2, weights1, weights2):
+    x = numpy.unique(numpy.concatenate([data1, data2]))
+    inds1 = numpy.searchsorted(x, data1)
+
+
 
 
 def test_ks2samp_fast(size=1000):

### 2014-10-07 assets/paperdraft/paper_draft.tex
Mike edits

@@
 \author{
-Aleksandar Bukva$^e$, 
+Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~,
+Aleksandar Bukva$^c$, 
 Vladimir Gligorov$^d$,
-Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~,\newline
-Andrey Ustyuzhanin$^{b,f,g}$ and
-Mike Williams$^c$\\
+Andrey Ustyuzhanin$^{b,e,f}$ and
+Mike Williams$^g$\\
 \llap{$^a$}Lomonosov Moscow State University, Moscow, Russia\\
 \llap{$^b$}Yandex, Moscow, Russia\\
-\llap{$^c$}Massachusetts Institute of Technology, Cambridge, MA, United States \\
+\llap{$^c$}Faculty of Physics, Belgrade, Serbia \\
 \llap{$^d$}Organisation Europ\'eenne pour la Recherche Nucl\'eaire (CERN), Geneva, Switzerland  \\
-\llap{$^e$}Faculty of Physics, Belgrade, Serbia \\
-\llap{$^f$}Moscow Institute of Physics and Technology, Moscow, Russia\\
-\llap{$^g$}Imperial College, London, UK\\
+\llap{$^e$}Moscow Institute of Physics and Technology, Moscow, Russia\\
+\llap{$^f$}Imperial College, London, UK\\
+\llap{$^g$}Massachusetts Institute of Technology, Cambridge, MA, United States \\
 E-mail: \email{alex.rogozhnikov@yandex.ru}}
@@
 
-\section{Timings}
+\section{CPU Resources}
 
@@
 
-%\textcolor{red}{xxx}
-
-The link on the repository ans links to final notebooks with results in the article.
+UPDATE THIS WITH CODE LOCATION.
+%The link on the repository ans links to final notebooks with results in the article.
 
@@
 We particularly acknowledge the work of the LHCb Simulation and Core Computing teams in tuning the simulation software and managing the productions of simulated events.
+MW is supported by NSF grant PHY-1306550.
 
@@
 
+\bibitem{ref:FS1997}
+{Y. Freund, and R. Schapire},
+\emph{A Decision-Theoretic Generalization of On-Line Learning and an Application to Boosting}, 
+{\emph{Journal of Computer and System Sciences} {\bf 55(1)} (August 1997) 119-139}.
+
 \bibitem{ref:uboost} 
@@
 
+\bibitem{ref:F1999}
+{Jerome H. Friedman},
+\emph{Greedy Function Approximation: A Gradient Boosting Machine},
+\href{http://statweb.stanford.edu/~jhf/ftp/trebst.pdf}
+{\emph{Annals of Statistics} {\bf 29} (2000) 1189-1232}.
+
 \bibitem{Sjostrand:2006za}
-{"Sj\"{o}strand, Torbj\"{o}rn and Mrenna, Stephen and Skands, Peter"},
+T.~Sj\"{o}strand, S.~Mrenna and P.~Skands,
 \emph{PYTHIA 6.4 physics and manual},
@@
 \bibitem{LHCb-PROC-2010-056}
-{"Belyaev, I. and others"},
+I.~Belyaev, I. {\em et al.},
 \emph{Handling of the generation of primary events in GAUSS, the LHCb simulation framework},
@@
 \bibitem{Lange:2001uf}
-{"Lange, D. J."},
+D.J.~Lange,
 \emph{The EvtGen particle decay simulation package},
@@
 \bibitem{Golonka:2005pn}
-{"Golonka, Piotr and Was, Zbigniew"},
+P.~Golonka and Z.~Was,
 \emph{PHOTOS Monte Carlo: a precision tool for QED corrections in $Z$ and $W$ decays},
@@
 \bibitem{Allison:2006ve}
-{"Allison, John and Amako, K. and Apostolakis, J. and Araujo, H. and Dubois, P.A. and others"},
+J.~Allison {\em et al.},
 \emph{Geant4 developments and applications},
@@
 \bibitem{Agostinelli:2002hh}
-{"Agostinelli, S. and others"},
+S.~Agostinelli {\em et al.},
 \emph{Geant4: a simulation toolkit},
@@
 \bibitem{LHCb-PROC-2011-006}
-{"Clemencic, M and others"},
+M~Clemencic {\em et al.},
 \emph{The LHCb simulation application, GAUSS : design, evolution and experience},
@@
 {\emph{J. Phys. Conf. Ser.} {\bf 331} (2011) 032023}.
-
-\bibitem{ref:FS1997}
-{Y. Freund, and R. Schapire},
-\emph{A Decision-Theoretic Generalization of On-Line Learning and an Application to Boosting}, 
-{\emph{Journal of Computer and System Sciences} {\bf 55(1)} (August 1997) 119-139}.
-
-\bibitem{ref:F1999}
-{Jerome H. Friedman},
-\emph{Greedy Function Approximation: A Gradient Boosting Machine},
-\href{http://statweb.stanford.edu/~jhf/ftp/trebst.pdf}
-{\emph{Annals of Statistics} {\bf 29} (2000) 1189-1232}.
 
@@
 %\section{Approaches Proposed}
-\input{parts/approaches}
+%\input{parts/approaches}
 

### 2014-10-07 assets/paperdraft/parts/dataset.tex
Mike edits

@@
   \includegraphics[width=0.49\textwidth]{DP_bkgd.pdf}
-  \caption{\label{fig:dalitz} Dalitz-plot distributions for (left) signal and (right) background for the $D_s^\pm\to\pi^+\pi^-\pi^\pm$.  The three pions are labeled here as 1, 2 and 3.}
+  \caption{\label{fig:dalitz} Dalitz-plot distributions for (left) signal and (right) background for the $D_s^\pm\to\pi^+\pi^-\pi^\pm$.
+The three pions are labeled here as 1, 2 and 3 and ordered according to increases momentum.}
 \end{figure}
@@
 Figure~\ref{fig:dalitz_rocs} shows the ROC curves obtained for the various classifiers studied in this paper.
-For the UGBFL algorithms, there is a choice to be made for the value $\alpha$ which defines the relative weight of the flatness loss {\em vs} AdaLoss.
+For the uGBFL algorithms, there is a choice to be made for the value $\alpha$ which defines the relative weight of the flatness loss {\em vs} AdaLoss.
 As expected, increasing $\alpha$, which increases the weight of AdaLoss, drives the ROC curve to be similar to AdaBoost.
@@
   \includegraphics[width=0.49\textwidth]{ROC_DP_Alpha.pdf}
-  \caption{\label{fig:dalitz_rocs} (left) ROC curves for classifier algorithms studied in this paper.  (right) ROC curves for UGBFL(bin) for differnet values of $\alpha$.}
+  \caption{\label{fig:dalitz_rocs} (left) ROC curves for classifier algorithms studied in this paper.
+For the uGBFL algorithms $\alpha=0.02$ is shown.
+(right) ROC curves for uGBFL(bin) for differnet values of $\alpha$.}
 \end{figure}
@@
 This is likely due to the fact that while kNNAdaBoost uses non-local kNN information, it does not utilize global information.
-The UGBFL (binned and unbinned kNN) and uBoost algorithms each produce an efficiency which is statistically consistent with uniform across the Dalitz plot.  
-As stated above, the analyst is free to optimize the choice of $\alpha$ for UGBFL by defining a metric that involves signal efficiency, background rejection and uniformity, {\em e.g.}, using uniformity metrics discussed in detail in Appendix~A.
+The uGBkNN algorithm overcompensates and drives the efficiency higher at the corners.
+This suggests that if this algorithm is to be used some stopping criteria or {\em throttle} of the event-weighting updating should be implemented.
+The uGBFL (binned and unbinned kNN) and uBoost algorithms each produce an efficiency which is statistically consistent with uniform across the Dalitz plot.  
+As stated above, the analyst is free to optimize the choice of $\alpha$ for uGBFL by defining a metric that involves signal efficiency, background rejection and uniformity, {\em e.g.}, using uniformity metrics discussed in detail in Appendix~A.
 
@@
   \includegraphics[width=0.49\textwidth]{DP_compare.pdf}
-  \caption{\label{fig:dalitz_results} Efficiency {\em vs} distance to a corner of the Dalitz-plot.  An arbitrary working point of 50\% integrated efficiency is displayed.}
+  \caption{\label{fig:dalitz_results} Efficiency {\em vs} distance to a corner of the Dalitz-plot.
+An arbitrary working point of 50\% integrated efficiency is displayed.
+For the uGBFL algorithms $\alpha=0.02$ is shown.}
 \end{figure}
@@
 This is just an artifact of how the background sample is obtained and will not be true for background candidates {\em under} the signal peak.
-Figure~\ref{fig:md_results} shows the background mis-indentifcation rate {\em vs} $D$ candidate mass.
+Figure~\ref{fig:md_results} shows the background mis-identification rate {\em vs} $D$ candidate mass.
 AdaBoost has clearly learned to use this mis-match in signal and background candidate masses in the training.
@@
 
-Figure~\ref{fig:md_results} also shows the background mis-indentifcation rate {\em vs} $D$ candidate mass for the various uniform classifiers where $y = m(D)$ and the choice is for uniformity in the background efficiency\footnote{The algorithms in this paper can easily be made uniform on signal, background or both.}.
+Figure~\ref{fig:md_results} also shows the background mis-identification rate {\em vs} $D$ candidate mass for the various uniform classifiers where $y = m(D)$ and the choice is for uniformity in the background efficiency\footnote{The algorithms in this paper can easily be made uniform on signal, background or both.}.
 The uBoost algorithm does better than AdaBoost here but is still not optimal.
@@
 The algorithms presented in this paper each does well in achieving similar performance in the training and signal regions.
-Consider, {\em e.g.}, the UGBFL approach to achieving uniform selection efficiency.
+Consider, {\em e.g.}, the uGBFL approach to achieving uniform selection efficiency.
 In this case the training drives the BDT response itself to have the same PDF everywhere in the region $1.75 < m(D) < 1.85$~GeV (the training region).
-This does not gaurantee that the BDT response is truly independent of $m(D)$ outside the training region, but does strongly suppress learning to use $m(D)$ and in this example results in the desired behavior.
+This does not guarantee that the BDT response is truly independent of $m(D)$ outside the training region, but does strongly suppress learning to use $m(D)$ and in this example results in the desired behavior.
 Finally, if both high and low $m(D)$ sidebands had been used, it is possible for a BDT to create a fake peak near the signal peak location.
-The use of UGBFL greatly reduces the chances and possible size of such an effect.
+The use of uGBFL greatly reduces the chances and possible size of such an effect.

### 2014-10-07 assets/paperdraft/parts/metrics.tex
Mike edits

@@
 
-\subsection{Standard Deviation of Efficiency on Bins (SDE)}
+\subsection*{Standard Deviation of Efficiency on Bins (SDE)}
 
@@
 
-\subsection{Theil Index of Efficiency}
+\subsection*{Theil Index of Efficiency}
 \def\theil{\text{Theil}}
@@
 
-\subsection{Distribution Similarity Approach}
+\subsection*{Distribution Similarity Approach}
 \label{sec:similarity}
@@
 
-\subsection{Knn-based modifications}
+\subsection*{Knn-based modifications}
 
@@
 
-\subsection{Advantages and Disadvantages of Different Metrics}
-\subsubsection{Theil and SDE}
+\subsection*{Advantages and Disadvantages of Different Metrics}
+\subsubsection*{Theil and SDE}
 Let us compare two metrics that have proven most appropriate for our problem, SDE and Theil. We have some masses distributed uniformly in $[0,1]$, some constant $\alpha$ from interval $[0,1]$. 
@@
 
-\subsubsection{$D \to hhh$}
+\subsubsection*{$D \to hhh$}
 When our classifiers where applied to data set $D \to hhh$ we got uniformity measures for the SDE and Theil

### 2014-10-07 assets/paperdraft/parts/timings.tex
Mike edits

@@
        \item kNNAdaBoost: $O(k \times N)$;
-	\item UGBkNNknn: $O(k \times N)$ for $A_{\rm knn}$, and  
+	\item uGBkNNknn: $O(k \times N)$ for $A_{\rm knn}$, and  
 	$O( \text{\#nonzero elements in the matrix})$ for arbitrary matrix $A$;
-	\item UGBFL(bin): $O(N \ln N)$;
-	\item UGBFL(kNN): $O(N \ln N + N k \ln k) $.
+	\item uGBFL(bin): $O(N \ln N)$;
+	\item uGBFL(kNN): $O(N \ln N + N k \ln k) $.
 \end{itemize}
 In the example analysis studied in this paper, we find that the training time for these new algorithms is within a factor of two the same as AdaBoost.
-The CPU-resource usage of these new algorithms is not prohibative.
+The CPU-resource usage of these new algorithms is not prohibitive.

### 2014-10-07 assets/paperdraft/tmp/abstract.tex
Mike edits

The use of multivariate classifiers has become commonplace in particle physics.
To enhance the performance, a series of classifiers is typically trained; this is a technique known as boosting. 
This paper explores several novel boosting methods that have been designed to produce a uniform selection efficiency in a chosen multivariate space.  
Such algorithms have a wide range of applications in particle physics,
from producing uniform signal selection efficiency across a Dalitz-plot to avoiding the creation of false signal peaks in an invariant mass distribution when searching for new particles.

### 2014-10-07 assets/paperdraft/tmp/metrics.tex
Mike edits

In this section we come up with some approaches on how to measure uniformity of prediction. The typical way of 'checking' uniformity of prediction used by physicists 
is fitting the distribution of the events that were classified as signal (or background) over the mass (or some other variable).

This approach is hardly formalizable, and not automatable --- each time you should assume some kind of distribution. 
Ideally we want to have some easy-to-use out-of-the-box metrics like FOMs in machine learning (like area under the ROC, f1 or ).

\subsection{Ideal uniformity}

We start from the simplest case --- when we are fully satisfied by predictions of our classifier. Remember that output of 
classification is probabilities of each event being a signal and background event, only after we select some cut on probability we get classification.

Ideal uniformity of signal prediction means that whichever cut we select, the efficiency (part of signal event that passed the cut) doesn't depend on uniform variables: 
in every region of uniform variables space the part of signal events that passed the cut is the same.

In practice, of course, this never happens.

There is uniformity of predictions on signal and uniformity of efficiency on background, which can be defined one from another by swapping classes. 
In what follows in this section we are writing about the efficiency on signal events.

A good example of classifier that has close to ideal uniformity is classifier which returns a random probability in [0, 1] range. What a pity: in practice it's absolutely useless.

% The most significant drawback of the upcoming metrics is they are ill-defined if there are not too many events of the named class.

\subsection{Some Restrictions on Uniformity Metrics}

There are some additional conditions which we expect metric to meet:
\begin{enumerate}
\item
shouldn't depend much on the number of events (i.e., if we randomly select half of the events, the metrics should roughly be the same)
\item
shouldn't depend on the weights renormalization: if we multiply all the weight by some arbitrary number, it shouldn't change at all.
\item 
depends only on the order of predictions, not the exact values of probabilities.
This is because we care about which events pass the cut and which don't, not about the exact values of predictions.

Example: correlation of prediction and mass doesn't satisfy this restriction.
\item
parameter stability: if it uses bins, changing the number of bins shouldn't affect the metrics value much, if it uses $k$-nearest neighbors, it should be stable to small deviations of $k$.
\end{enumerate}

\subsection{Standard Deviation of Efficiency on Bins (SDE)}

\def\bineff{\text{eff}_\text{bin}}
\def\binweight{\text{weight}_\text{bin}}
\def\globaleff{\text{eff}}
\def\SDE{\text{SDE}}
\def\bin{\text{bin}}

Let's split the space of uniform features into bins, the uniformity in these terms is following: when we select some probability cut, 
the part of signal events that passes the cut is equal in all bins. Assume we selected some cut, then we have global efficiency
\[
	\globaleff = \dfrac{
		\text{total weight of signal events that passed the cut}}
		{\text{total weight of signal events}}
\]

Efficiency in every bin is defined respectively, 
\[
	\bineff = \dfrac{
		\text{weight of signal events in bin that passed the cut}}
		{\text{weight of signal events in this bin}} 
\]

So, basically, what we want to have in our dreams:
\[
	\bineff = \text{global efficiency} \qquad \forall \; \text{bin}
\]

To measure how far we are from ideal situation we use standard deviation:
\[
	\sqrt{\sum_{\bin} \left( \bineff - \globaleff \right)^2  }
\]

What is bad in this formula that every bin has some impact in the result, which does not depend on how many events are there, so metrics becomes very unstable to deviations in bins with only few events. To cure this, we add weights to the bins (note that $\sum_\bin \binweight = 1$):
\[
	\binweight = \dfrac{\text{total weight of signal events in bin}}
		{\text{total weight of signal events}},
\]
so we have SDE formula:
\[
	\SDE(\globaleff) = 
	\sqrt{\sum_{\bin} \binweight \times \left(\bineff - \globaleff \right)^2}. 
\] 
In fact, the expression depends on the cut, but for cuts which produce equal efficiency, this is

Finally we note that the weighted average of $\bineff$ is $\globaleff$:
\[
	\globaleff = < \bineff > =  \sum_{\bin} \binweight \times \bineff,
\]
and this is why the introduced metrics was named SDE --- this is a weighted standard deviations of array of bin efficiencies.

But this is how we measure the non-uniformity for only one fixed cut, to measure the overall non-flatness, we take several global efficiencies (for instance, [0.5, 0.6, 0.7, 0.8, 0.9], because in practice usually we are interested in cuts with high global efficiency) and use 
\[
	\SDE^2  =  \frac{1}{k} 
	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
		\text{SDE}^2(\globaleff)
\]

Some other power $p \neq 2$ can be used as well, but $p=2$ is considered as the default value: 
\[
	\SDE^p(\globaleff) = 
	\sum_{\bin} \binweight \times \abs{\bineff - \globaleff}^p,
\qquad
	\SDE^p  =  \frac{1}{k} 
	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
		\text{SDE}^p(\globaleff).
\]

\subsection{Theil Index of Efficiency}
\def\theil{\text{Theil}}

One more measure uses Theil Index frequently used to measure economic inequality:
\[
	\theil = \frac{1}{N} \sum_i \frac{x_i}{<x>} \ln{\frac{x_i}{<x>}}, 
		\qquad <x> = \frac{1}{N} \sum_i x_i
\]
In our case we have to alter formula a bit to take into account that different bins have different impact, thus the formula turns into
\[
	\theil(\globaleff) = \sum_\bin \binweight \; \frac{\bineff}{\globaleff} \; \ln{\frac{\bineff}{\globaleff}}
\]

TODO how to combine Theil for different global efficiencies?
\[
	\theil = ??? \text{from} \theil(\globaleff)
\]

\subsection{Distribution Similarity Approach}
\label{sec:similarity}

Let's start from reformulation of what is uniform predictions in signal. First we split all signal events into some bins in uniform variables. There is some empirical distribution $F_\bin$ of predictions in each bin. Ideal uniformity means that all the distributions $F_\bin$ are equal and hence equal to the global distribution $F(x)$.

To 'measure' non-flatness we can use some distribution distance, like Kolmogorov-Smirnov:
\[
	 \sum_{\bin} \binweight \max_x \abs{F_{\bin}(x) - F(x)},
\]
but Cram\'er--von Mises similarity is more informative (usually $p=2$ is used):
\[
	 \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x),
\]

The good point is we don't need to select some global efficiencies like in the other metrics.

\subsection{Connection Between SDE and Distance Similarity Approach}

SDE and DSA based on Cram\'er--von Mises similarity can be shown to have connection.
Let's consider the SDE with global efficiencies $= [1/N, 2/N, \dots, N/N]$. In the limit $N \to \infty$
\[
	\lim_{N \to \infty} \SDE^2 = 
	\lim_{N \to \infty} \frac{1}{N} \sum_{\globaleff}\SDE^2(\globaleff) = 
	\int_0^1 \SDE^2(\globaleff) d\, \globaleff = 
	\int_0^1 \sum_{\bin} \binweight \abs{\bineff - \globaleff}^2 d\, \globaleff
\]

From the other side, we can write the expression for similarity-based measure (for $p=2$) 
\[
	\sum_{\bin} \binweight \int \abs{F_\text{bin}(x) - F(x)}^2 dF(x) =
	\int \sum_{\bin} \binweight \abs{F_\text{bin}(x) - F(x)}^2 dF(x) 
\] The hard thing now is to believe this is literally the same and these two expressions are equal.

\subsection{Knn-based modifications}

\def\knni{\text{knn}(i)}
\def\effknni{\text{eff}_{\knni}}
\def\weightknni{\text{weight}_{\knni}}
\def\Fknn{F_{\knni}}

\def\knnSDE{\text{knnSDE}}

Though operating with bins is usually both simple and very efficient, 
in many cases it is hard to find optimal size of bins in the space of uniform variables (specifically in the case of more than two dimensions).
One more situation when bins-based approach fails, is when we have too few events to obtain a good statistics at least in several bins.

In these cases we can switch to $k$-nearest neighbors: for each signal event we find $k$ nearest signal events (including the event itself) in the space of uniform variables. Now we can compute the efficiency $\effknni$, empirical distribution $\Fknn$ of nearest neighbors. 
The weights for $\knni$ are proportional to the total weight of events in $\knni$:
\[
	\weightknni = \alpha \sum_{j \in \knni} w_j, \qquad \alpha^{-1} = \sum_i \sum_{j \in \knni} w_j,
\]
so again weights are normed to 1: $\sum_{i} \knni = 1$.

Now we are ready to write knn version of SDE:
\[
	\knnSDE^2(\globaleff)
		= \sum_{i \in \text{events}} \weightknni \abs{\effknni - \globaleff}^2
\]
\[
	\knnSDE^2 = \sum_{\globaleff \in [\globaleff_1, \dots \globaleff_k]}
		\knnSDE^2(\globaleff),
\]
knn version of Theil index of Efficiency
\[
	\text{knnTheil}(\globaleff) = \sum_{i \in \text{events}} \weightknni \; \frac{\effknni}{\globaleff} \; \ln{\frac{\effknni}{\globaleff}}
\]
\[
	\text{knnTheil} = ??? \text{knnTheil}(\globaleff)
\]
and knn version of similarity-based measure:

\[
	 \sum_{i \in \text{events}} \weightknni \int \abs{\Fknn(x) - F(x)}^p dF(x),
\]

$K$-nearest neighbors approach suffers from the other drawback: the impact of different events has very little connection with the weights, because some events are met in knn of other events much more frequently while the other.
This effect can be suppressed by dividing initial weight of the event by the number of times it is met in knn.

\subsection{Advantages and Disadvantages of Different Metrics}

... TODO, here some plots with comparison, maybe timings ...

### 2014-10-07 assets/paperdraft/tmp/paper_draft.tex
Mike edits

--- 2014-09-20 assets/paperdraft/paper_draft.tex
@@
 \author{
-Aleksandar Bukva$^e$, 
+Aleksandar Bukva$^{\rm ADD}$, 
 Vladimir Gligorov$^d$,
-Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~,\newline
-Andrey Ustyuzhanin$^{b,f}$ and
+Alex Rogozhnikov$^{a,b}$\thanks{Corresponding author.}~ ,
+Andrey Ustuzhanin$^b$ and
 Mike Williams$^c$\\
 \llap{$^a$}Lomonosov Moscow State University, Moscow, Russia\\
-\llap{$^b$}Yandex, Moscow, Russia\\
+\llap{$^b$}Yandex LLC, Moscow, Russia\\
 \llap{$^c$}Massachusetts Institute of Technology, Cambridge, MA, United States \\
 \llap{$^d$}Organisation Europ\'eenne pour la Recherche Nucl\'eaire (CERN), Geneva, Switzerland  \\
-\llap{$^e$}Faculty of Physics, Belgrade, Serbia \\
-\llap{$^f$}Moscow Institute of Physics and Technology, Moscow, Russia\\
 E-mail: \email{alex.rogozhnikov@yandex.ru}}
@@
 % AMS packages:
-
 \usepackage{amsmath, amsthm, amsfonts, xspace}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{float}
 
@@
 
-\def\x {\ensuremath{\vec{x}}\xspace}
-\def\y {\ensuremath{\vec{y}}\xspace}
+\def \x {\ensuremath{\vec{x}}\xspace}
+\def \y {\ensuremath{\vec{y}}\xspace}
 \def\knn{\text{knn}}
@@
 \input{parts/introduction}
+
+%\section{Uniformity Measurement}
+\input{parts/metrics}
+%\section{Approaches Proposed}
+\input{parts/approaches}
 
@@
 
-These results were obtained using events generated with the official LHCb simulation, and we are grateful to the LHCb collaboration for this privilege. 
-We particularly acknowledge the work of the LHCb Simulation and Core Computing teams in tuning the simulation software and managing the productions of simulated events.
-
 \begin{thebibliography}{9}
 
-\bibitem{ref:lhcbhlt} 
-R. Aaij {\em et al.} [LHCb Trigger Group], 
-{\em The LHCb trigger and its performance}, 
-JINST {\bf 8}, P04022 (2013). 
-[arXiv:1211.3055]
+\bibitem{ref:lhcbhlt} R. Aaij {\em et al.} [LHCb Trigger Group], {\em The LHCb trigger and its performance}, JINST {\bf 8}, P04022 (2013). [arXiv:1211.3055]
 
-\bibitem{ref:bbdt}  
-V. Gligorov and M. Williams, 
-{\em Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree}, 
-JINST {\bf 8}, P02013 (2013). 
-[arXiv:1210.6861]
+\bibitem{ref:bbdt}  V. Gligorov and M. Williams, {\em Efficient, reliable and fast high-level triggering using a bonsai boosted decision tree}, JINST {\bf 8}, P02013 (2013). [arXiv:1210.6861]
 
-\bibitem{ref:uboost} 
-J. Stevens and M. Williams,
-{\em uBoost: A boosting method for producing uniform selection efficiencies from multivariate classifiers}, 
-JINST {\bf 8}, P12013 (2013). 
-[arXiv:1305.7248]
-
-\bibitem{Sjostrand:2006za}
-{"Sj\"{o}strand, Torbj\"{o}rn and Mrenna, Stephen and Skands, Peter"},
-\emph{PYTHIA 6.4 physics and manual},
-\href{http://dx.doi.org/10.1088/1126-6708/2006/05/026},
-{\emph{JHEP} {\bf 05} (2006) 026}.
-[\hepph{0603175}]
-
-\bibitem{LHCb-PROC-2010-056}
-{"Belyaev, I. and others"},
-\emph{Handling of the generation of primary events in GAUSS, the LHCb simulation framework},
-\href{http://dx.doi.org/10.1109/NSSMIC.2010.5873949},
-{\emph{NSS/MIC} {\bf IEEE} (2010) 1155}.
-
-\bibitem{Lange:2001uf}
-{"Lange, D. J."},
-\emph{The EvtGen particle decay simulation package},
-\href{http://dx.doi.org/10.1016/S0168-9002(01)00089-4},
-{\emph{NIM} {\bf A462} (2001) 152-155}.
-
-\bibitem{Golonka:2005pn}
-{"Golonka, Piotr and Was, Zbigniew"},
-\emph{PHOTOS Monte Carlo: a precision tool for QED corrections in $Z$ and $W$ decays},
-\href{http://dx.doi.org/10.1140/epjc/s2005-02396-4},
-{\emph{EPJ} {\bf C45} (2006) 97-107}.
-[\hepph{0506026}]
-
-\bibitem{Allison:2006ve}
-{"Allison, John and Amako, K. and Apostolakis, J. and Araujo, H. and Dubois, P.A. and others"},
-\emph{Geant4 developments and applications},
-\href{http://dx.doi.org/10.1109/TNS.2006.869826},
-{\emph{IEEE TNS} {\bf 53} (2006) 270}.
-
-\bibitem{Agostinelli:2002hh}
-{"Agostinelli, S. and others"},
-\emph{Geant4: a simulation toolkit},
-\href{http://dx.doi.org/10.1016/S0168-9002(03)01368-8},
-{\emph{NIM} {\bf A506} (2003) 250}.
-
-\bibitem{LHCb-PROC-2011-006}
-{"Clemencic, M and others"},
-\emph{The LHCb simulation application, GAUSS : design, evolution and experience},
-\href{http://dx.doi.org/10.1088/1742-6596/331/3/032023},
-{\emph{J. Phys. Conf. Ser.} {\bf 331} (2011) 032023}.
-
-% The examples of items:
-% \bibitem{bib1}
-% Authors,
-% \emph{Title},
-% \emph{J. Ref.} \textbf{vol} (year) page.
-
-% \bibitem{bib2}
-% A. Pietropaolo et al., \emph{DINS measurements on VESUVIO in the
-%     resonance detector configuration: proton mean kinetic energy in
-%     water}, \jinst{1}{2006}{P04001}.
-
-% \bibitem{bib3}
-% A.I. Harris,
-% \emph{Spectroscopy with multichannel correlation radiometers},
-% \href{http://dx.doi.org/10.1063/1.1898643}
-% {\emph{Rev.\ Sci.\ Instrum.} {\bf 76} (2005) 054503}
-% [\astroph{0504449}].
-
-% \bibitem{bib4}
-% G.F. Knoll, \emph{Radiation detection and measurements}, John Wiley
-%     and Sons, Inc., New York 2000.
-
-% \bibitem{bib5}
-% V. Dangendorf, \emph{Time-resolved fast-neutron imaging with a
-% pulse-counting image intensifier}, in proceedings of
-% \emph{International workshop on fast neutron detectors and
-% applications}, April, 3--6, 2006 University of Cape Town, South Africa
-% \pos{PoS(FNDA2006)008}.
+\bibitem{ref:uboost} J.
+Stevens and M.
+Williams, {\em uBoost: A boosting method for producing uniform selection efficiencies from multivariate classifiers}, JINST {\bf 8}, P12013 (2013) [arXiv:1305.7248]
 
@@
 
-Clean all this up...
-
-%\section{Uniformity Measurement}
-\input{parts/metrics}
-%\section{Approaches Proposed}
-\input{parts/approaches}
+ADD THEIL, ETC DESCRIPTIONS AND HOW TO IMPLEMENT IN THE ALGORITHM HERE.
 

### 2014-10-07 assets/paperdraft/tmp/timings.tex
Mike edits

The main drawback of uBoost technique is it's high computational complexity: 
while simple AdaBoost trains $M$ trees, uBoost builds $100 \times M$ trees (contribution of other operations usually can be neglected).

Presented in this paper classifiers are building $M$ trees, though there is more complicated boosting, it takes at each iteration ($k$ is number of neighbors, $N$ is number of events in training sample)

\begin{itemize}
	\item meanAdaBoost: $O(k \times N)$
	\item knnAdaLoss: $O(k \times N)$, for the arbitrary matrix $A$ it is 
	$O( \text{\#nonzero elements in the matrix})$
	\item FlatnessLoss: $O(N \ln N)$
\end{itemize}

### 2014-10-07 assets/paperdraft/tmp/ugb.tex
Mike edits

--- 2014-09-17 assets/paperdraft/ugb.tex
@@
 
-Another approach to obtain uniformity, introduced in this paper, involves defining a more general expression of the AdaBoost criteria:
+Another approach to obtain uniformity involves defining a more general expression of the AdaBoost criteria:
 \begin{equation}
@@
 \end{equation}
-where $F_{(b)}(s)$ denotes the cummulative distribution of $f_{(b)}(s)$ and $w_b = \sum \delta({\rm bin}_i - b)/n_{\rm signal}$, {\em i.e.} $w_b$ is the fraction of signal events in the bin\footnote{If weighted events are used, then the fractional sum of weights should be used for $w_b$.}.
+where $F_{(b)}(s)$ denotes the cummulative distribution of $f_{(b)}(s)$ and $w_b = \sum \delta({\rm bin}_i - b)/n_{\rm signal}$, {\em i.e.} $w_b$ is the fraction of signal events in the bin.
 
@@
 \begin{equation}
- -\frac{\partial L_{\rm flat}}{\partial s_k} \approx - 2\left[F_b(s_k)-F(s_k)\right]/n_{\rm signal}.
+ -\frac{\partial L_{\rm flat}}{\partial s_k} = - w_b {\rm sign}\left[F_b(s_k)-F(s_k)\right] \left( w_k/w_b - w_k \right).  
 \end{equation}
@@
 
-The loss function given in Eq.~\ref{eq:L_ada_flat} can also be constructed without binning the data using k-nearest-neighbor events.
-The cummulative distribution $F_{\rm knn}(s)$ is easily obtained and the bin weight, $w_b$, is replaced by ???
+ADD KNN HERE ONCE ALGORITHM IS TESTED.

### 2014-10-07 assets/paperdraft/ugb.tex
Mike edits

@@
 The variates used in the BDT are denoted by \x, while the variates in which uniformity is desired are denoted by \y.
-Some (perhaps all) of the \x variantes will be {\em biasing} in \y, {\em i.e.} they provide discriminating power between signal and background that varies in \y.
+Some (perhaps all) of the \x variates will be {\em biasing} in \y, {\em i.e.} they provide discriminating power between signal and background that varies in \y.
 A uniform BDT selection efficiency can be obtained by removing all such variates; however, this will also reduce the power of the BDT.
@@
 
-One category of boosting works by assigning training events more weight based on classification erros made by previous members of the series.
-For example, the AdaBoost~\cite{ref:adaboost} algorithm updates the weight of event $i$, $w_i$, according to
+One category of boosting works by assigning training events more weight based on classification errors made by previous members of the series.
+For example, the AdaBoost~\cite{ref:FS1997} algorithm updates the weight of event $i$, $w_i$, according to
 \begin{equation}
@@
 where $\gamma = +1(-1)$ for signal(background) events and $p$ is the prediction for each event produced by last classifier in the series.  
-The uBoost technique, described in detail in Ref.~\cite{cit:uboost}, alters the event-weight updating procedure to achieve uniformity in the signal-selection efficiency.
+The uBoost technique, described in detail in Ref.~\cite{ref:uboost}, alters the event-weight updating procedure to achieve uniformity in the signal-selection efficiency.
 
@@
 
-The gradient boosting~\cite{ref:gradboost} (GB) algorithm category requires the analyst to choose a differentiable loss function with the goal of building a classifier that minimizes the loss.  
-A popular choice of loss function is the AdaLoss~\cite{ref:adaboost} function 
+The gradient boosting~\cite{ref:F1999} (GB) algorithm category requires the analyst to choose a differentiable loss function with the goal of building a classifier that minimizes the loss.  
+A popular choice of loss function is the so-called AdaLoss function 
 \begin{equation}
@@
 \end{equation}
-So-called scores $s$ are obtained for each event as sum of predictions of all regressots in the series
+The {\em scores} $s$ are obtained for each event as sum of predictions of all elements in the series. 
 At each stage in the gradient boosting process, a {\em regressor} (a decision tree in our case) is trained whose purpose is to decrease the loss.
@@
 \end{equation}
-which are positive(negative) for signal(background) events and have larger modulai for poorly classified events.
+which are positive(negative) for signal(background) events and have larger moduli for poorly classified events.
 
@@
 One can see that the direction of the gradient will be influenced the most by events whose $k$-nearest-neighbor events are classified poorly. 
+We generically refer to GB methods designed to achieve uniform selection efficiency as uniform GB (uGB).  The specific algorithm that uses kNNAdaLoss will be called uGBkNN.
 
@@
 
-The gradient of the Kolmogorov-Smirnov loss function is zero for events with responses not equal to than the value of $s$ at which ${\rm max}|F_b(s)-F(s)|$ occurs.
+The gradient of the Kolmogorov-Smirnov loss function is zero for events with responses greater than the value of $s$ at which ${\rm max}|F_b(s)-F(s)|$ occurs.
 Thus, it is not suitable for gradient boosting due to its instability.
@@
 \end{equation}
-whose pseudo-residuals are ($b$ is bin containing $k$th event)
+whose pseudo-residuals are ($b$ is the bin containing the $k$th event)
 \begin{equation}
@@
 The first term in Eq.~\ref{eq:L_ada_flat} penalizes non-uniformity, while that second term penalizes poor classification.
+We refer to this algorithm as uGB with flatness loss (uGBFL). 
+In principle, many different flatness loss functions can be defined and could be substituted for our choice here.  See Appendix~A for a detailed discussion on this topic.
 
@@
 This procedure is followed to offset the fact that some events are found in more k-nearest-neighbor samples than other events.
+We study two versions of uGBFL below: uGBFL using bins denoted by uGBFL(bin) and uGBFL using kNN collections denoted by uGBFL(kNN).  
+The algorithms are summarized in Table~\ref{tab:algs}.
+
+\begin{table}
+  \begin{center}
+    \caption{\label{tab:algs} Description of uniform boosting algorithms.}
+    \begin{tabular}{c|c}
+      %\toprule
+      \hline
+      Name & Description \\
+      \hline
+      %\midrule
+        uBoost & algorithm introduced in Ref.~\cite{ref:uboost}\\
+      \hline
+      %\midrule 
+        kNNAda & AdaBoost modification using matrix $A_{\rm knn}$ \\
+        uGBkNN & gradient boost using kNNAdaLoss loss function \\
+        uGBFL(bin) & gradient boost using flatness loss $+ \alpha$ AdaLoss as in Eq.~\ref{eq:L_ada_flat} (data binned for FL) \\
+        uGBFL(kNN) & same as uGBFL(bin) except kNN events are used rather than bins \\
+      \hline
+      %\bottomrule
+    \end{tabular}
+  \end{center}
+\end{table}

### 2014-10-08 IPythonWorkflow/metrics.py
deleted multiplication by 10 in SDE, minor fixes

@@ -6,8 +6,6 @@
 from __future__ import division
 from __future__ import print_function
 
-import warnings
-
 import numpy
 import pandas
 from sklearn.metrics import auc
@@ -27,11 +25,13 @@ __all__ = ['sde', 'cvm_flatness', 'cvm_2samp', 'theil']
 
 #region Utilities
 
+
 def compute_cdf(ordered_weights):
     """Computes cumulative distribution function (CDF) by ordered weights,
     be sure that sum(ordered_weights) == 1
     """
     return numpy.cumsum(ordered_weights) - 0.5 * ordered_weights
+
 
 def generate_test_dataset(n_samples, n_bins):
     random = RandomState()
@@ -369,7 +369,7 @@ def compute_sde_on_bins(y_pred, mask, bin_indices, target_efficiencies, power=2.
                                                     cut=cut, sample_weight=sample_weight)
         result += weighted_deviation(bin_efficiencies, weights=bin_weights, power=power)
 
-    return 10 * (result / len(cuts)) ** (1./power)
+    return (result / len(cuts)) ** (1./power)
 
 
 def compute_sde_on_groups(y_pred, mask, groups_indices, target_efficiencies, sample_weight=None, power=2.):
@@ -382,7 +382,7 @@ def compute_sde_on_groups(y_pred, mask, groups_indices, target_efficiencies, sam
         group_efficiencies = compute_group_efficiencies(y_pred, groups_indices=groups_indices,
                                                         cut=cut, sample_weight=sample_weight)
         sde += weighted_deviation(group_efficiencies, weights=group_weights, power=power)
-    return 10 * (sde / len(cuts)) ** (1./power)
+    return (sde / len(cuts)) ** (1./power)
 
 
 def sde(y, proba, X, uniform_variables, sample_weight=None, label=1, knn=30):
@@ -516,11 +516,18 @@ def _ks_2samp_fast(prepared_data1, data2, prepared_weights1, weights2, F1):
     F2 = compute_cdf(prepared_weights2)
     return numpy.max(numpy.abs(F1 - F2))
 
+
 def ks_2samp_weighted(data1, data2, weights1, weights2):
     x = numpy.unique(numpy.concatenate([data1, data2]))
[…]

### 2014-10-09 IPythonWorkflow/metrics.py
Created DemoMetrics notebook

@@ -730,7 +730,7 @@ def check_limit(size=2000):
     val2 = compute_sde_on_bins(y_pred, mask=mask, bin_indices=bin_indices, target_efficiencies=effs,
                                sample_weight=sample_weight)
 
-    assert numpy.allclose(val1, (val2 / 10) ** 2, atol=1e-3, rtol=1e-2)
+    assert numpy.allclose(val1, val2 ** 2, atol=1e-3, rtol=1e-2)
 
 
 check_limit()

### 2014-10-12 assets/paperdraft/parts/metrics.tex
update plots

@@
 		\centering
-		\begin{subfigure}[b]{0.8\textwidth}
+		\begin{subfigure}[b]{0.48\textwidth}
 			\includegraphics[width=\textwidth]{graphs/bins_uniform_distribution.pdf}
-			\caption{Predictions are uniform in mass, the distribution of predictions in the bin (yellow) is close to the global (blue).
-			Yellow rectangle shows the events in the bin over mass.}
 		\end{subfigure}
-		\begin{subfigure}[b]{0.8\textwidth}
+		\begin{subfigure}[b]{0.48\textwidth}
 			\includegraphics[width=\textwidth]{graphs/bins_nonuniform_distribution.pdf}
-			\caption{Distribution with peak in the middle, the distribution in the bin is quite different from global.}
 		\end{subfigure}
-		\caption{Demonstration of distribution similarity approach. \label{fig:dsavisualization}}
+		\caption{Demonstration of the distribution similarity approach.
+(left) Predictions are uniform in mass, the distribution of predictions in the bin (yellow) is close to the global (blue).
+(right) Distribution with peak in the middle, the distribution in the bin is quite different from the global distribution.
+In both cases the yellow rectangle shows the events in the bin over mass.\label{fig:dsavisualization}}
 \end{figure}
@@
 \centering
-	\includegraphics[width=0.5\textwidth]{graphs/TheilVsMSE.png}
+	\includegraphics[width=0.9\textwidth]{graphs/TheilVsMSE.png}
 	\caption{Comparison of SDE and Theil uniformity measures for different values of the constant $\alpha$ for the peak and pit distributions.
@@
 metrics shown in Fig.~\ref{fig:d2hhhsdetheil}.
+Both metrics show similar results, so there is no significant difference between them, nor the CvM metric, on this dataset.
+
 \begin{figure}[h]
@@
 		% TODO use metrics on signal for signal, use metrics on bck for bck.
-		\begin{subfigure}[b]{0.95\textwidth}
+		\begin{subfigure}[b]{0.48\textwidth}
 			\includegraphics[width=\textwidth]{graphs/D23hSignalEffs.png}
@@
 		\end{subfigure}
-		\begin{subfigure}[b]{0.95\textwidth}
+		\begin{subfigure}[b]{0.48\textwidth}
 			\includegraphics[width=\textwidth]{graphs/D23hBgEffs.png}
@@
 \end{figure}
-
-Both metrics shown similar results, and that there is no significant difference on real data sets (as well as CvM).

### 2014-10-12 IPythonWorkflow/metrics.py

@@ -733,6 +733,6 @@ def check_limit(size=2000):
     assert numpy.allclose(val1, val2 ** 2, atol=1e-3, rtol=1e-2)
 
 
-check_limit()
+#check_limit()
 
 #endregion

### 2014-10-12 assets/paperdraft/graphs/plotd23hflatnessinpaperstyle.py

@@ -4,13 +4,13 @@ from ROOT import *
 from math import *
 from array import *
 
-basedir = './dto3h/Bg/'
+basedir = './dto3h/Signal/'
 marker = 20
 size = 0.5
 colour = 1
 
-filestoplot = { 'data_sde_bg'   : {'filename':'data_sde_bg'  ,'subclassifiers' : {}},
-                'data_theil_bg' : {'filename':'data_theil_bg','subclassifiers' : {}}}
+filestoplot = { 'data_sde_sig'   : {'filename':'data_sde_sig'  ,'subclassifiers' : {}},
+                'data_theil_sig' : {'filename':'data_theil_sig','subclassifiers' : {}}}
 
 subclassifiercols = {'ada':1,'uGB+knnFL':5,'uBoost':6}
 
@@ -29,7 +29,10 @@ for entry in filestoplot :
       continue
     for subfile in ['ada','uGB+knnFL','uBoost'] :    
       filestoplot[entry]['subclassifiers'][subfile]['x'].append(float(line.split(',')[0]))
-      filestoplot[entry]['subclassifiers'][subfile]['y'].append(float(line.split(',')[subclassifiercols[subfile]].strip('\n')))
+      if entry.find('sde') > -1 :
+        filestoplot[entry]['subclassifiers'][subfile]['y'].append(float(line.split(',')[subclassifiercols[subfile]].strip('\n')))
+      else :
+        filestoplot[entry]['subclassifiers'][subfile]['y'].append(100.*float(line.split(',')[subclassifiercols[subfile]].strip('\n')))
     num += 1
 
   i = 0   
@@ -41,19 +44,23 @@ for entry in filestoplot :
     filestoplot[entry]['subclassifiers'][subfile]['graph'].SetMarkerSize(size)
     filestoplot[entry]['subclassifiers'][subfile]['graph'].SetMarkerStyle(marker+i)
     filestoplot[entry]['subclassifiers'][subfile]['graph'].GetXaxis().SetTitle('classifier stage')
-    filestoplot[entry]['subclassifiers'][subfile]['graph'].GetXaxis().SetNdivisions(510)
+    filestoplot[entry]['subclassifiers'][subfile]['graph'].GetXaxis().SetNdivisions(505)
[…]

### 2014-10-12 assets/paperdraft/graphs/plotdistsinpaperstyle.py

@@ -4,15 +4,15 @@ from ROOT import *
 from math import *
 from array import *
 
-basedir = './data/raspodela1/raspodele/'
+basedir = './data/raspodela2/raspodele/'
 marker = 20
 size = 0.5
 colour = 1
 
 filestoplot = { '#alpha=0.1':{'filename':'0_raspodela'},
-                '#alpha=0.28':{'filename':'1_raspodela'},
+                #'#alpha=0.28':{'filename':'1_raspodela'},
                 '#alpha=0.47':{'filename':'2_raspodela'},
-                '#alpha=0.65':{'filename':'3_raspodela'},
+                #'#alpha=0.65':{'filename':'3_raspodela'},
                 '#alpha=0.84':{'filename':'4_raspodela'}}
 
 for entry in filestoplot :
@@ -34,22 +34,25 @@ for entry in filestoplot :
   filestoplot[entry]['graph'].GetXaxis().SetTitle('mass')
   filestoplot[entry]['graph'].GetXaxis().SetNdivisions(510)
   filestoplot[entry]['graph'].GetYaxis().SetTitle('prediction')
-  filestoplot[entry]['graph'].GetXaxis().SetTitleSize(0.08)
+  filestoplot[entry]['graph'].GetXaxis().SetLabelSize(0.075)
+  filestoplot[entry]['graph'].GetXaxis().SetLabelOffset(0.015)
+  filestoplot[entry]['graph'].GetXaxis().SetTitleSize(0.10)
   filestoplot[entry]['graph'].GetXaxis().SetTitleOffset(0.6)
-  filestoplot[entry]['graph'].GetYaxis().SetTitleSize(0.08)
-  filestoplot[entry]['graph'].GetYaxis().SetTitleOffset(0.85)
+  filestoplot[entry]['graph'].GetYaxis().SetTitleSize(0.10)
+  filestoplot[entry]['graph'].GetYaxis().SetLabelSize(0.075)
+  filestoplot[entry]['graph'].GetYaxis().SetTitleOffset(0.65)
 
   thisfile.close() 
 
 #print filestoplot
 
-c1 = TCanvas("c1","c1",1600,800)
-c1.Divide(5)
+c1 = TCanvas("c1","c1",1920,500)
+c1.Divide(3)
 i=1
 for entry in filestoplot :
   c1.cd(i)
[…]

### 2014-10-12 assets/paperdraft/graphs/ploteffsinpaperstyle.py

@@ -4,15 +4,15 @@ from ROOT import *
 from math import *
 from array import *
 
-basedir = './data/raspodela2/efikasnosti/'
+basedir = './data/raspodela1/efikasnosti/'
 marker = 20
 size = 0.5
 colour = 1
 
 filestoplot = { '#alpha=0.1':{'dirname':'0','subfiles' : {}},
-                '#alpha=0.28':{'dirname':'1','subfiles' : {}},
+                #'#alpha=0.28':{'dirname':'1','subfiles' : {}},
                 '#alpha=0.47':{'dirname':'2','subfiles' : {}},
-                '#alpha=0.65':{'dirname':'3','subfiles' : {}},
+                #'#alpha=0.65':{'dirname':'3','subfiles' : {}},
                 '#alpha=0.84':{'dirname':'4','subfiles' : {}}}
 
 suffix = 'eff'
@@ -42,18 +42,21 @@ for entry in filestoplot :
     filestoplot[entry]['subfiles'][subfile]['graph'].GetXaxis().SetTitle('mass')
     filestoplot[entry]['subfiles'][subfile]['graph'].GetXaxis().SetNdivisions(510)
     filestoplot[entry]['subfiles'][subfile]['graph'].GetYaxis().SetTitle('efficiency')
-    filestoplot[entry]['subfiles'][subfile]['graph'].GetXaxis().SetTitleSize(0.08)
+    filestoplot[entry]['subfiles'][subfile]['graph'].GetXaxis().SetLabelSize(0.075)
+    filestoplot[entry]['subfiles'][subfile]['graph'].GetXaxis().SetLabelOffset(0.015)
+    filestoplot[entry]['subfiles'][subfile]['graph'].GetXaxis().SetTitleSize(0.10)
     filestoplot[entry]['subfiles'][subfile]['graph'].GetXaxis().SetTitleOffset(0.6)
-    filestoplot[entry]['subfiles'][subfile]['graph'].GetYaxis().SetTitleSize(0.08)
-    filestoplot[entry]['subfiles'][subfile]['graph'].GetYaxis().SetTitleOffset(0.85)
+    filestoplot[entry]['subfiles'][subfile]['graph'].GetYaxis().SetTitleSize(0.10)
[…]

### 2014-10-13 assets/paperdraft/parts/introduction.tex
small textual edits

@@
 The classifiers are trained in series with the inputs of each member being augmented based on the performance of its predecessors.
-This augmentation is designed such that each new classifier targets more those events which were poorly classified by previous members of the series.
+This augmentation is designed such that each new classifier targets those events which were poorly classified by previous members of the series.
 The classifier obtained by combining all members of the series is typically much more powerful than any of the individual members.
@@
 This approach is optimal for a counting experiment; however, in many cases the BDT-based selection obtained in this way is not optimal.
-For example, in a Dalitz-plot (or any angular or amplitude analysis) analysis, obtaining a selection efficiency on signal candidates that is uniform across the Dalitz-plot is more important than any integrated FOM.
+For example, in a Dalitz-plot (or any angular or amplitude analysis) analysis, obtaining a selection efficiency for signal candidates that is uniform across the Dalitz-plot is more important than any integrated FOM.
 Similarly, when measuring a mean particle lifetime, obtaining an efficiency that is uniform in lifetime is what is desired.
@@
 When searching for a new particle, an analyst may want a uniform efficiency in mass for selecting background candidates so that the BDT-based selection does not generate a fake signal peak.
-Furthermore, the analyst may also desire a uniform selection efficiency of signal candidates in mass (or other variates) since the new particle mass is not known.
+Furthermore, the analyst may also desire a uniform selection efficiency of signal candidates in mass (or other variates) since the mass of the new particle is not known.
 In such cases, the BDT is often trained on simulated data generated with several values of mass (lifetime, {\em etc.}).

### 2014-10-13 assets/paperdraft/parts/metrics.tex
small textual edits

@@
 		\begin{subfigure}[b]{0.95\textwidth}
-			\includegraphics[width=\textwidth]{graphs/PeakEffs.png}
+			\includegraphics[width=\textwidth]{graphs/PeakEffs.pdf}
 			\caption{Mass vs efficiency. The different coloured lines correspond to
@@
 		\begin{subfigure}[b]{0.95\textwidth}
-			\includegraphics[width=\textwidth]{graphs/PitEffs.png}
+			\includegraphics[width=\textwidth]{graphs/PitEffs.pdf}
 			\caption{Mass vs efficiency. The different coloured lines correspond to
@@
 \centering
-	\includegraphics[width=0.9\textwidth]{graphs/TheilVsMSE.png}
+	\includegraphics[width=0.9\textwidth]{graphs/TheilVsMSE.pdf}
 	\caption{Comparison of SDE and Theil uniformity measures for different values of the constant $\alpha$ for the peak and pit distributions.
@@
 \subsubsection*{$D \to hhh$}
-When our classifiers where applied to data set $D \to hhh$ we got uniformity measures for the SDE and Theil
-metrics shown in Fig.~\ref{fig:d2hhhsdetheil}.
-Both metrics show similar results, so there is no significant difference between them, nor the CvM metric, on this dataset.
+Finally, we compare the SDE and Theil uniformity measures for the classifiers applied to the $D \to hhh$ data set, as shown in Fig.~\ref{fig:d2hhhsdetheil}.
+Both measures show similar results, so there is no significant difference between them, nor the CvM metric, on this dataset.
 
@@
 		\begin{subfigure}[b]{0.48\textwidth}
-			\includegraphics[width=\textwidth]{graphs/D23hSignalEffs.png}
+			\includegraphics[width=\textwidth]{graphs/D23hSignalEffs.pdf}
 			\caption{Metrics for $D \to hhh$ signal.}
@@
 		\begin{subfigure}[b]{0.48\textwidth}
-			\includegraphics[width=\textwidth]{graphs/D23hBgEffs.png}
+			\includegraphics[width=\textwidth]{graphs/D23hBgEffs.pdf}
 			\caption{Metrics for $D \to hhh$ background.}
@@
 The (top) signal and (bottom) background uniformities are plotted as a function of the training stage of a given classifier, listed in the legends on the leftmost plots.
+Within each figure the leftmost plot is the SDE and the rightmost plot is the Theil uniformity metric.
 \label{fig:d2hhhsdetheil}}

### 2014-10-13 assets/paperdraft/ugb.tex
small textual edits

@@
 \end{equation}
-The {\em scores} $s$ are obtained for each event as sum of predictions of all elements in the series. 
+The {\em scores} $s$ are obtained for each event as the sum of predictions of all elements in the series. 
 At each stage in the gradient boosting process, a {\em regressor} (a decision tree in our case) is trained whose purpose is to decrease the loss.
@@
 
-Another approach is to include in the definition of the loss function some uniformity metric.
+Another approach is to include some uniformity metric in the definition of the loss function.
 Consider first the case where the data have been binned in \y.
@@
 where $\alpha$ is a real-valued parameter that is typically chosen to be small.
-The first term in Eq.~\ref{eq:L_ada_flat} penalizes non-uniformity, while that second term penalizes poor classification.
+The first term in Eq.~\ref{eq:L_ada_flat} penalizes non-uniformity, while the second term penalizes poor classification.
 We refer to this algorithm as uGB with flatness loss (uGBFL). 

### 2014-10-14 assets/paperdraft/ugb.tex
minor changes

@@
 \end{equation}
-where $a_{ij}$ are the elements of some matrix $A$\footnote{A natural choice is a square $n\times n$ matrix, but this is not required.}.
-For the case where $A$ is the identity matrix, the AdaBoost weighting procedure is recovered.
+where $a_{ij}$ are the elements of some square matrix $A$.  For the case where $A$ is the identity matrix, the AdaBoost weighting procedure is recovered.  
 Other choices of $A$ will induce non-local effects, {\em e.g.}, consider the sparse matrix $A_{\rm knn}$ given by
@@
 \end{equation} 
-For example, the loss function obtained from Eq.~\ref{eq:loss_general} using $A_{\rm knn}$, which we refer to as kNNAdaLoss and denote $L_{\rm knn}$, accounts for the score of each event's $k$ nearest neighbors and not just each event individually.
+For example, the loss function obtained from Eq.~\ref{eq:loss_general} using\footnote{A natural choice is a square $n\times n$ matrix, but this is not required.} 
+$A_{\rm knn}$, which we refer to as kNNAdaLoss and denote $L_{\rm knn}$, accounts for the score of each event's $k$ nearest neighbors and not just each event individually.
 The pseudo-residuals of $L_{\rm knn}$ are 

### 2014-10-14 assets/paperdraft/parts/metrics.tex
updated the paper draft
minor change

@@
 \[
-	p_{\rm peak} \thicksim \text{Beta}(\alpha \mean{m}, \abs{m - \mean{m}})
+	p_{\rm peak} \thicksim \text{Beta}(1 + \alpha f(m), 1), \qquad  f(m) = 5 \times e^{- 100  \abs{m - \overline{m}}^2}
 \]
@@
 \[
-	p_{\rm pit} \thicksim 1 - \text{Beta}(\alpha \mean{m}, \abs{m - \mean{m}})
+	p_{\rm pit} \thicksim 1 - \text{Beta}(1 + \alpha f(m), 1)
 \]
@@
 	\caption{Comparison of SDE and Theil uniformity measures for different values of the constant $\alpha$ for the peak and pit distributions.
-The Theil can clearly distinguish between these two while the SDE measure cannot.}
+The Theil can clearly distinguish between these two while the SDE measure cannot.\label{fig:pitvspeak}}
 \end{figure}
 
-From these figures we can see that SDE doesn't make any difference between these distributions, while Theil has lower value in the second case 
+From figure~\ref{fig:pitvspeak} we can see that SDE doesn't make any difference between these distributions, while Theil has lower value in the second case 
 which indicates that distribution is flatter.

### 2014-10-15 assets/paperdraft/parts/dataset.tex
added sde v alpha graph

@@
 In general, the ROC curves for the uniform-driven BDTs are not too different from AdaBoost.
+Figure~\ref{fig:dalitz_sde_v_alpha} shows how the uniformity of the selection efficiency depends on $\alpha$.
+As expected, as $\alpha$ is decreased the selection becomes more uniform.
 
@@
   \includegraphics[width=0.49\textwidth]{ROC_DP.pdf}
-  \includegraphics[width=0.49\textwidth]{ROC_DP_Alpha.pdf}
+  \includegraphics[width=0.49\textwidth]{ROC_DP_Alpha.pdf}\\
   \caption{\label{fig:dalitz_rocs} (left) ROC curves for classifier algorithms studied in this paper.
@@
 
+\begin{figure}[] 
+  \centering 
+  \includegraphics[width=0.49\textwidth]{sde_v_alpha.pdf}
+  \caption{\label{fig:dalitz_sde_v_alpha} Uniformity of the selection efficiency across the Dalitz plot, as measured using the so-called SDE metric described in detail in the appendix, {\em vs} $\alpha$ for uGBFL(bin).
+The dashed line indicates the SDE value for AdaBoost.
+Lower values of $\alpha$ produce more uniform selection efficiencies.}
+\end{figure}
+
 Figure~\ref{fig:dalitz_results} shows the efficiency obtained for each classifier {\em vs} distance from the a corner of the Dalitz-plot\footnote{The Dalitz-plot is essentially a triangle with three corners.
-Our defnition of this distance is ${\rm MIN}\left[(m(D_s)-m(\pi))^2 - m_{ij}^2\right]$, where $ij$ is 12, 13 and 23.}.
+Our definition of this distance is ${\rm MIN}\left[(m(D_s)-m(\pi))^2 - m_{ij}^2\right]$, where $ij$ is 12, 13 and 23.}.
 The AdaBoost algorithm, as expected, produces a much lower efficiency in the interesting corner regions.

### 2014-10-24 IPythonWorkflow/reports.py
nothing (test commit)

@@ -211,6 +211,11 @@ class Predictions(object):
             pylab.legend(loc="lower left")
         return self
 
+    def root_roc(self):
+        # TODO implement
+        pass
+        # proba_on_stages = pandas.DataFrame(self._get_stages())
+
     def prediction_pdf(self, stages=None, histtype='step', bins=30, show_legend=False):
         proba_on_stages = pandas.DataFrame(self._get_stages(stages))
         for stage_name, proba_on_stage in proba_on_stages.iterrows():
@@ -308,7 +313,6 @@ class Predictions(object):
         pylab.legend(loc='upper center', bbox_to_anchor=(0.5, 1.00), ncol=3, fancybox=True, shadow=True)
         if return_data:
             return result
-
 
     def theil_curves(self, uniform_variables, target_efficiencies=None, n_bins=20, label=1, step=3, return_data=True):
         mask = self.y == label

### 2014-11-09 hep_ml/toymc.py
minor fix in toymc

@@ -214,7 +214,7 @@ def compare_toymc(data, clustering_features=None, integer_features=None):
     for i, column in enumerate(data.columns, 1):
         pylab.subplot(n_rows, n_cols, i)
         pylab.title(column)
-        pylab.hist([data[column], toy_data[column]], histtype='step', bins=20)
+        pylab.hist([data[column].values, toy_data[column].values], histtype='step', bins=20)
     pylab.show()
 
     print_title("Means and std")

### 2014-11-10 assets/presentation/presentation.tex
first version of presentation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%   PACKAGES AND THEMES
%----------------------------------------------------------------------------------------
\batchmode
\documentclass{beamer}

\mode<presentation> {

% The Beamer class comes with a number of default slide themes
% which change the colors and layouts of slides. Below this is a list
% of all the themes, uncomment each in turn to see what they look like.

%\usetheme{default}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg} 
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

% As well as themes, the Beamer class has a number of color themes
% for any slide theme. Uncomment each of these in turn to see how it
% changes the colors of your current slide theme.

%\usecolortheme{albatross}
\usecolortheme{beaver}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dolphin}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{lily}
%\usecolortheme{orchid}
%\usecolortheme{rose}
%\usecolortheme{seagull}
%\usecolortheme{seahorse}
%\usecolortheme{whale}
%\usecolortheme{wolverine}

%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}

\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{epsf}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}

% \newcounter{eqn}
% \renewcommand*{\theeqn}{\alph{eqn})}
% \newcommand{\num}{\refstepcounter{eqn}\text{\theeqn}\;}

%----------------------------------------------------------------------------------------
%   TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Uniformity: metrics and classifiers]{New approaches for boosting to uniformity} % The short title appears at the bottom of every slide, the full title is only on the title page
\author[Alex Rogozhnikov]{
\underline{Alex Rogozhnikov}$^{a,b}$,
Aleksandar Bukva$^c$, 
Vladimir Gligorov$^d$,
Andrey Ustyuzhanin$^{b,e,f}$ and
Mike Williams$^g$
}

%\institute[SINP MSU] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space

\institute[]{
$^a$  Lomonosov Moscow State University, Moscow\\
$^b$  Yandex School of Data Analysis, Moscow\\
$^c$  Faculty of Physics, Belgrade \\
$^d$  Organisation Europ\'eenne pour la Recherche Nucl\'eaire (CERN), Geneva  \\
$^e$  Moscow Institute of Physics and Technology, Moscow\\
$^f$  Imperial College, London\\
$^g$  Massachusetts Institute of Technology, Cambridge 
\medskip
\\
\textit{alex.rogozhnikov@yandex.ru} % Your email address
}
\date{11 November, 2014}

\begin{document}
\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

%\begin{frame}
%\frametitle{Overview} % Table of contents slide, comment this block out to remove it
%\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
%\end{frame}

%   PRESENTATION SLIDES

\def\knn{\text{knn}}
\def\knni{\text{knn}(i)}
\newcommand{\abs}[1]{\left\vert#1\right\vert}
\newcommand{\sgn}{\operatorname{sgn}}

\def\bineff{\text{eff}_\text{bin}}
\def\binweight{\text{weight}_\text{bin}}
\def\globaleff{\text{eff}}
\def\SDE{\text{SDE}}
\def\bin{\text{bin}}
\def\theil{\text{Theil}}
\def\score{\text{score}}
\def\knn{\text{knn}}
\def\FL{\text{FL}}
\def\AdaLoss{L_\text{ada}}
\def\knnAdaLoss{L_\text{knn-ada}}
\def\generalAda{L_\text{general}}

\begin{frame}
    \frametitle{Outline}
    \begin{itemize}
    \item What is uniformity (of predictions)? 
    \item How to measure it? (metric functions)
    \item How to achieve it? (classifiers proposed)
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Uniformity}
    In particle physics, apart from optimizing some FOM of classifier (BDT, ANN), 
    there are cases when we need to have uniformity of predictions
    \begin{itemize}
    \item Dalitz-plot analysis (or any angular or amplitude analysis)

    \item search for a new particle (not to get fake peak)

    \item sensitivity for new signal 
        in wide range of vars (mass, lifetime ...)

    \end{itemize}

    \textbf{Uniform variables} --- variables, along which uniformity of selection is desired (Dalitz variables, mass variable).

    Typical solution: choose such features which don't give an ability to reconstruct 'mass' (or other selected 'uniform variables').
\end{frame}

\begin{frame}
    \frametitle{What is uniformity?}
    Predictions of some classifier are called \textit{uniform} in variables $var_1, \dots, var_n$ if prediction and set of this variables is \textit{statistically independent}.

    This (and only this) guarantees that any cut of prediction of classifier will produce the same efficiency in every region over $var_1, \dots, var_n$

    % definition of uncorrelated variables
    \pause
    % picture with absolutely uniform predictions
    % left bottom right top
	% picture with skew

    \begin{figure}
        \centering
        \begin{subfigure}[b]{0.48\textwidth}
            \includegraphics[width=\textwidth, height=4cm]{img/uniform.pdf}
            \caption{Uniform predictions}
        \end{subfigure}
        \begin{subfigure}[b]{0.48\textwidth}
            \includegraphics[width=\textwidth, height=4cm]{img/nonuniform.pdf}
            \caption{Non-uniform}
        \end{subfigure}
    \end{figure}

\end{frame}

\begin{frame}
    \frametitle{Desirable properties of metrics}
    The metric should ...
    \begin{itemize}
    \item not depend strongly on the number of events used to test uniformity

    \item not depend on the total weight

    \item depend on order of predictions, not the exact values of predictions \\
        (example: Pearson correlation does not satisfy this property)

    \item be stable against free parameters (number of bins, $k$ in knn)
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Similarity-based approach}
    Idea: uniformity means that distribution of predictions in every bin is equal. 
    \bigskip \\ 
    {\small
  	Let's compare the global distribution (blue hist) 
  	with distibution in one bin (yellow hist).
  	Yellow rectangle shows the events in selected bin over mass.
  	}

    \includegraphics[width=0.5\textwidth, height=4cm]{img/uniform_hist.pdf}
    \includegraphics[width=0.5\textwidth, height=4cm]{img/nonuniform_hist.pdf}

\end{frame}

\begin{frame}[t]
    \frametitle{Similarity-based approach}

    Let $F(x)$ -- cdf of all predictions, $F_{\bin}(x)$ --- cdf of predictions in bin over mass. Hereinfter $\binweight = \frac{\text{weight of events in bin}}{\text{weight of all events}}$ 
    \bigskip \\

    % $\binweight$ --- weights of bins.

    Kolmogorov-Smirnov measure (uninformative)
    \[
    	\sum_{\bin} \binweight \max_x \abs{F_{\bin}(x) - F(x)},
    \]
    Cram\'er--von Mises similarity 
    \[
    	 \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x)
    \]
\end{frame}

\begin{frame}
    \frametitle{Cut-based approach (1/2)}
    Select some set of efficiencies (in examples: 0.1, 0.3, 0.5, 0.7, 0.9), 
    for each one can compute global cut and look at efficincies in each bin:

    \includegraphics[width=0.5\textwidth, height=3.8cm]{img/uniform.pdf}
    \includegraphics[width=0.5\textwidth, height=3.8cm]{img/uniform_effs.pdf}

    \includegraphics[width=0.5\textwidth, height=3.8cm]{img/nonuniform.pdf}
    \includegraphics[width=0.5\textwidth, height=3.8cm]{img/nonuniform_effs.pdf}

\end{frame}

\begin{frame}
    \frametitle{Cut-based approach (2/2)}
    Standard deviation of efficiency
    \[
	\SDE^2(\globaleff) = 
		\sum_{\bin} \binweight \times \left(\bineff - \globaleff \right)^2 
	\]
	\[
	\SDE^2  =  \frac{1}{k} 
	\sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
		\text{SDE}^2(\globaleff)
	\] 
	Theil index of $x_1, \dots, x_n$
	\[
		\theil = \frac{1}{N} \sum_i \frac{x_i}{<x>} \ln{\frac{x_i}{<x>}}, 
	\]
    Theil index of efficiency
	\[
	\theil(\globaleff) = \sum_\bin \binweight \; \frac{\bineff}{\globaleff} \; \ln{\frac{\bineff}{\globaleff}}
	\]
    \[
    \theil  =  \frac{1}{k} 
    \sum_{\globaleff \in [\globaleff_1 \dots \globaleff_k] }  
        \theil(\globaleff).
    \]
\end{frame}

\begin{frame}
    \frametitle{Example}

    {
    \centering

    \includegraphics[width=0.9\textwidth]{img/metrics_comp1.png}

    \begin{figure}
    \centering

    \includegraphics[width=0.5\textwidth]{img/metrics_comp2.png}
    \end{figure}
    }

\end{frame}
% TODO demonstration of metrics

\begin{frame}
    \frametitle{Summary on metrics}
    \begin{enumerate}
        \item two basic approaches were introduced (distribution-based and cut-based)
        \item despite their difference, the results obtained with metrics proposed are \textbf{similar}.
        \item for higher dimensions: $k$nn modifications of metrics are available 
        {\small (instead of binning over uniform variables, we can compute nearest neighbours in the space of uniform variables).}
    \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{Boosting to uniformity}
    % TODO link
    Previuos work: uBoost \\
    {\small J. Stevens and M. Williams,
        {\em uBoost: A boosting method for producing uniform selection efficiencies from multivariate classifiers}, 
        JINST {\bf 8}, P12013 (2013). [arXiv:1305.7248]
    }
    \bigskip \\
    The classifiers we propose alter the \textbf{boosting} procedure as well
\end{frame}

\begin{frame}[t]
    \frametitle{Boosting: \textit{knn}AdaBoost}
    Usual AdaBoost reweighting procedure ($p_i$ is prediction of last classifier):
    \[
        w_i' = w_i \times \exp[-y_i \, p_i],
    \]

    \textit{knn}AdaBoost uses mean of predictions of neighbours 
    \[
        w_i = w_i \times \exp[-y_i \, \dfrac{1}{k} \sum_{j \in \knni} p_j]
    \]
    (neighbours are of the same class).

    Thus boosting focuses not on the events that were poorly classified, but on the regions with poor classification.
\end{frame}

\begin{frame}[t]
    \frametitle{Boosting: Gradient Boosting with \textit{knn}AdaLoss (1/2) }
    Gradient boosting on trees is widely used algorithm, 
    it's built upon decision tree regressors with usage of some loss function.

    Usual AdaLoss:
    \[
        \AdaLoss = \sum_{i \in \text{events}} w_i \times \exp [- \score_i \, y_i] 
    \]
    Pseudo-residual of AdaLoss:
    \[
        -\dfrac{\partial \, \AdaLoss}{\partial \, \score_i} = w_i \, y_i \exp[- \score_i \, y_i],
    \]

    \textit{knn}AdaLoss:
    \[
        \knnAdaLoss = \sum_{i \in events} \exp[-y_i \times \sum_{j \in \text{knn}(i)} \score_j],
    \]

\end{frame}

\begin{frame}[t]
    \frametitle{Boosting: Gradient Boosting with \textit{knn}AdaLoss (2/2)}
    \textit{knn}AdaLoss:
    \[
        \knnAdaLoss = \sum_{i \in events} \exp[-y_i \times \sum_{j \in \text{knn}(i)} \score_j],
    \]
    It can be written as particular case of:
    \[
        \generalAda = \sum_i \exp [- y_i \sum_j a_{ij} \, \score_j ],
    \]
    \[
        a_{ij} = 
        \begin{cases} 
            1, & j \in \knn(i), \text{ events $i$ and $j$ belong to the same class} \\
            0, & \text{otherwise},
        \end{cases}
    \]

    This is one particular choice of $a_{ij}$; \\
    in general case matrix $a_{ij}$ even may be non-square.

\end{frame}

\begin{frame}[t]
    \frametitle{Boosting: Gradient Boosting with FlatnessLoss (uGBFL)}
    CvM measure of non-uniformity:
    \[
        \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x),
    \]
    Let's modify this function:
    \[
        \FL = \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p {\textcolor{red}{dx}}
    \]
    so that it becomes differentiable
    \[
        \dfrac{\partial} {\partial \, \score_i} \FL
        \cong 
        w_i \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1}
        \sgn[F_{\bin(i)}(x) - F(x)]
        \Bigg|_{x=\score_i}
    \]
\end{frame}

\begin{frame}
    \frametitle{Boosting: Gradient Boosting with FlatnessLoss (uGBFL)}

    FL doesn't take into account the quality of predictions, only uniformity.

    So what we use in practice is linear combination of FlatnessLoss and AdaLoss:
    \[
        \text{loss} = \FL + \alpha \, \AdaLoss
    \]

    First one penalizes non-uniformity, second one --- poor predictions, 
    $\alpha$ is usually taken small.

\end{frame}

% \begin{frame}[t]
%     \frametitle{Crash test}

% \end{frame}

\begin{frame}[t]
    \frametitle{Tests on Dalitz data}
    Testing on dataset from paper about uBoost

    \center
    \includegraphics[width=0.6\textwidth, height=3cm]{img/dalitz_distr.png}

    \begin{figure}
        \centering
        \begin{subfigure}[b]{0.48\textwidth}
            \includegraphics[width=\textwidth, height=4cm]{img/dalitz_roc_stage.png}
            \caption{ROC vs \#trees \\ 
             \; (greater is better)}
        \end{subfigure}
        \begin{subfigure}[b]{0.48\textwidth}
            \includegraphics[width=\textwidth, height=4cm]{img/dalitz_sde.png}
            \caption{SDE vs \#trees \\ 
             \; (less is better)}
        \end{subfigure}

    \end{figure}

\end{frame}

\begin{frame}[t]
    \frametitle{Tradeoff uniformity vs quality}
    In uGBFL we can choose different values of $alpha$ thus adjusting quality/uniformity.

    \begin{figure}
        \centering
        \begin{subfigure}[b]{0.32\textwidth}
            \includegraphics[width=\textwidth, height=4cm]{img/trade_roc.png}
            \caption{ROC curves \\  
            \; $ \;$
            }
        \end{subfigure}
        \begin{subfigure}[b]{0.32\textwidth}
            \includegraphics[width=\textwidth, height=4cm]{img/trade_roc_stage.png}
            \caption{ROC vs \#trees \\ 
             \; (greater is better)}
        \end{subfigure}
        \begin{subfigure}[b]{0.32\textwidth}
            \includegraphics[width=\textwidth, height=4cm]{img/trade_sde.png}
            \caption{SDE vs \#trees \\ 
              \;  (less is better)
            }
        \end{subfigure}
    \end{figure}

\end{frame}

\begin{frame}[t]
    \frametitle{Summary on classifiers}
    New classifiers
    \begin{itemize}
        \item Faster (than uBoost)
        \item Introduces classifiers can target at uniformity in \textit{both} signal and bck
        \item $knn$AdaBoost and uGB + knnAdaLoss can be easily implemented, but don't seem to produce good uniformity
        \item uGBFL is highly tunable and proved to be able to fight severe correlation
    \end{itemize}

\end{frame}

\begin{frame}

    Read: \\
    \url{http://arxiv.org/abs/1410.4140} 
    \bigskip
    \\

    Try out (python implementation): \\
    \url{https://github.com/anaderi/lhcb_trigger_ml}

\end{frame}

\begin{frame}
\Huge{\centerline{Q\&A}}
\end{frame}

\begin{frame}
    \Huge{\centerline{Backup}}
\end{frame}

\begin{frame}[t]
    \frametitle{title}

\end{frame}

% \input{backup.tex}
%------------------------------------------------

\end{document}

### 2014-11-11 hep_ml/commonutils.py
fixed description

@@ -183,7 +183,7 @@ def compute_bdt_cut(target_efficiency, y_true, y_pred, sample_weight=None):
     """Computes cut which gives fixed efficiency.
     :type target_efficiency: float from 0 to 1 or numpy.array with floats in [0,1]
     :type y_true: numpy.array, of zeros and ones, shape = [n_samples]
-    :type y_pred: numpy.array, prediction probabilities returned by classifier, shape = [n_samples, 2]
+    :type y_pred: numpy.array, prediction probabilities returned by classifier, shape = [n_samples]
     """
     assert len(y_true) == len(y_pred), "different size"
     signal_proba = y_pred[y_true > 0.5]

### 2014-11-11 assets/presentation/presentation.tex

@@
     \item What is uniformity (of predictions)? 
-    \item How to measure it? (metric functions)
+    \item How to measure it? 
     \item How to achieve it? (classifiers proposed)
@@
     In particle physics, apart from optimizing some FOM of classifier (BDT, ANN), 
-    there are cases when we need to have uniformity of predictions
+    there are cases when we want to have uniformity of predictions
     \begin{itemize}
@@
     \item sensitivity for new signal 
-        in wide range of vars (mass, lifetime ...)
+        in wide range of mass (lifetime ...), \\
+        ({\small i.e. train only one classifier, not separate for each mass})
 
     \end{itemize}
+    $ $
 
     \textbf{Uniform variables} --- variables, along which uniformity of selection is desired (Dalitz variables, mass variable).
-
+    \bigskip \\
     Typical solution: choose such features which don't give an ability to reconstruct 'mass' (or other selected 'uniform variables').
@@
     \frametitle{What is uniformity?}
+
     Predictions of some classifier are called \textit{uniform} in variables $var_1, \dots, var_n$ if prediction and set of this variables is \textit{statistically independent}.
@@
 \begin{frame}
-    \frametitle{Desirable properties of metrics}
-    The metric should ...
-    \begin{itemize}
-    \item not depend strongly on the number of events used to test uniformity
+    \frametitle{Boosting to uniformity}
+    % TODO link
+    Previous work: \textbf{uBoost} \\
+    {\small J. Stevens and M. Williams,
+        {\em uBoost: A boosting method for producing uniform selection efficiencies from multivariate classifiers}, 
+        JINST {\bf 8}, P12013 (2013). [arXiv:1305.7248]
+    }
+    \bigskip \\
+    % The classifiers we propose alter the \textbf{boosting} procedure as well
 
-    \item not depend on the total weight
+    \center
+    \includegraphics[width=0.7\textwidth]{img/uboost.png}
 
-    \item depend on order of predictions, not the exact values of predictions \\
-        (example: Pearson correlation does not satisfy this property)
+    In this talk I present alternate methods for producing BDTs with uniform selection efficiency.
+\end{frame}
 
-    \item be stable against free parameters (number of bins, $k$ in knn)
-    \end{itemize}
+\begin{frame}[t]
+    \frametitle{Boosting: Gradient Boosting with plain AdaLoss}
+    Gradient boosting on trees is widely used algorithm, 
+    it's built upon decision tree regressors with usage of some loss function.
+
+    Usual AdaLoss ($y_i = +1$ for signal, $-1$ for background):
+    \[
+        \AdaLoss = \sum_{i \in \text{events}} \exp [- \score_i \, y_i] 
+    \]
+
+    Gradient boosting tries to minimize the loss function by 'steepest descent'-like procedure.
+
+    Pseudo-residual of AdaLoss:
+    \[
+        -\dfrac{\partial \, \AdaLoss}{\partial \, \score_i} = y_i \exp[- \score_i \, y_i],
+    \]
+
+    We need to introduce some loss, which penalizes non-unifomity.
+
 \end{frame}
@@
 \begin{frame}
-    \frametitle{Similarity-based approach}
+    \frametitle{Similarity-based approach for measuring uniformity}
     Idea: uniformity means that distribution of predictions in every bin is equal. 
@@
 
-    Let $F(x)$ -- cdf of all predictions, $F_{\bin}(x)$ --- cdf of predictions in bin over mass. Hereinfter $\binweight = \frac{\text{weight of events in bin}}{\text{weight of all events}}$ 
+    Let $F(x) = \text{P}(prediction < x)$ --- cdf of all predictions, \\
+    $F_{\bin}(x) = \text{P}(prediction \; in \; \text{bin} < x)$ --- cdf of predictions in bin over mass. Define $\binweight = \frac{\text{weight of events in bin}}{\text{weight of all events}}$ 
     \bigskip \\
@@
     \]
+\end{frame}
+
+\begin{frame}[t]
+    \frametitle{Boosting: Gradient Boosting with FlatnessLoss (uGBFL)}
+    CvM measure of non-uniformity:
+    \[
+        \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x),
+    \]
+    Let's modify this function:
+    \[
+        \FL = \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p {\textcolor{red}{dx}}
+    \]
+    so that it becomes differentiable
+    \[
+        \dfrac{\partial} {\partial \, \score_i} \FL
+        \cong 
+        w_i \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1}
+        \sgn[F_{\bin(i)}(x) - F(x)]
+        \Bigg|_{x=\score_i}
+    \]
+\end{frame}
+
+\begin{frame}
+    \frametitle{Boosting: Gradient Boosting with FlatnessLoss (uGBFL)}
+
+    FL doesn't take into account the quality of predictions, only uniformity.
+
+    So what we use in practice is linear combination of FlatnessLoss and AdaLoss:
+    \[
+        \text{loss} = \FL + \alpha \, \AdaLoss
+    \]
+
+    First one penalizes non-uniformity, second one --- poor predictions, 
+    $\alpha$ is usually taken small.
+
+\end{frame}
+
+\begin{frame}[t]
+    \frametitle{Boosting: Gradient Boosting with \textit{knn}AdaLoss (uGBkNNAdaLoss) }
+    As another approach we define:\textit{kNN}AdaLoss:
+    \[
+        \knnAdaLoss = \sum_{i \in events} \exp[-y_i \times \sum_{j \in \text{knn}(i)} \score_j],
+    \]
+    It can be written as particular case of:
+    \[
+        \generalAda = \sum_i \exp [- y_i \sum_j a_{ij} \, \score_j ],
+    \]
+    \[
+        a_{ij} = 
+        \begin{cases} 
+            1, & j \in \knn(i), \text{ events $i$ and $j$ belong to the same class} \\
+            0, & \text{otherwise},
+        \end{cases}
+    \]
+
+    This is one particular choice of $a_{ij}$; \\
+    in general case matrix $a_{ij}$ even may be non-square.
+
+\end{frame}
+
+\begin{frame}[t]
+    \frametitle{\textit{knn}AdaBoost}
+    Usual AdaBoost reweighting procedure ($p_i$ is prediction of last classifier):
+    \[
+        w_i' = w_i \times \exp[-y_i \, p_i],
+    \]
+
+    \textit{knn}AdaBoost uses mean of predictions of neighbours 
+    \[
+        w_i' = w_i \times \exp[-y_i \, \dfrac{1}{k} \sum_{j \in \knni} p_j]
+    \]
+    (neighbours are of the same class).
+
+    Thus boosting focuses not on the events that were poorly classified, but on the regions with poor classification.
+\end{frame}
+
+% \begin{frame}[t]
+%     \frametitle{Boosting: Gradient Boosting with \textit{knn}AdaLoss }
+%     Gradient boosting on trees is widely used algorithm, 
+%     it's built upon decision tree regressors with usage of some loss function.
+
+%     Usual AdaLoss:
+%     \[
+%         \AdaLoss = \sum_{i \in \text{events}} w_i \times \exp [- \score_i \, y_i] 
+%     \]
+%     Pseudo-residual of AdaLoss:
+%     \[
+%         -\dfrac{\partial \, \AdaLoss}{\partial \, \score_i} = w_i \, y_i \exp[- \score_i \, y_i],
+%     \]
+%     \textit{knn}AdaLoss:
+%     \[
+%         \knnAdaLoss = \sum_{i \in events} \exp[-y_i \times \sum_{j \in \text{knn}(i)} \score_j],
+%     \]
+
+% \end{frame}
+
+\begin{frame}[t]
+    \frametitle{Example analysis: distibutions}
+
+    \begin{figure}[] 
+          \centering 
+          \includegraphics[width=0.49\textwidth]{../paperdraft/graphs2/DP_sig.pdf}
+          \includegraphics[width=0.49\textwidth]{../paperdraft/graphs2/DP_bkgd.pdf}
+    \end{figure}
+    Dalitz-plot distributions for (left) signal and (right) background for the $D_s^\pm\to\pi^+\pi^-\pi^\pm$.
+The three pions are labeled here as 1, 2 and 3 and ordered according to increases momentum.
+\end{frame}
+
+\begin{frame}[t]\frametitle{Abbreviations}
+    \begin{table}
+      \begin{center}
+        \caption{\label{tab:algs} Description of uniform boosting algorithms.}
+        \begin{tabular}{c|c}
+          %\toprule
+          \hline
+          Name & Description \\
+          \hline
+          %\midrule
+            uBoost & algorithm introduced in [arXiv:1305.7248]\\
+          \hline
+          %\midrule 
+            kNNAda & AdaBoost modification using averaging over knn \\
+            uGBkNN & gradient boost using kNNAdaLoss loss function \\
+            uGBFL(bin) & gradient boost using flatness loss $+ \alpha$ AdaLoss  \\
+            uGBFL(kNN) & same except kNN events are used rather than bins \\
+          \hline
+          %\bottomrule
+        \end{tabular}
+      \end{center}
+    \end{table}
+
+\end{frame}
+
+\begin{frame}[t]
+    \frametitle{Example analysis}
+    \begin{figure}[] 
+        \centering 
+        \includegraphics[width=0.49\textwidth]{../paperdraft/graphs2/ROC_DP.pdf}
+        \includegraphics[width=0.49\textwidth]{../paperdraft/graphs2/DP_compare.pdf}
+        \\
+    \end{figure}
+
+    (left) ROC curves for classifiers. \\
+    (right) Efficiency {\em vs} distance to a corner of the Dalitz-plot.  An arbitrary working point of 50\% integrated efficiency is displayed. 
+    \bigskip \\
+    AdaBoost produces a much lower efficiency near corners of Dalitz plot, while uGBFL and uBoost are consistent with flat.
+\end{frame}
+
+\begin{frame}[t]
+    \frametitle{Example analysis, tradeoff }
+    Increasing of $\alpha$ parameter in uGBFL results in better quality of classification, but less uniformity of predictions.
+
+    \begin{figure}[] 
+        \centering 
+        \includegraphics[width=0.49\textwidth]{../paperdraft/graphs2/sde_v_alpha.pdf}
+        \includegraphics[width=0.49\textwidth]{../paperdraft/graphs2/ROC_DP_Alpha.pdf}
+    \end{figure}
+    (left) SDE for uGBFL(bin) for different values of $\alpha$, \\
+    SDE = Standard Deviation of Efficiency (0 = perfectly uniform).
+    \medskip    \\
+    (right) ROC curves for uGBFL(bin) for different values of $\alpha$.
+
+\end{frame}
+
+\begin{frame}[t]
+    \frametitle{Example analysis 2: training on mass sideband}
+
+    \begin{figure}[] 
+        \centering 
+        \includegraphics[width=0.49\textwidth]{../paperdraft/graphs2/mD_train.pdf}
+        \includegraphics[width=0.49\textwidth]{../paperdraft/graphs2/MD_eff.pdf}
+    \end{figure}
+
+    (left) Signal and (filled) background samples used in training.  \\
+    (right) Background mis-identification {\em vs} $D$ candidate mass for an arbitrary working point of 10\% background mis-identification in the training region $1.75 < m(D) < 1.85$~GeV is displayed.
+\end{frame}
+
+% \begin{frame}[t]
+%     \frametitle{Tests on Dalitz data}
+%     Testing on dataset from paper about uBoost
+
+%     \center
+%     \includegraphics[width=0.6\textwidth, height=3cm]{img/dalitz_distr.png}
+
+%     \begin{figure}
+%         \centering
+%         \begin{subfigure}[b]{0.48\textwidth}
+%             \includegraphics[width=\textwidth, height=4cm]{img/dalitz_roc_stage.png}
+%             \caption{ROC vs \#trees \\ 
+%              \; (greater is better)}
+%         \end{subfigure}
+%         \begin{subfigure}[b]{0.48\textwidth}
+%             \includegraphics[width=\textwidth, height=4cm]{img/dalitz_sde.png}
+%             \caption{SDE vs \#trees \\ 
+%              \; (less is better)}
+%         \end{subfigure}
+
+%     \end{figure}
+ % \end{frame}
+
+% \begin{frame}[t]
+%     \frametitle{Tradeoff uniformity vs quality}
+%     In uGBFL we can choose different values of $alpha$ thus adjusting quality/uniformity.
+
+%     \begin{figure}
+%         \centering
+%         \begin{subfigure}[b]{0.32\textwidth}
+%             \includegraphics[width=\textwidth, height=4cm]{img/trade_roc.png}
+%             \caption{ROC curves \\  
+%             \; $ \;$
+%             }
+%         \end{subfigure}
+%         \begin{subfigure}[b]{0.32\textwidth}
+%             \includegraphics[width=\textwidth, height=4cm]{img/trade_roc_stage.png}
+%             \caption{ROC vs \#trees \\ 
+%              \; (greater is better)}
+%         \end{subfigure}
+%         \begin{subfigure}[b]{0.32\textwidth}
+%             \includegraphics[width=\textwidth, height=4cm]{img/trade_sde.png}
+%             \caption{SDE vs \#trees \\ 
+%               \;  (less is better)
+%             }
+%         \end{subfigure}
+%     \end{figure}
+% \end{frame}
+
+\begin{frame}[t]
+    \frametitle{Summary on classifiers}
+    \small
+    New classifiers
+    \begin{itemize}
+        \item faster (than uBoost)
+        \item can target at uniformity in \textit{both} signal and bck
+        \item $knn$AdaBoost and gradient boosting with knnAdaLoss can be easily implemented, but don't seem to produce good uniformity
+        \item uGBFL is highly tunable and proved to be able to fight severe correlation
+    \end{itemize}
+    $ $ \\
+    uGBFL will be useful for:  
+    \begin{itemize}
+        \item Dalitz-plot or angular/amplitude analyses (as shown);
+        \item mass measurements since both signal and background can be kept unbiased near mass peak
+        \item searches for new particles since efficiency can be kept optimal for mass, lifetime values not used in the generated training samples
+        \item etc
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Summary on metrics}
+    \begin{itemize}
+        \item three metrics were introduced (details in the backup)
+        \item despite their difference, the results obtained with metrics proposed are \textbf{similar}.
+        \item for higher dimensions: $k$-nearest neighbours modifications of metrics are available 
+        {\small (instead of binning over uniform variables, we can compute nearest neighbours in the space of uniform variables).}
+    \end{itemize}
+
+Details: backup or article.
+\end{frame}
+
+\begin{frame}
+
+    \textbf{Read}: \\
+    \url{http://arxiv.org/abs/1410.4140} 
+    \bigskip
+    \\
+
+    \textbf{Try out (python implementation)}: \\
+    \url{https://github.com/anaderi/lhcb_trigger_ml}
+    \bigskip
+    \\
+
+    Feel free to email  \\
+    \url{mwill@mit.edu} \\
+    \url{alex.rogozhnikov@yandex.ru} \\
+    with any questions or comments.
+\end{frame}
+
+\begin{frame}
+\Huge{\centerline{Q\&A}}
+\end{frame}
+
+\begin{frame}
+    \Huge{\centerline{Backup}}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Desirable properties of metrics}
+    The metric should ...
+    \begin{enumerate}
+    \item not depend strongly on the number of events used to test uniformity
+
+    \item not depend on the total weight
+
+    \item depend on order of predictions, not the exact values of predictions \\
+        (example: Pearson correlation does not satisfy this property)
+
+    \item be stable against free parameters (number of bins, $k$ in knn)
+    \end{enumerate}
 \end{frame}
@@
     }
-
-\end{frame}
-% TODO demonstration of metrics
-
-\begin{frame}
-    \frametitle{Summary on metrics}
-    \begin{enumerate}
-        \item two basic approaches were introduced (distribution-based and cut-based)
-        \item despite their difference, the results obtained with metrics proposed are \textbf{similar}.
-        \item for higher dimensions: $k$nn modifications of metrics are available 
-        {\small (instead of binning over uniform variables, we can compute nearest neighbours in the space of uniform variables).}
-    \end{enumerate}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Boosting to uniformity}
-    % TODO link
-    Previuos work: uBoost \\
-    {\small J. Stevens and M. Williams,
-        {\em uBoost: A boosting method for producing uniform selection efficiencies from multivariate classifiers}, 
-        JINST {\bf 8}, P12013 (2013). [arXiv:1305.7248]
-    }
-    \bigskip \\
-    The classifiers we propose alter the \textbf{boosting} procedure as well
 \end{frame}
@@
 \begin{frame}[t]
-    \frametitle{Boosting: \textit{knn}AdaBoost}
-    Usual AdaBoost reweighting procedure ($p_i$ is prediction of last classifier):
-    \[
-        w_i' = w_i \times \exp[-y_i \, p_i],
-    \]
+    \frametitle{Timings of classifiers}
+    One drawback of the uBoost technique is that it has a high degree of computational complexity: 
+    while AdaBoost trains $M$ trees (a user-defined number), uBoost builds $100 \times M$ trees.  
+    The algorithms presented only build $M$ trees; however, the boosting involves some more complicated algorithms.  Training each of the $M$ trees scales as follows for $N$ training events:
+    \begin{itemize}
+           \item kNNAdaBoost: $O(k \times N)$;
+        \item uGBkNNknn: $O(k \times N)$ for $A_{\rm knn}$, and  
+        $O( \text{\#nonzero elements in the matrix})$ for arbitrary matrix $A$;
+        \item uGBFL(bin): $O(N \ln N)$;
+        \item uGBFL(kNN): $O(N \ln N + N k \ln k) $.
+    \end{itemize}
 
-    \textit{knn}AdaBoost uses mean of predictions of neighbours 
-    \[
-        w_i = w_i \times \exp[-y_i \, \dfrac{1}{k} \sum_{j \in \knni} p_j]
-    \]
-    (neighbours are of the same class).
-
-    Thus boosting focuses not on the events that were poorly classified, but on the regions with poor classification.
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Boosting: Gradient Boosting with \textit{knn}AdaLoss (1/2) }
-    Gradient boosting on trees is widely used algorithm, 
-    it's built upon decision tree regressors with usage of some loss function.
-
-    Usual AdaLoss:
-    \[
-        \AdaLoss = \sum_{i \in \text{events}} w_i \times \exp [- \score_i \, y_i] 
-    \]
-    Pseudo-residual of AdaLoss:
-    \[
-        -\dfrac{\partial \, \AdaLoss}{\partial \, \score_i} = w_i \, y_i \exp[- \score_i \, y_i],
-    \]
-
-    \textit{knn}AdaLoss:
-    \[
-        \knnAdaLoss = \sum_{i \in events} \exp[-y_i \times \sum_{j \in \text{knn}(i)} \score_j],
-    \]
+    Usually the training time for these new algorithms is within a factor of two the same as AdaBoost.
 
@@
 
-\begin{frame}[t]
-    \frametitle{Boosting: Gradient Boosting with \textit{knn}AdaLoss (2/2)}
-    \textit{knn}AdaLoss:
-    \[
-        \knnAdaLoss = \sum_{i \in events} \exp[-y_i \times \sum_{j \in \text{knn}(i)} \score_j],
-    \]
-    It can be written as particular case of:
-    \[
-        \generalAda = \sum_i \exp [- y_i \sum_j a_{ij} \, \score_j ],
-    \]
-    \[
-        a_{ij} = 
-        \begin{cases} 
-            1, & j \in \knn(i), \text{ events $i$ and $j$ belong to the same class} \\
-            0, & \text{otherwise},
-        \end{cases}
-    \]
-
-    This is one particular choice of $a_{ij}$; \\
-    in general case matrix $a_{ij}$ even may be non-square.
-
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Boosting: Gradient Boosting with FlatnessLoss (uGBFL)}
-    CvM measure of non-uniformity:
-    \[
-        \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p dF(x),
-    \]
-    Let's modify this function:
-    \[
-        \FL = \sum_{\bin} \binweight \int \abs{F_{\bin}(x) - F(x)}^p {\textcolor{red}{dx}}
-    \]
-    so that it becomes differentiable
-    \[
-        \dfrac{\partial} {\partial \, \score_i} \FL
-        \cong 
-        w_i \, p \,  \abs{F_{\bin(i)}(x) - F(x)}^{p-1}
-        \sgn[F_{\bin(i)}(x) - F(x)]
-        \Bigg|_{x=\score_i}
-    \]
-\end{frame}
-
-\begin{frame}
-    \frametitle{Boosting: Gradient Boosting with FlatnessLoss (uGBFL)}
-
-    FL doesn't take into account the quality of predictions, only uniformity.
-
-    So what we use in practice is linear combination of FlatnessLoss and AdaLoss:
-    \[
-        \text{loss} = \FL + \alpha \, \AdaLoss
-    \]
-
-    First one penalizes non-uniformity, second one --- poor predictions, 
-    $\alpha$ is usually taken small.
-
-\end{frame}
-
-% \begin{frame}[t]
-%     \frametitle{Crash test}
-
-% \end{frame}
-
-\begin{frame}[t]
-    \frametitle{Tests on Dalitz data}
-    Testing on dataset from paper about uBoost
-
-    \center
-    \includegraphics[width=0.6\textwidth, height=3cm]{img/dalitz_distr.png}
-
-    \begin{figure}
-        \centering
-        \begin{subfigure}[b]{0.48\textwidth}
-            \includegraphics[width=\textwidth, height=4cm]{img/dalitz_roc_stage.png}
-            \caption{ROC vs \#trees \\ 
-             \; (greater is better)}
-        \end{subfigure}
-        \begin{subfigure}[b]{0.48\textwidth}
-            \includegraphics[width=\textwidth, height=4cm]{img/dalitz_sde.png}
-            \caption{SDE vs \#trees \\ 
-             \; (less is better)}
-        \end{subfigure}
-
-    \end{figure}
-
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Tradeoff uniformity vs quality}
-    In uGBFL we can choose different values of $alpha$ thus adjusting quality/uniformity.
-
-    \begin{figure}
-        \centering
-        \begin{subfigure}[b]{0.32\textwidth}
-            \includegraphics[width=\textwidth, height=4cm]{img/trade_roc.png}
-            \caption{ROC curves \\  
-            \; $ \;$
-            }
-        \end{subfigure}
-        \begin{subfigure}[b]{0.32\textwidth}
-            \includegraphics[width=\textwidth, height=4cm]{img/trade_roc_stage.png}
-            \caption{ROC vs \#trees \\ 
-             \; (greater is better)}
-        \end{subfigure}
-        \begin{subfigure}[b]{0.32\textwidth}
-            \includegraphics[width=\textwidth, height=4cm]{img/trade_sde.png}
-            \caption{SDE vs \#trees \\ 
-              \;  (less is better)
-            }
-        \end{subfigure}
-    \end{figure}
-
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{Summary on classifiers}
-    New classifiers
-    \begin{itemize}
-        \item Faster (than uBoost)
-        \item Introduces classifiers can target at uniformity in \textit{both} signal and bck
-        \item $knn$AdaBoost and uGB + knnAdaLoss can be easily implemented, but don't seem to produce good uniformity
-        \item uGBFL is highly tunable and proved to be able to fight severe correlation
-    \end{itemize}
-
-\end{frame}
-
-\begin{frame}
-
-    Read: \\
-    \url{http://arxiv.org/abs/1410.4140} 
-    \bigskip
-    \\
-
-    Try out (python implementation): \\
-    \url{https://github.com/anaderi/lhcb_trigger_ml}
-
-\end{frame}
-
-\begin{frame}
-\Huge{\centerline{Q\&A}}
-\end{frame}
-
-\begin{frame}
-    \Huge{\centerline{Backup}}
-\end{frame}
-
-\begin{frame}[t]
-    \frametitle{title}
-
-\end{frame}
-
-% \input{backup.tex}
-%------------------------------------------------
-
 \end{document}

### 2014-11-13 hep_ml/commonutils.py
added docstrings, + changes in commnutils

@@ -76,7 +76,10 @@ def generate_sample(n_samples, n_features, distance=2.0):
 
 
 def check_sample_weight(y_true, sample_weight):
-    """Checks the weights, returns normalized version """
+    """Checks the weights, returns normalized version
+    :param y_true: numpy.array of shape [n_samples]
+    :param sample_weight: array-like of shape [n_samples] or None
+    :returns: numpy.array with weights of shape [n_samples]"""
     if sample_weight is None:
         return numpy.ones(len(y_true), dtype=numpy.float)
     else:
@@ -84,6 +87,17 @@ def check_sample_weight(y_true, sample_weight):
         assert len(y_true) == len(sample_weight), \
             "The length of weights is different: not {0}, but {1}".format(len(y_true), len(sample_weight))
         return sample_weight
+
+
+def check_uniform_label(uniform_label):
+    """ Convert to numpy.array
+    :param uniform_label: label or list of labels (examples: 0, 1, [0], [1], [0, 1])
+    :return: numpy.array (with [0], [1] or [0, 1])
+    """
+    if isinstance(uniform_label, numbers.Number):
+        return numpy.array([uniform_label])
+    else:
+        return numpy.array(uniform_label)
 
 
 def reorder_by_first(*arrays):
@@ -123,6 +137,15 @@ def train_test_split(*arrays, **kw_args):
 
 
 def weighted_percentile(array, percentiles, sample_weight=None, array_sorted=False, old_style=False):
+    """ Very close to numpy.precentile, but supports weights.
+    NB: percentiles should be in [0, 1]
+    :param array: numpy.array with data
+    :param percentiles: array-like with many percentiles
+    :param sample_weight: array-like of the same length as `array`
+    :param array_sorted: bool, if True, then will avoid sorting
+    :param old_style: if True, will correct output to be consistent with numpy.percentile.
+    :return: numpy.array with computed percentiles.
+    """
     array = numpy.array(array)
     percentiles = numpy.array(percentiles)
     sample_weight = check_sample_weight(array, sample_weight)
@@ -258,16 +281,6 @@ def memory_usage():
     return result
 
 
-def subdict(start_dict, keys=None):
-    """Returns the ordered dictionary from initial one which contains only listed keys """
-    if keys is None:
-        return OrderedDict(start_dict)
-    result = OrderedDict()
-    for key in keys:
-        result[key] = start_dict[key]
-    return result
-
-
 def indices_of_values(array):
[…]

### 2014-11-14 hep_ml/experiments/gradient_boosting.py
temporally added higgs algos to experiments

@@ -1,0 +1,683 @@
+from __future__ import division
+from __future__ import print_function
+from collections import OrderedDict
+
+import copy
+from itertools import islice
+import numbers
+import numpy
+import pandas
+import sklearn
+from scipy.special import expit, logit
+from sklearn.base import BaseEstimator, ClassifierMixin, TransformerMixin
+from sklearn.cross_validation import train_test_split, StratifiedKFold
+from sklearn.ensemble._gradient_boosting import _random_sample_mask
+from sklearn.ensemble.forest import RandomForestClassifier
+from sklearn.ensemble.gradient_boosting import LossFunction
+from sklearn.ensemble.weight_boosting import AdaBoostClassifier
+from sklearn.tree.tree import DecisionTreeRegressor, DTYPE
+from sklearn.utils.random import check_random_state
+from sklearn.utils.validation import check_arrays, column_or_1d, array2d
+from hep_ml.commonutils import check_sample_weight, generate_sample, map_on_cluster
+import hep_ml.commonutils
+from transformations import enhance_data, Shuffler, indices_of_values
+
+real_s = 691.988607712
+real_b = 410999.847322
+
+
+#region Functions for measurements
+
+def get_higgs_data(train_file = '/Users/axelr/ipython/datasets/higgs/training.csv'):
+    data = pandas.read_csv(train_file, index_col='EventId')
+    answers_bs = numpy.ravel(data.Label)
+    weights = numpy.ravel(data.Weight)
+    data = data.drop(['Label', 'Weight'], axis=1)
+    answers = numpy.zeros(len(answers_bs), dtype=numpy.int)
+    answers[answers_bs == 's'] = 1
+    return data, answers, weights
+
+
+def AMS(answers, predictions, sample_weight):
+    """ Predictions are classes """
+    assert len(answers) == len(predictions) == len(sample_weight)
+    predictions = column_or_1d(predictions)
+    total_s = numpy.sum(sample_weight[answers > 0.5])
+    total_b = numpy.sum(sample_weight[answers < 0.5])
+    s = numpy.sum(sample_weight[answers * predictions > 0.5])
+    b = numpy.sum(sample_weight[(1 - answers) * predictions > 0.5])
+    s *= real_s / total_s
+    b *= real_b / total_b
+    br = 10.
+    radicand = 2 * ( (s+b+br) * numpy.log(1.0 + s/(b+br)) - s)
+    if radicand < 0:
+        raise ValueError('Radicand is negative')
+    else:
+        return numpy.sqrt(radicand)
+
[…]

### 2014-11-14 hep_ml/experiments/transformations.py

@@ -1,0 +1,482 @@
+from __future__ import division
+from __future__ import print_function
+from collections import OrderedDict
+import copy
+
+import numpy
+import pandas
+import math
+from numpy.random import normal as randn
+from sklearn.utils.validation import check_random_state
+
+__author__ = 'Alex Rogozhnikov'
+
+
+
+#region Special functions
+
+
+def to_nan(data):
+    assert isinstance(data, pandas.DataFrame)
+    return data.replace(-999, numpy.nan)
+
+
+def from_nan(data):
+    assert isinstance(data, pandas.DataFrame)
+    return data.replace(numpy.nan, -999)
+
+
+def to_projections_(phi, eta, pt):
+    return pt * numpy.cos(phi), pt * numpy.sin(phi), pt * numpy.sinh(eta)
+
+
+def to_projections(df, prefix):
+    df[prefix+'px'], df[prefix+'py'], df[prefix+'pz'] = \
+        to_projections_(df[prefix+'phi'], df[prefix+'eta'], df[prefix+'pt'])
+
+
+def from_projections_(px, py, pz):
+    p_t = numpy.sqrt(px*px + py*py)
+    return numpy.arctan2(py, px), numpy.arcsinh(pz / p_t), p_t
+
+
+def from_projections(df, prefix):
+    df[prefix+'phi'], df[prefix+'eta'], df[prefix+'pt'] = \
+        from_projections_(df[prefix+'px'], df[prefix+'py'], df[prefix+'pz'])
+
+
+def test_projections(size=100):
+    triple = (randn(size=size), randn(size=size), randn(size=size))
+    res = to_projections_(*from_projections_(*triple))
+
+    for i in [0, 1, 2]:
+        assert numpy.allclose(res[i], triple[i])
+
+test_projections()
+
+
+def compute_energy(m, px, py, pz):
+    return numpy.sqrt(m*m + px*px + py*py + pz*pz)
+
+
+def compute_mass(E, px, py, pz):
+    return numpy.sqrt(E*E - px*px - py*py - pz*pz)
+
+
+def dot(df, prefix1, prefix2):
+    return df[prefix1 + "px"] * df[prefix2 + "px"] \
[…]

### 2014-11-14 hep_ml/experiments/losses.py
old losses moved to experiments/losses. Many changes, old FL was deleted.

@@ -1,21 +1,21 @@
-from __future__ import print_function
-from __future__ import division
+from __future__ import print_function, division, absolute_import
 
-from collections import defaultdict
+import numpy
 import numbers
+from collections import defaultdict
 
 import scipy.sparse as sparse
 from sklearn.base import BaseEstimator
 from sklearn.ensemble.forest import RandomForestRegressor
-from sklearn.ensemble.gradient_boosting import LogOddsEstimator
-import numpy
 from sklearn.metrics.pairwise import pairwise_distances
 from sklearn.neighbors.unsupervised import NearestNeighbors
 from sklearn.utils.validation import column_or_1d
 
-from .. import commonutils, reports, metrics
+from .. import commonutils
 from ..commonutils import check_sample_weight
-from ..ugradientboosting import AbstractLossFunction, AbstractMatrixLossFunction, compute_positions
+from hep_ml.commonutils import check_uniform_label
+from hep_ml.losses import AbstractFlatnessLossFunction, KnnFlatnessLossFunction
+from ..ugradientboosting import AbstractLossFunction, AbstractMatrixLossFunction
 
 
 __author__ = 'Alex Rogozhnikov'
@@ -296,7 +296,7 @@ class PairwiseKnnLossFunction(AbstractMatrixLossFunction):
         knn = self.knn
         if self.exclude_self:
             knn_indices = \
-                commonutils.computeKnnIndicesOfSameClass(self.uniform_variables, trainX, is_signal, knn+1)[:, 1:]
+                commonutils.computeKnnIndicesOfSameClass(self.uniform_variables, trainX, is_signal, knn + 1)[:, 1:]
         else:
             knn_indices = commonutils.computeKnnIndicesOfSameClass(self.uniform_variables, trainX, is_signal, knn)
 
@@ -405,10 +405,10 @@ class DistanceBasedKnnFunction(AbstractMatrixLossFunction):
 
         if self.row_normalize:
             from sklearn.preprocessing import normalize
+
             A = normalize(A, norm='l1', axis=1)
 
         return A, numpy.ones(A.shape[0])
-
 
 
 def exp_margin(margin):
@@ -416,226 +416,49 @@ def exp_margin(margin):
     return numpy.exp(numpy.clip(margin, -1e5, 2))
 
 
-class FlatnessLossFunction(AbstractLossFunction, BaseEstimator):
-    def __init__(self, uniform_variables, bins=10, uniform_label=1, power=2., ada_coefficient=1.,
[…]

### 2014-11-14 hep_ml/experiments/losses.py
updated experimental simpleKnnLoss

@@ -1,21 +1,16 @@
 from __future__ import print_function, division, absolute_import
 
 import numpy
-import numbers
-from collections import defaultdict
-
 import scipy.sparse as sparse
-from sklearn.base import BaseEstimator
 from sklearn.ensemble.forest import RandomForestRegressor
 from sklearn.metrics.pairwise import pairwise_distances
 from sklearn.neighbors.unsupervised import NearestNeighbors
-from sklearn.utils.validation import column_or_1d
 
 from .. import commonutils
 from ..commonutils import check_sample_weight
 from hep_ml.commonutils import check_uniform_label
-from hep_ml.losses import AbstractFlatnessLossFunction, KnnFlatnessLossFunction
-from ..ugradientboosting import AbstractLossFunction, AbstractMatrixLossFunction
+from hep_ml.losses import SimpleKnnLossFunction, KnnFlatnessLossFunction
+from hep_ml.ugradientboosting import AbstractMatrixLossFunction
 
 
 __author__ = 'Alex Rogozhnikov'
@@ -27,25 +22,24 @@ __author__ = 'Alex Rogozhnikov'
 # algorithm of generating A and w
 
 
-class SimpleKnnLossFunction(AbstractMatrixLossFunction):
-    def __init__(self, uniform_variables, knn=10, uniform_label=1, distinguish_classes=True, row_norm=1.):
+class ExperimentalSimpleKnnLossFunction(SimpleKnnLossFunction):
+    def __init__(self, uniform_variables, knn=10, uniform_label=1, row_norm=1., diagonal=0.):
         """A matrix is square, each row corresponds to a single event in train dataset, in each row we put ones
         to the closest neighbours of that event if this event from class along which we want to have uniform prediction.
         :param list[str] uniform_variables: the features, along which uniformity is desired
         :param int knn: the number of nonzero elements in the row, corresponding to event in 'uniform class'
         :param int|list[int] uniform_label: the label (labels) of 'uniform classes'
-        :param bool distinguish_classes: if True, 1's will be placed only for
+        :param int diagonal: float, A + diagonal * Identity is used.
         """
         self.knn = knn
-        self.distinguish_classes = distinguish_classes
         self.row_norm = row_norm
-        self.uniform_label = [uniform_label] if isinstance(uniform_label, numbers.Number) else uniform_label
-        AbstractMatrixLossFunction.__init__(self, uniform_variables)
[…]

### 2014-11-14 hep_ml/experiments/losses.py
returned distingish_classes param

@@ -23,7 +23,7 @@ __author__ = 'Alex Rogozhnikov'
 
 
 class ExperimentalSimpleKnnLossFunction(SimpleKnnLossFunction):
-    def __init__(self, uniform_variables, knn=10, uniform_label=1, row_norm=1., diagonal=0.):
+    def __init__(self, uniform_variables, knn=10, uniform_label=1, row_norm=1., diagonal=0., distinguish_classes=True):
         """A matrix is square, each row corresponds to a single event in train dataset, in each row we put ones
         to the closest neighbours of that event if this event from class along which we want to have uniform prediction.
         :param list[str] uniform_variables: the features, along which uniformity is desired
@@ -35,7 +35,8 @@ class ExperimentalSimpleKnnLossFunction(SimpleKnnLossFunction):
         self.row_norm = row_norm
         self.diagonal = diagonal
         self.uniform_label = check_uniform_label(uniform_label)
-        AbstractMatrixLossFunction.__init__(self, uniform_variables,)
+        self.distinguish_classes = distinguish_classes
+        AbstractMatrixLossFunction.__init__(self, uniform_variables)
 
     def compute_parameters(self, trainX, trainY):
         n_samples = len(trainX)
